### 2009-01-30 tex/somm.tex
Initial check-in of the code and latex file

\documentclass[aps,prd,twocolumn,amsmath,amssymb,floatfix,nofootinbib,10pt]{revtex4}
%\documentclass[aps,prd,amsmath,amssymb,floatfix,nofootinbib,12pt]{revtex4}

\usepackage{graphicx,epsfig}
\usepackage{amsmath,amssymb}
\usepackage{subfigure,epsfig}
%\usepackage{feynmp}
\usepackage{float}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}

\newcommand{\eg}{e.g.}
\newcommand{\Fermi}{\emph{Fermi}}
\newcommand{\VL}{Via Lactea}
\newcommand{\aquarius}{Aquarius}
\newcommand{\CDM}{CDM}
\newcommand{\LCDM}{\ensuremath{\Lambda}\CDM}
\newcommand{\NFW}{NFW}
\newcommand{\GNFW}{G\NFW}
\newcommand{\DM}{DM}
\newcommand{\SM}{SM}
\newcommand{\OM}{OM}
\newcommand{\mpl}{\ensuremath{m_{\mathrm{Pl}}}}
\newcommand{\somm}{\ensuremath{S}}
\newcommand{\mdm}{\ensuremath{m_{\chi}}}
\newcommand{\mv}{\ensuremath{m_{\phi}}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\Eqnname}{Equation}
\newcommand{\eqnname}{equation}
\newcommand{\Ngamma}{\ensuremath{N_{\gamma}}}
\newcommand{\Ngammai}{\ensuremath{N_{\gamma,i}}}
\newcommand{\Eth}{\ensuremath{E_{\mathrm{th}}}}
\newcommand{\sigmaannv}{\ensuremath{\langle\sigma v\rangle}}
\newcommand{\los}{los}
\newcommand{\order}{\ensuremath{\mathcal{O}}}
\newcommand{\sigv}{\ensuremath{\sigma_v}}
\newcommand{\lum}{\ensuremath{\mathcal{L}}}
\newcommand{\lumsmooth}{\ensuremath{\lum_{\mbox{{\footnotesize sm}}}}}
\newcommand{\lumsmoothc}{\ensuremath{\lum_{\mbox{{\footnotesize sm}},v=c}}}
\newcommand{\lumsmoothtoo}{\ensuremath{\overline{\mathcal{L}}}}
\newcommand{\lumsub}{\ensuremath{\lum_{\mbox{{\footnotesize sub}}}}}
\newcommand{\boost}{\ensuremath{B}}
\newcommand{\redr}{\ensuremath{\tilde{r}}}
\newcommand{\rhos}{\ensuremath{\rho_s}}
\newcommand{\rs}{\ensuremath{r_s}}
\newcommand{\dist}{\ensuremath{D}}
\newcommand{\norm}{\ensuremath{\mathcal{N}}}
\newcommand{\normnfw}{\ensuremath{\norm_1}}
\newcommand{\normeinasto}{\ensuremath{\norm_2}}
\newcommand{\rhominustwo}{\ensuremath{\rho_{-2}}}
\newcommand{\rminustwo}{\ensuremath{r_{-2}}}
\newcommand{\sommboost}{\ensuremath{\mathcal{S}}}
\newcommand{\clumpboost}{\ensuremath{\mathcal{C}}}
\newcommand{\alphaEinasto}{\ensuremath{\alpha}}
\newcommand{\Msol}{\ensuremath{M_{\odot}}}
\newcommand{\Mvir}{\ensuremath{M_{\mbox{{\footnotesize vir}}}}}
\newcommand{\Rvir}{\ensuremath{r_{\mbox{{\footnotesize vir}}}}}
\newcommand{\conc}{\ensuremath{c}}
\newcommand{\rhoc}{\ensuremath{\rho_{\mbox{{\footnotesize c}}}}}
\newcommand{\omegam}{\ensuremath{\Omega_{m}}}
\newcommand{\vcirc}{\ensuremath{v_{\mbox{{\footnotesize c}}}}}
\newcommand{\vmax}{\ensuremath{v_{\mbox{{\footnotesize max}}}}}
\newcommand{\rmax}{\ensuremath{r_{\mbox{{\footnotesize max}}}}}

%Journals
\newcommand{\aap}{Astron.~Astrophys.}
\renewcommand{\nat}{Nature}
\newcommand{\jhep}{J.~High.~Energy.~Phys.}
\newcommand{\astropartphys}{Astropart.~Phys.}
\newcommand{\mnras}{Mon.~Not.~Roy.~Astron.~Soc.}
\newcommand{\apjs}{Astrophys.~J.~Suppl.}
\newcommand{\apjl}{\apj}
\newcommand{\aj}{Astron.~J.}
\newcommand{\jcap}{J.~Cosmol.~Astropart.~Phys.}

\begin{document}

\title{Substructure Boosts to Dark Matter Annihilation from Sommerfeld Enhancement}
\author{Jo Bovy} \affiliation{Center for
Cosmology and Particle Physics, Department of Physics\\ New York
University, New York, NY 10003}

\date{\today}

\begin{abstract}
We discuss the implications of the Sommerfeld enhancement of the dark
matter annihilation cross section for the detection of dark matter
annihilation from subhalos in the Galactic halo. A robust prediction
of \CDM\ is that the Galactic halo is teeming with cold substructure,
and that the subhalos themselves contain even smaller substructure,
leading to an approximately self-similar distribution of substructure
all the way down to earth-mass halos. In addition to the boost to the
dark matter annihilation cross section from the high densities of
these subhalos, an additional boost due to Sommerfeld enhancement
results from the fact that they are kinematically cold. We take
subhalos from the large numerical $N$-body simulation \VL\ to study
the expected distribution of dark matter annihilation signals from
subhalos at a Sun-like position in a large dark matter halo hosting a
late-type galaxy. We also study the expected Sommerfeld enhanced
annihilation signal for the smallest, most-dark-matter-dominated, and
kinematically coldest dwarf spheroidal satellite galaxy Segue
1. Finally, we discuss the prospects for detecting annihilation
signals from microhalos.
\end{abstract}

%\keywords{}
\maketitle

\section{Introduction}

First two paragraphs straight from my paper with Glennys:

In the standard flat, Gaussian, adiabatic, and scale-invariant \LCDM\
paradigm $\sim\!80$ \% of the matter density of the Universe is in the
form of dark matter (\DM) \cite{2008arXiv0803.0547K}. The properties
of dark matter are largely unconstrained, although a weakly
interacting massive particle (WIMP) with a mass $\sim\!100$ GeV -- 1
TeV has many attractive features, including that it leads to a relic
density of dark matter that is remarkably close to the measured value
\cite{1996PhR...267..195J}. No interaction of dark matter either with
itself or with standard model (\SM) particles other than gravity has
ever been observed. However, at the present day we cannot exclude that
the dark sector is brimming with a rich phenomenology at
weak-interaction scales or lower.

One exciting possibility is that dark matter self-annihilates at a
level observable in the Galactic halo today. Detecting this
self-annihilation seems unlikely considering the standard
self-annihilation cross section --- $\sigmaannv_{\mathrm{ann}}
\lesssim 3 \times 10^{-26} $ cm$^3$ s$^{-1}$ for consistency with the
observed relic density. However, there are various mechanisms to boost
the annihilation signal to a value larger by a few orders of
magnitude, e.g., clumpiness of the dark halo
\cite{1993ApJ...411..439S,1999PhRvD..59d3506B,2008A&A...479..427L,2008ApJ...686..262K}
or a combination of coannihilations and Sommerfeld enhancement
\cite{2005PhRvD..72j3521P,2008arXiv0812.0360L}. The possibility of
detecting the \DM\ annihilation signal has attracted much attention
recently because of the anomalous excesses of high energy electrons
and positrons reported by two different experimental groups: the
Payload for Antimatter Exploration and Light-nuclei Astrophysics
(PAMELA) satellite, which saw a sharp upturn in the positron fraction
$e^+/(e^++e^-)$ from 10 -- 100 GeV \cite{Adriani:2008zr}, and the
Advanced Thin Ionization Calorimeter (ATIC) group, which reported an
excess in the combined number of electrons and positrons at energies
500 -- 800 GeV \cite{2008Natur.456..362C}. The PAMELA result on the
positron fraction confirms earlier results
\cite{1969ApJ...158..771F,1975ApJ...199..669B,1987ApJ...312..183M,1994ApJ...436..769G,Barwick:1997ig,Beatty:2004cy,Aguilar:2007yf}
and can be explained by \DM\ annihilation
\cite{2002PhLB..536..263K,2004PhRvD..69j3509H,2008arXiv0810.5344C}. There
is still a good chance that these anomalies can be obviated by
conventional astrophysical effects, e.g., pulsars
\cite{1995A&A...294L..41A,2008arXiv0810.1527H,2008arXiv0810.2784Y,2008arXiv0812.4457P}. Nevertheless,
a plethora of models has been proposed in the last few months
explaining the PAMELA and ATIC results in terms of \DM\ annihilation
into electron-positron pairs (e.g.,
\cite{2008arXiv0810.5557H,ArkaniHamed:2008qn,2008arXiv0811.0399F,2008arXiv0811.1555I,2008arXiv0811.3357C,2008arXiv0811.3641C,2008arXiv0812.2196A}),
a description which also naturally accounts for other experimental
anomalies, such as the ``WMAP haze''
\cite{Finkbeiner:2003im,Dobler:2007wv,Hooper:2007kb}, EGRET
observations of an excess of high-energy gamma-rays in the Galactic
center \cite{Strong:2005zx}, and recent HESS measurements of the flux
of very high-energy electrons \cite{2008arXiv0811.3894H}. Model
building is complicated by the fact that PAMELA did not observe a
similar excess in the ratio of anti-protons to protons
\cite{2008arXiv0810.4994A}, which means that the coupling of the \DM\
particles to quarks must somehow be suppressed.

Of course, the largest obstacle currently to modeling the PAMELA/ATIC
signal with \DM\ annihilation is the very large boost factors ($10^4$
-- $10^5$) with respect to the fiducial $\langle \sigma v
\rangle_{\mathrm{ann}} \approx 3 \times 10^{-26} $ cm$^3$ s$^{-1}$
needed to lead to an observable annihilation flux. The boost factors
from the clumpiness of the dark matter halo is expected to be small
\cite{2008A&A...479..427L,2008Natur.454..735D,2008ApJ...686..262K,2008Natur.456...73S},
typically of the order 1 -- 10. Therefore, a different mechanism must
be reponsible for the large boosts necessary. One such mechanism that
has received much attention lately is the Sommerfeld enhancement
\cite{sommerfeld31a,2003PhRvD..67g5014H,2004PhRvL..92c1303H,2005PhRvD..71f3528H,2005PhRvD..71a5007H,2006PhRvD..73e5004H,2008NuPhB.800..204C,2008JHEP...07..058M,2008arXiv0812.0559M,2008arXiv0812.0360L}. The
Sommerfeld enhancement is a generic effect present whenever there is
an attractive force acting between the dark matter particles, \eg, due
to a Yukawa interaction or a gauge interaction through vector bosons
\cite{ArkaniHamed:2008qn}. The main effect of this attractive force is
to enhance the annihilation cross section with a factor proportional
to $\beta^{-1}$ ($\beta \equiv v/c$), this is known as ``$1/v$''
enhancement. There are regions in the parameter space made up of the
dark matter mass \mdm, the force carrier mass \mv, and the force
``fine-structure constant'' $\alpha \equiv$ coupling$^2$/$4 \pi$ in
which the enhancement is near a resonance, where it approximately
behaves as $\beta^{-2}$. However, the enhancement levels off at a
certain $\beta$ due to the finite range of the attractive force ---
large ($\alpha \sim\!10^{-1}$ -- $10^{-3}$) attractive forces with
infinite range lead to a burst of \DM\ annihilation in the first \DM\
halos formed at $z$ $\sim$ 100 -- 200 which would lead to observable
effects incompatible with measurements of the diffuse extragalactic
gamma-ray background today and the cosmic microwave background
\cite{2008arXiv0810.3233K}.

BOVY: Say something here about clumpiness of the halo: much more than
the following: \cite{1998MNRAS.300..146G,johnston98a,2003MNRAS.339..834H,2008ApJ...688..277T,2002MNRAS.330..778W,2002MNRAS.330..792K,koposov,Koposov:2009ru}

\section{The Sommerfeld enhancement}

The simplest example of the Sommerfeld effect that contains all of the
generic features described in the introduction is that of the
Sommerfeld enhancement due to an attractive Yukawa type force mediated
by a spin-0 boson $\phi$ between the \DM\ particles, and we adopt that
description here. The Sommerfeld effect is a non-perturbative effect
in the sense that in the language of quantum field theory it comes
about from Feynman diagrams in which the force carrier is exchanged
many times before the annihilation actually happens, giving rise to so
called ``ladder diagrams''. However, we can study the Sommerfeld
effect in simple, non-relativistic quantum mechanics (see the appendix
in Ref.~\cite{ArkaniHamed:2008qn}). We are instructed to solve the
radial Schr{\"o}dinger equation (in natural units)
\begin{equation}\label{eq:sommdiff1}
\frac{1}{\mdm} \frac{\dd^2 \psi(r)}{\dd r^2} + \frac{\alpha}{r}e^{-\mv r} \psi(r) = -\mdm \beta^2 \psi(r)\, ,
\end{equation}
subject to the boundary condition $\dd\psi/\dd r = i \mdm \beta \psi$
as $r \rightarrow \infty$. Here $\mdm$ and $\mv$ are the mass of the
dark matter and the force carrier, respectively. Using the
substitution $r \rightarrow \alpha \mdm r$ this becomes
\begin{equation}\label{eq:sommdiff2}
\frac{\dd^2 \psi(r)}{\dd r^2} + \frac{1}{r}e^{-\mv r/(\alpha \mdm)} \psi(r) = -\frac{\beta^2}{\alpha^2} \psi(r)\, ,
\end{equation}
with boundary condition
\begin{equation}
\psi \propto e^{i\beta r/\alpha} \qquad \mbox{ as } r \rightarrow\infty\, .
\end{equation}
The Sommerfeld enhancement \somm\ is then given by
\begin{equation}
\somm = \frac{|\psi(\infty)|^2}{|\psi(0)|^2}\, .
\end{equation}
From \eqnname\ (\ref{eq:sommdiff2}) it can be seen that the Sommerfeld
enhancement is only a function of two variables, which can be chosen
as $\alpha/\beta$ and $\alpha \mdm/\mv$. The qualitative features of
\eqnname\ (\ref{eq:sommdiff2}) have been widely discussed before and
are easily derived, and, therefore, we will merely mention them here
without proof (see, \eg,
Refs.~\cite{ArkaniHamed:2008qn,2008arXiv0812.0360L}). When the kinetic
energy of the collision is much larger than the product of the force
carrier mass and the coupling, then the Yukawa interaction can be
approximated as an attractive Coulomb interaction, for which an
analytical solution in terms of hypergeometric functions exists. The
Sommerfeld enhancement is simply given by
\begin{equation}
\somm = \frac{\pi \alpha}{\beta} (1 - e^{-\pi\alpha/\beta})^{-1}\, .
\end{equation}
From this we see why the Sommerfeld enhancement is often referred to
as a ``$1/v$'' enhancement for small velocities $\beta \ll 1$. In the
limit of a vanishing interaction ($\alpha/\beta \ll 1$) the Sommerfeld
enhancement disappears as $\somm \rightarrow 1$. The $1/v$ behavior,
which diverges for small $\beta$ and leads to an overproduction of
high-energy photons in the early Universe \cite{2008arXiv0810.3233K},
levels off due to the finite range of the Yukawa interaction at a
velocity $\beta \approx \mv/\mdm$ and settles on a value $\somm \approx
\alpha \mdm/\mv$. However, at some parameter combinations the Yukawa
potential has bound states, at which there are large boosts to the
annihilation cross section which go approximately as $\somm \propto
\beta^{-2}$. These resonance regions, however, are cut-off at small
$\beta$ due to the finite lifetime of the bound state. We expect these
three different behaviors ($1/v$, saturation at low $v$, and resonance
effects due to the existence of bound states) to be generic among
different models of the Sommerfeld enhancement (see also the
discussion in the appendix of Ref.~\cite{ArkaniHamed:2008qn}).

\Eqnname\ (\ref{eq:sommdiff2}) can also be numerically integrated
efficiently, by integrating the solution $\psi(r) = e^{i\beta
r/\alpha}$ inwards from a radius satisfying
\begin{equation}
\frac{1}{r}e^{-\mv r/(\alpha \mdm)} \ll \frac{\beta^2}{\alpha^2} \, .
\end{equation}
In the Coulomb regime ($\mdm \beta^2 \gg \alpha \mv$) this means that
the starting value should be such that $r \gg \alpha^2/\beta^2$;
otherwise we should start at $r \gg \alpha \mdm/\mv$. Figure
\ref{fig:sommb34} shows the Sommerfeld enhancement in the parameter plane
made up off $\alpha$ and $\mv/\mdm$. All of the qualitative features
we discussed above are present: In the Coulomb regime (lower part of
the two panels) we see that the enhancement scales as $\alpha$,
leveling off at $\somm \approx 1$ when $\alpha$ becomes small; The
fact that the upper left corner of both panels is approximately the
same shows that the Sommerfeld enhancement saturates at $\beta \approx
\mv/\mdm$; The sharp diagonal lines in the upper right of both panels
is the resonance region, in which the enhancement is sharply
peaked. Because of the scaling properties of \eqnname\
(\ref{eq:sommdiff2}) shifting $\log\beta$ is equivalent to shifting
both $\log\alpha$ and $\log \mv/\mdm$ by the same amount. Therefore,
the $\beta$ behavior of the enhancement can also be read off from
these figures along diagonals. Inspection of the resonance diagonals
in the second panel of Figure \ref{fig:sommb34} shows that the enhancement
grows as $\beta^{-2}$ in these regions.

\begin{figure}[t]
\centering
\includegraphics{2d-3.eps}
\\
\includegraphics{2d-4.eps}
\caption{Sommerfeld enhancement as a function of the particle physics
parameters for two different values of the relative velocity
$\beta$. Because of the scaling properties of \eqnname\
(\ref{eq:sommdiff2}) the behavior of the Sommerfeld enhancement as a
function of $\log_{10}\beta$ can actually be read off along diagonals
in the plane.}%\
\label{fig:sommb34}%
\end{figure}

Of course, when considering the Sommerfeld enhancement of a halo of a
certain velocity dispersion $\sigv$ we need to average the Sommerfeld
enhancement over the velocity distribution of that halo. We will
approximate the velocity distribution of the halo as a single Gaussian
distribution with velocity dispersion $\sigv$, in which case we need
to average the enhancement over a truncated Maxwell-Boltzmann
distribution
\begin{equation}\label{eq:maxboltz}
f(v) \propto \left\{
\begin{array}{rl}
v^2\, e^{-v^2/2\sigv^2}  & \qquad v \leq v_{\mathrm{esc}}\\
0 & \qquad v > v_{\mathrm{esc}}\, ,\\
\end{array} \right.
\end{equation}
in which we set the escape velocity to $v_{\mathrm{esc}} =
\sqrt{2}\,\sigv$. Note that in the pure $1/v$ regime this averaging
will tend to sustain the $1/v$ behavior up to a factor $\order(1)$,
since then
\begin{equation}
\int_0^\infty \dd v \, v^2\, e^{-v^2/2\sigv^2}
\somm\left/\int_0^\infty \dd v \, v^2\, e^{-v^2/2\sigv^2} \right.\propto
\frac{1}{\sigv}\, ,
\end{equation}
since the truncation does not significantly influence the
result. Similarly, in the resonance region the $1/v^2$ behavior is
conserved by the averaging. When the enhancement has leveled off the
averaging has no effect.

\begin{figure*}[t]
\includegraphics[clip=]{avg-3_64.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{avg-4_128.eps}\\%%BoundingBox: 190 291 418 474
\includegraphics[clip=]{avg-5_128.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{avg-6_128.eps}%%BoundingBox: 190 291 418 474
\caption{Sommerfeld enhancement factor as a function of the particle
physics parameters averaged over a Maxwell-Boltzman velocity
distribution with velocity dispersion \sigv\ for four different orders
of magnitude of the velocity dispersion (all velocities are relative
to the speed of light).}%
\label{fig:sommavg}%
\end{figure*}

\section{Substructure boosts from Sommerfeld enhancement}
The photon flux from \DM\ annihilation from a solid angle
$\Delta\Omega$ along a given line-of-sight (\los) is given by
\begin{equation}\label{eq:flux}
\begin{split}
\frac{\dd \Ngamma}{\dd A \,\dd t}  = &
\int_{\Eth}^{\mdm}\dd E\, \sum_i \frac{\dd
\Ngammai}{\dd E} \,\frac{\sigmaannv_i}{\mdm^2}\,
\somm\left(\mv/\mdm,\alpha,\sigv\right)\,\\
& \times \int_{\Delta\Omega}\,\frac{\dd\Omega}{4\pi}\,\int_{\mathrm{\los}} \dd l\,
\rho^2(l)\, .
\end{split}
\end{equation}
In this \Eth\ is the treshold energy of the detector, $i$ denotes
(possibly) different final states of the \DM\ annihilation, $\frac{\dd
\Ngammai}{\dd E}$ is the spectrum from the annihilation to that state,
$\sigmaannv_i$ is the annihilation cross section to state $i$ (without
the Sommerfeld enhancement), $\somm\left(\mv/\mdm,\alpha,\beta\right)$
is the Sommerfeld enhancement which we have factored out of the cross
section, and $\rho(l)$ is the \DM\ density along the line-of-sight. We
emphasize that this framework also allows us to consider different
final annihilation products, such as the electrons and positrons
supposedly seen by PAMELA/ATIC.

We will focus our analysis on the quantity on the ``structure quantity''
\lum($M$), which we define here as
\begin{equation}\label{eq:structquant}
\lum(M)\equiv\somm\left(\mv/\mdm,\alpha,\sigv\right)\,\int_{\Delta\Omega}\,\frac{\dd\Omega}{4\pi}\int_{\mathrm{\los}}
\dd l\, \rho^2(l)
\end{equation}
This structure quantity does not only depend on the mass and internal
properties of the \DM\ halos in which we are interest, it depends on
the particle physics parameters as well. We will keep all of the other
factors in \eqnname\ (\ref{eq:flux}) fixed. However, in order to
derive actual numbers for the photon flux from \DM\ annihilation, we
briefly describe the other quantities appearing in \eqnname\
(\ref{eq:flux}) here.

\subsection{Some details of the particle physics that do not particularly concern us here}

For the annihilation cross section without taking account of any
Sommerfeld enhancement we will adopt the fiducial value $\sigmaannv =
3 \times 10^{-26} $ cm$^3$ s$^{-1}$. The energy spectrum of the \DM\
annihilation depends on the model used to describe the \DM\ and its
interactions and it is outside of the scope of this paper to provide a
detailed treatment of this for different \DM\ models. Therefore, we
use an annihilation spectrum appropriate for neutralino dark matter,
which provides a reasonable fit to a variety of different annihilation
products. Since the energy spectrum contributes a multiplicative
factor to the \DM\ annihilation flux, the effect of different
annihilation spectra can be incorporated by simply multiplying our
results with the appropriate factor. The neutralino dark matter
annihilation spectrum we use has the form
\cite{1998APh.....9..137B,2004PhRvD..70j3529F}
\begin{equation}\label{eq:annspectrum}
\frac{\dd \Ngamma}{\dd E} = \frac{a_1}{\mdm} \left(\frac{E}{\mdm}\right)^{-3/2}\, \exp\left[-a_2\frac{E}{\mdm}\right]\, ,
\end{equation}
with parameters $a_1\sim 1$ and $a_2 \sim 10$, depending on the
details of the annihilation process
\cite{2004PhRvD..70j3529F}. Integrating this spectrum using a
threshold energy appropriate for \Fermi\, $\Eth = 4$ GeV, and choosing
a \DM\ mass, $\mdm = 700$ GeV, we find that
\begin{equation}
\int_{\Eth}^{\mdm}\dd E\, \sum_i \frac{\dd
\Ngammai}{\dd E} \approx 17\, ,
\end{equation}
with any differences due to variations in $(a_1,a_2)$ and $\mdm$
between different models leading to changes of a factor $\order(10)$.

\subsection{The Dark Matter profile}

The density profile of dark matter halos has been discussed
extensively in the last few years. While the shape of the outer
density profile seems to have converged in the various numerical
simulations of \DM\ halos and a consensus has been reached, the
detailed shape of the inner density profile is still a hot topic of
debate. The main point of contention is whether the density profile is
cusped near the center of the \DM\ halo, and if so, how steep this
cusp is, or whether the density stays finite near the center. Since
this discussion has a non-negligible impact on the prospects for \DM\
annihilation detection, we will briefly discuss the various plausible
density profiles.

The inner structure of \DM\ halos is one of the key predictions of the
\CDM\ paradigm, and over a decade ago it was shown that the
spherically averaged density profiles of \DM\ halos are well described
by a universal profile, the so-called \NFW\ profile
\cite{1995MNRAS.275..720N,1996ApJ...462..563N,1997ApJ...490..493N},
depending only on the virial mass of the halo and its
concentration. Moreover, the concentration was shown to depend
systematically on the halo mass
\cite{1996ApJ...462..563N,1997ApJ...490..493N,2001MNRAS.321..559B,2001ApJ...554..114E,2007MNRAS.381.1450N,2008MNRAS.387..536G}. This
profile was found to provide a good fit to many numerical simulations
(\eg,
\cite{1996MNRAS.281..716C,1997MNRAS.286..865T,1997ApJS..111...73K}).
While there seems to be a consensus between different numerical
simulations on the shape of the outer density profile, no agreement
was reached on the inner density structure of the \DM\ halos found in
the simulations. The \NFW\ profile has an inner slope of -1, other
studies, however, found evidence for steeper inner slopes
\cite{1997ApJ...477L...9F,1998ApJ...499L...5M,1999MNRAS.310.1147M,2000ApJ...544..616G,2001ApJ...554..903K,2001ApJ...557..533F,2004MNRAS.353..624D},
shallower inner slopes \cite{1998ApJ...502...48K}, or didn't find any
convergence toward a power-law behavior
\cite{2003MNRAS.338...14P}. Even the universality of the density
profile was brought into question
\cite{2000ApJ...529L..69J,2000ApJ...535...30J}.

More recently, high-resolution numerical simulations have essentially
ruled out the steepest cusps, such as Moore's profile, which has an
inner cusp with logarithmic slope of -1.5, however steeper inner cusps
than the original \NFW\ profile are still observed in numerical
simulations \cite{2005MNRAS.364..665D}. Other recent simulations find
inner slopes that are shallower than an \NFW\ profile
\cite{2008MNRAS.385..545K,2008arXiv0808.2981S,2008arXiv0810.1522N},
although there is a growing consensus that the logarithmic slope of
the inner density profile is a quantity that hasn't fully converged
yet in any numerical simulation. It also has become apparent in recent
years that the inner density profile of \DM\ halos might not be
described by a power-law divergence, but that, on the contrary, the
logarithmic slope becomes ever shallower, leading to a cored density
profile
\cite{2004MNRAS.349.1039N,2005ApJ...624L..85M,2006AJ....132.2685M,2006AJ....132.2701G,2006MNRAS.365..147S,2008arXiv0808.2981S,2008MNRAS.391.1685S,2008arXiv0810.1522N}. Therefore,
density profiles such as the Einasto profile
\cite{einasto65a,1989A&A...223...89E} have been shown to provide an
excellent fit to the numerical simulations with the highest resolution
yet \cite{2008MNRAS.391.1685S,2008arXiv0808.2981S}.

Since the photon flux from \DM\ annihilations depends on the density
squared of the \DM\, the exact form of the inner density profile has
important ramifications for the detectability of \DM\
annihilation. This has been widely discussed before in the context of
various \DM\ annihilation models (\eg,
\cite{1998APh.....9..137B,1999PhRvL..83.1719G,2000PhRvD..62l3005C,2000PhLB..494..181G,2002PhRvD..66b3509T,2002PhRvD..66l3502U,2003MNRAS.339..505T,2004ApJ...601...47A,2008JCAP...07..013B,2008arXiv0811.3744B,2008arXiv0812.3895B}). Roughly
speaking, the annihilation flux is smaller for inner density profiles
which are cored, and larger for inner density profiles which show a
cuspy behavior. The steeper the slope of the inner cusp, the larger
the \DM\ annihilation flux (for obvious reasons). Since the core
vs.~cusp debate is still undecided we will consider \DM\ distribution
models of both types.

The cuspy \DM\ profile we will use is a generalized \NFW\ profile (\GNFW):
\begin{equation}\label{eq:NFW}
\rho(r) = \frac{\rhos}{\redr^\gamma\left(1+\redr\right)^{3-\gamma}};\qquad \redr = r/\rs\, ,
\end{equation}
in which \rhos\ is a characteristic density and \rs\ a characteristic
(scale-)radius.  The original \NFW\ profile is recovered for $\gamma$
= 1 \cite{1997ApJ...490..493N}, which we will use as the fiducial
cuspy model in what follows. Varying the value of the inner
logarithmic slope $\gamma$ will allow us to describe the effect of
steeper and shallower inner slopes than the \NFW\ model. Note that
this functional form does \emph{not} include the exact form of Moore's
profile; this need not concern us since inner cusps as steep as
Moore's profile have been as good as excluded by the latest numerical
simulations.

The contribution of the \DM\ density profile on the structure quantity
\lum\ can be simply evaluated in the context of any density model. As
in Ref.~\cite{2007PhRvD..75h3526S} we assume that the distance
\dist\ is much larger than the scale-radius of the \DM\ halo. For an
\NFW\ profile 90\% of the flux originates from the region with one
scale-radius, and we will, conservatively, only take contributions to
the \DM\ annihilation flux coming rom this region for all the
\GNFW\ profiles. Therefore, we find that 
\begin{equation}\label{eq:lumrhosrsNFW}
\int_{\Delta\Omega}\,\frac{\dd\Omega}{4\pi}\,\int_{\mathrm{\los}} \dd
l\, \rho^2(l)= \normnfw(\gamma)\, \frac{\rhos^2\, \rs^3}{\dist^2}\, .
\end{equation}
The normalization of this relationship depends on the logarithmic
slope of the inner cusp and is given by
\begin{equation}\label{eq:normgammaNFW}
\normnfw(x) = 4 \pi  \, \frac{2^{2x-5}\left(-16+11x-2x^2\right)}{-30+47x-24x^2+4x^3}\, .
\end{equation}
The normalization for the \NFW\ profile is $7\pi/6$ and it depends
strongly on the value of the logarithmic slope of the inner density
profile, as can be seen in Figure \ref{fig:normgammaNFW}. The lower
panel of Fig.~\ref{fig:normgammaNFW} shows the fraction of the \DM\
annihilation flux coming from region within one scale-radius of the
halo. For reasonable values of the inner slope this fraction is in the
range 80 -- 95\%.

\begin{figure}[t]
\centering
\includegraphics{normgammaNFW.eps}
\caption{Top: Normalization of the $\lum \propto \rhos^2\,
\rs^3/\dist^2$ relation of \eqnname\ (\ref{eq:lumrhosrsNFW}) for the
\GNFW\ profile. The functional dependence of the normalization on the
exponent $\gamma$ in the profile given in \eqnname\ (\ref{eq:NFW}) is
given in \eqnname\ (\ref{eq:normgammaNFW}). Bottom: Fraction of the
emission that comes from the region within \rs.}%\
\label{fig:normgammaNFW}%
\end{figure}

The Einasto profile is characterized by an ever-decreasing logarithmic
inner slope. It's functional dependence on the radius is given by
\begin{equation}\label{eq:einasto}
\rho (r) = \rhominustwo\, \exp\left[-\frac{2}{\alpha}\left\{\left(\frac{r}{\rminustwo}\right)^{\alpha}-1\right\}\right]\, .
\end{equation}
Here, \rhominustwo\ is the density at the radius \rminustwo\ where the
local slope is -2 and \alphaEinasto\ is a parameter often fixed at a
value $\sim\!0.17$ \cite{2008MNRAS.391.1685S}. As above, assuming
$\dist \gg \rminustwo$, calculating the flux originating from within
the radius \rminustwo, we find that
\begin{equation}\label{eq:lumrhosrsEinasto}
\int_{\Delta\Omega}\,\frac{\dd\Omega}{4\pi}\,\int_{\mathrm{\los}} \dd
l\, \rho^2(l)= \normeinasto(\alpha)\, \frac{\rhominustwo^2\, \rminustwo^3}{\dist^2}\, .
\end{equation}
The normalization depends on the value of \alphaEinasto\ and is given
by
\begin{equation}\label{eq:normgammaEinasto}
\normeinasto(x) = 4 \pi \, 2^{-6/\alpha}\,e^{4/\alpha}\,
\alpha^{-1+3/\alpha}\,
\gamma\left[\frac{3}{\alpha},\frac{4}{\alpha}\right]\,
,
\end{equation}
where $\gamma[a,z]$ is the lower incomplete gamma function. The
dependence of this normalization on \alphaEinasto\ is shown in Figure
\ref{fig:normgammaEinasto}. In order to compare the Einasto profile as
described here, we can relate the parameters $(\rhos,\rs)$ of the
profile in \eqnname\ (\ref{eq:NFW}) to the parameters
$(\rhominustwo,\rminustwo)$ of the Einasto model by calculating the
radius at which the logarithmic slope of the density versus radius
profile equals -2. Thus, we find that
\begin{equation}
\rminustwo = \rs,  \qquad \qquad \rhominustwo = \frac{\rhos}{2^{3-\gamma}}\, .
\end{equation}

Note that the \DM\ annihilation flux from the Einasto profile,
although it seems larger from Figs.~\ref{fig:normgammaNFW} and
\ref{fig:normgammaEinasto}, is actually smaller in general, since
there is a factor of 2$^{6-2\gamma}$ involved in going from one to the
other.

\begin{figure}[t]
\centering
\includegraphics{normgammaEinasto.eps}
\caption{Top: Normalization of the $\lum \propto \rhominustwo^2\,
\rminustwo^3/\dist^2$ relation of \eqnname\
(\ref{eq:lumrhosrsEinasto}) for the Einasto profile. The functional
dependence of the normalization on $\alpha$ in the profile given in
\eqnname\ (\ref{eq:einasto}) is given in \eqnname\
(\ref{eq:normgammaEinasto}). Bottom: Fraction of the emission that
comes from the region within \rminustwo.}%\
\label{fig:normgammaEinasto}%
\end{figure}

Further on it will sometimes be useful to describe the \DM\ density
profiles by different parameters than a characteristic density and
radius. Another set of parameters is given by a concentration
parameter and a mass. The concentration for both the \GNFW\ profile as
for the Einasto profile is given by
\begin{equation}\label{eq:concdef}
\conc \equiv \frac{\Rvir}{\rminustwo}\, .
\end{equation}
A characteristic mass is given by the virial mass \Mvir. The virial
mass of the \GNFW\ profile is given by
\begin{equation}\label{eq:Mvirdef}
\Mvir = 4\pi\rhos\rs^3 \left[ - (-\conc)^\gamma\,\conc^{-\gamma} B(-c;3-\gamma,\gamma-2)\right]\, ,
\end{equation}
in which $B(z;a,b)$ is the incomplete Beta function, given by the
familiar form $f(c) = \ln(1+c)-c/(1+c)$ in the special case of the
\NFW\ profile. The virial mass for the Einasto profile is given by
\begin{equation}
\Mvir = 4\pi\rhominustwo\rminustwo^3 \frac{1}{\alphaEinasto} \exp\left(\frac{3\ln \alphaEinasto + 2 - \ln 8}{\alphaEinasto}\right) \gamma\left[\frac{3}{\alphaEinasto},\frac{2}{\alphaEinasto}\conc^\alphaEinasto\right]\, .
\end{equation}
In this last expression $\gamma\left[a,z\right]$ is the lower
incomplete gamma function. The density can then be expressed in terms
of the concentration by using the definition of the virial mass:
$\Mvir \equiv 4\pi/3 \Delta \rhoc \Rvir^3$, where $\Delta = 178\,
\omegam^{0.45} \approx 100$
\cite{1996MNRAS.282..263E,1998ApJ...495...80B}, and \rhoc\ is the
critical density.

Another pair of parameters that one can use to characterize the \DM\
density profile is the maximum circular velocity \vmax\ and the radius
at which this maximum occurs, \rmax. These parameters are especially
well-suited when dealing with numerical simulations, since they can be
readily read off from the raw simulation data. The circular velocity
is given by
\begin{equation}
\vcirc^2(r) \equiv \frac{G M(r)}{r} = \vcirc^2(\Rvir)\frac{c}{f(c)}\frac{f(x)}{x}\, ,
\end{equation}
in which $x \equiv r/\rminustwo$ and $f(x)$ is given by
$B(-x;3-\gamma,\gamma-2)$ in the case of a \GNFW\ profile and by
$\gamma\left[\frac{3}{\alphaEinasto},\frac{2}{\alphaEinasto}x^\alphaEinasto\right]$
in the case of an Einasto profile. One can find \rmax\ by (numerically) solving the equation
\begin{equation}
\frac{\dd}{\dd x}\left[ \frac{f(x)}{x}\right] = 0\, ,
\end{equation}
and \vmax\ follows then immediatly.

\subsection{Calculating the total boost}

As is customary, we will write the total structure quantity \lum($M$) as a
sum of a smooth component and a substructure component:
\begin{equation}
\lum(M) = \lumsmooth(M) + \lumsub(M)\, .
\end{equation}
The smooth contribution to the structure quantity is the product of
the Sommerfeld enhancement \somm\ and the line-of-sight integral over
the \DM\ density squared, given for the different \DM\ density
profiles in \eqnname s (\ref{eq:lumrhosrsNFW}) and
(\ref{eq:lumrhosrsEinasto}). Therefore,
\begin{equation}
\lumsmooth(M) = \somm \, \lumsmoothc(M)\, ,
\end{equation}
in which \lumsmoothc\ is the luminosity that would come from the
smooth component of the halo in the absence of Sommerfeld
enhancememnt.

We can now write
\begin{equation}
\begin{split}
\lum(M) & = \somm\,\lumsmoothc(M) + \lumsub(M)\\
&= \left[ \somm + \frac{\lumsub(M)}{\lumsmoothc(M)}\right] \lumsmoothc(M)\, .
\end{split}
\end{equation}
We will from now on write the substructure boost as \boost(M). Then we
have the following
\begin{equation}\label{eq:boost}
\begin{split}
\boost(M) & \equiv \frac{\lumsub(M)}{\lumsmoothc(M)} = \frac{1}{\lumsmoothc(M)} \int \dd m \, \frac{\dd N}{\dd m} \lum(m)\\
&= \frac{1}{\lumsmoothc} \int \dd m\, \frac{\dd N}{\dd m} \left[ \somm(m) + \boost(m) \right] \lumsmoothc(m)\, .
\end{split}
\end{equation}
Here we have introduced the subhalo mass function $\dd N/\dd m$, which
gives the abundance of subhalos of mass $m$.

Before proceeding we should specify the upper and lower limits of the
integral appearing in \eqnname\ (\ref{eq:boost}). The integration over
ever smaller substructures is cut-off at the size of the smallest
subhalos. The mass scale at which this cut-off occurs is the thermal
free-streaming scale, a \DM\ model-dependent quantity, which in many
\DM\ models lies in the range 10$^{-6}$ -- 10$^{-12}$ \Msol
\cite{1999PhRvD..59d3517S,2001PhRvD..64h3507H,2004MNRAS.353L..23G,2005PhRvD..71j3520L,2006PhRvL..97c1301P}. The
upper limit to this integral is some fraction $q$ of the halo mass
$M$, since the abundance of substructure does not continue up to this
mass scale.

In order to calculate the total boost from substructure to the
annihilation cross section, we will write all of the quantities
appearing in \eqnname\ (\ref{eq:boost}) as a function of the mass
$m$. For instance, from numerical simulations, the subhalo mass
function is well constrained to follow a power-law type behavior:
\begin{equation}\label{eq:massfraction}
\frac{\dd N}{\dd \ln m}
\propto \left(\frac{m}{M}\right)^n\, .
\end{equation}
The value of the exponent $n$ is still under debate. The two highest
resolution $N$-body simulations yet, the \aquarius\ and \VL\
projects, both find values for $n$ that put roughly equal amounts of
mass per subhalo mass decade. While the \VL\ simulation seems to
prefer a value $n = -1$
\cite{2006ApJ...649....1D,2008ApJ...686..262K}, which would lead to a
logarithmically diverging total mass in substructures if the subhalo
abundance were not cut off at the thermal free-streaming limit of the
dark matter, the \aquarius\ project prefers a value $n = -0.9$
\cite{2008MNRAS.391.1685S}. As the \aquarius\ project has made
detailed convergence studies that support their claim for a value of
$n$ that is shallower than -1, we will adopt the value $n = -0.9$ as
our fiducial value, although even shallower slopes have also been
observed \cite{2002PhRvD..66f3502H}. The same slope has also been
observed in other simulations \cite{2004MNRAS.355..819G}.

With the exponent $n$ fixed, we turn to the normalization of the
relation given in \eqnname\ (\ref{eq:massfraction}). As one might
expect by now, different high-resolution numerical simulations give
different answers to this question as it relates to the abundance of
subhalos of Galactic subhalos. The problem of the normalization is
complicated by the fact that numerical simulations have only recently
begun to be able to resolve substructure in subhalos in a Galaxy-size
halo. While the simulations generally agree that the resolved
substructure abundance of subhalos in the host halo adds up to
$\sim\!10$\% of the total mass of the halo
\cite{1998MNRAS.300..146G,1999ApJ...522...82K,2004MNRAS.353..624D,2003ApJ...598...49Z,2007ApJ...657..262D,2007ApJ...667..859D},
the question of whether the substructure abundance in subhalos is just
a scaled-down version of the substructure abundance in the host halo
remains unresolved. One would not expect the abundance to be
self-similar, since, while substructure in both the main host halo and
subhalos gets diminished by tidal disruptions, substructure in the
subhalos does not get replenished by the infall of halos from the
field, as is the case for the host halo. Nevertheless, many numerical
simulations do find approximately self-similar substructure abundances
\cite{1999ApJ...524L..19M,2007ApJ...659.1082S}. Recently, the
\aquarius\ simulation has concluded from its simulation that the
substructure abundance is not self-similar, although the deviation
from the self-similar relation is small
\cite{2008MNRAS.391.1685S}. Therefore, we will adopt a normalization
here which places 10\% of the mass of the subhalo in the form of
``resolved'' substructure, where we adopt the definition that resolved
substructure is in halos of mass $\gtrsim 10^{-5} M$.

In the previous section we calculated $\lumsmoothc$ for the different
\DM\ density profiles and found in all cases that
\begin{equation}
\lumsmoothc(m) \equiv \lumsmoothc(\rho(m),r(m))\, ,
\end{equation}
for a characteristic density $\rho$ and radius $r$, the exact
definition of which depends on the specific \DM\ density profile
used. In order to find the mass dependence of $\lumsmoothc$, we use
the description of the \DM\ density profile in terms of virial mass
\Mvir\ and concentration \conc. We then make use of the fact that many
simulations have found a relation between the concentration of halos
and their mass
\cite{2001MNRAS.321..559B,2001ApJ...554..114E,2008MNRAS.387..536G,2007MNRAS.381.1450N},
such that everything depends only on the mass. This relation is well
established over a large range in halo masses, and we will adopt the
particular relationship found in Ref.~\cite{2007MNRAS.378...55M}
\begin{equation}\label{eq:concscaling}
\conc = 10.5 \, \left(\frac{\Mvir}{10^{12}\, \Msol}\right)^{-0.11}\, ,
\end{equation}
in good agreement with the results of other, similar analyses
\cite{2001MNRAS.321..559B,2005MNRAS.357..387K}, although the
normalization of the concentration-mass relation is about 15\% lower
than the original relation. However, the concentration of subhalos is
generally larger than the concentration of field halos of the same
mass as a result of tidal mass-loss
\cite{1998MNRAS.300..146G,2008ApJ...673..226P,2004MNRAS.355..794H,2005ApJ...635..931B,2004ApJ...608..663K}. This
is a small effect, of order 10\%, and we will investigate later
whether it significantly influences our results.

The only remaining quantity in \eqnname\ (\ref{eq:boost}) that we
haven't related to the mass of the subhalo yet, is the Sommerfeld
enhancement factor \somm. The Sommerfeld enhancement depends on the
mass of the subhalo through the velocity dispersion. Therefore, we
need to relate the velocity dispersion of a \DM\ subhalo to its
velocity dispersion. It is important to note that all we really need
is an approximate relationship between the velocity dispersion and the
halo mass which fits well the range of Galactic size halos to a scale
not much smaller than the mass scale of the smallest dwarf Spheroidals
known today, since in most cases the Sommerfeld enhancement will
saturate at velocity dispersions of the order of the velocity
dispersion of the dwarf Spheroidals.

On the scale of clusters there is a well-defined relationship between
the mass of a \DM\ halo and its one-dimensional velocity dispersion
which one can derive from adiabatic scaling arguments
\cite{1998ApJ...495...80B} given by
\begin{equation}\label{eq:vdispscaling}
\sigma \propto m^{1/3}\, .
\end{equation}
This scaling relation can be easily derived in the case of an
isothermal distribution function \cite{2008gady.book.....B}, since
then
\begin{equation}
\rho(r) = \frac{\sigma^2}{2\pi G r^2}\, ,
\end{equation}
from which \eqnname\ (\ref{eq:vdispscaling}) immediatly follows. This
scaling relation has been found to describe numerical simulations of
cluster-sized halos extremely well
\cite{1991ApJ...383...95E,1995MNRAS.275..720N,1995AJ....110...21C,1996ApJ...469..494E,1996MNRAS.281..716C,1998ApJ...495...80B}.From
recent numerical simulations it seems that we can use this relation
even for subhalos of subhalos in a Galaxy sized dark matter halo. In
Ref.~\cite{2008MNRAS.391.1685S} it is shown that the relation between
the maximum circular velocity of a subhalo and its mass is well fit by
a power-law of the type
\begin{equation}
m \propto \vmax^{3.5}\, .
\end{equation}
By a simple estimate of the velocity dispersion using the virial
theorem, we can use the maximum circular velocity as a measure of the
halos velocity dispersion. This would give $\sigv \propto
m^{1/3.5}$. Alternatively, we could use the circular velocity at the
virial radius as a measure of the velocity dispersion. In the case of
an \NFW\ profile we have that
\begin{equation}
\frac{\vmax^2}{\vcirc(\Rvir)^2} \propto \frac{c}{f(c)}\, ,
\end{equation}
in which $f(x) = \ln(1+c) - c/(1+c) \approx 2.6 (c/33)^{0.4}$ in the
relevant concentration range. Using \eqnname\ (\ref{eq:concscaling}),
this gives
\begin{equation}
\sigv \propto M^{0.32}\, .
\end{equation}
Therefore, we will use (as a rough approximation) the scaling relation
given in \eqnname\ (\ref{eq:vdispscaling}).

\section{A typical distribution of signals in a halo}

\section{Segue 1}

\section{Microhalos}

{\bf Acknowledgments:}
NASA 07-ADP07-0099.

\bibliographystyle{apsrev}
\bibliography{ann}

\end{document}

### 2009-01-30 py/1denhance.py

@@ -1,0 +1,44 @@
+#! /usr/bin/env python
+from enhance import enhance
+from math import exp, cos, sin, sqrt
+from scipy import integrate
+from numpy import *
+from matplotlib.pyplot import *
+import os
+import pickle
+
+
+nb=5
+nsamples=20
+
+savefilename='1d.sav'
+if os.path.exists(savefilename):
+    savefile=open(savefilename,'r+')
+    data=pickle.load(savefile)
+else:
+    mv=.09
+    a=1/30.0
+    
+    m= linspace(0,2,nsamples)
+    m= 10**m
+    b= array([.1,.01,.001,.0001,.00001])
+    S= zeros((nsamples,nb))
+
+    for ii in range(nsamples):
+        for jj in range(nb):
+            S[ii,jj]= enhance(mv,m[ii],a,b[jj])
+            print str(ii*nb+jj)+'/'+str(nb*nsamples)
+
+    savefile=open(savefilename,'w')
+    data={'m':m,'S':S}
+    pickle.dump(data,savefile)
+
+
+            
+for jj in range(nb):
+    loglog(data['m'],data['S'][:,jj])
+
+axis([1,100,1,10**6])
+xlabel('m (TeV)')
+ylabel('S')
+savefig("enhancement.eps",format='eps')

### 2009-01-30 py/2dboost.py

@@ -1,0 +1,121 @@
+#Make density plot of total boost for a halo of mass Mh
+#function of m_V/m and alpha
+#argument1: log_10 Mh
+#argument2: grid (default: 64)
+
+from math import *
+from numpy import *
+import matplotlib
+matplotlib.use('PS')
+from pylab import *
+from matplotlib.pyplot import *
+from enhance import avg_enhance
+from boost import boost
+import os, sys
+import pickle
+import matplotlib.cm as cm
+from matplotlib import rc
+
+
+#parameters
+if len(sys.argv) > 1:
+    logMh= int(sys.argv[1])
+else:
+    print "No Mh given, returning..."
+    sys.exit(-1)
+Mh=10**logMh
+vdisp= 2.5*10**-2*pow(Mh,1./3.)/(3*10**5)
+m=1.0
+if len(sys.argv) > 2:
+    nms= int(sys.argv[2])
+    nas= int(sys.argv[2])
+else:
+    nms= 64
+    nas= 64
+mv= linspace(-5,-2,nms)
+mv= 10**mv
+a= linspace(-3,-1,nas)
+a= 10**a
+    
+#construct filenames
+savefilename= '2dboost'+str(logMh)+'_'+str(nas)+'.sav'
+plotfilename= '2dboost'+str(logMh)+'_'+str(nas)+'.eps'
+
+
+if os.path.exists(savefilename):
+    savefile=open(savefilename,'r+')
+    data=pickle.load(savefile)
+    ii= data['ii']
+    jj= data['jj']
+    if data['nms'] != nms:
+        print "Warning: suppplied nms does not equal savefile nms"
+        sys.exit(-1)
+    if data['nas'] != nas:
+        print "Warning: suppplied nas does not equal savefile nas"
+        sys.exit(-1)
+    mv= data['mv']
+    a= data['a']
+    S= data['S']
+    print 'restarting at ii= '+str(ii)+', jj= '+str(jj)+', at iteration '+str(ii*nms+jj+1)+'/'+str(nms*nas)
+else:
+    S= zeros((nas,nms))
+    ii= 0
+    jj= 0
+    savefile=open(savefilename,'w')
+
+while ii < nms:
+    while jj < nas:
+        boo= boost(Mh,mv[ii],a[jj])
+        so= avg_enhance(mv[ii],m,a[jj],vdisp)
+#        print mv[ii]
[…]

### 2009-01-30 py/2denhance.py

@@ -1,0 +1,95 @@
+#Make density plot of Sommerfeld enhancement as a
+#function of m_V/m and alpha, for fixed beta
+#argument1: log_10 beta
+#argument2: grid (default: 64)
+
+
+from math import *
+from numpy import *
+import matplotlib
+matplotlib.use('PS')
+from pylab import *
+from matplotlib.pyplot import *
+from enhance import enhance
+import os, sys
+import pickle
+import matplotlib.cm as cm
+from matplotlib import rc
+
+
+#construct filenames
+if len(sys.argv) > 1:
+    logb= int(sys.argv[1])
+else:
+    print "No beta given, returning..."
+    sys.exit(-1)
+
+savefilename= '2d'+str(logb)+'.sav'
+plotfilename= '2d'+str(logb)+'.eps'
+
+if os.path.exists(savefilename):
+    savefile=open(savefilename,'r+')
+    data=pickle.load(savefile)
+else:
+    b=10**-3
+    m=1.0
+    if len(sys.argv) > 2:
+        nms= int(sys.argv[2])
+        nas= int(sys.argv[2])
+    else:
+        nms= 64
+        nas= 64
+    mv= linspace(-5,-2,nms)
+    mv= 10**mv
+    a= linspace(-3,-1,nas)
+    a= 10**a
+    
+    S= zeros((nas,nms))
+    
+    for ii in range(nms):
+        for jj in range(nas):
+            S[ii,jj]= log10(enhance(mv[ii],m,a[jj],b))
+            sys.stdout.write('\r'+str(ii*nms+jj+1)+'/'+str(nms*nas))
+            sys.stdout.flush()
+    sys.stdout.write('\n')
+
+    savefile=open(savefilename,'w')
+    data={'mv':mv,'a':a,'S':S}
+    pickle.dump(data,savefile)
+
+
+
+
+#Plotting parameters
+fig_width = 3.25  # width in inches
+fig_height = 2.9      # height in inches
+fig_size =  [fig_width,fig_height]
+params = {'backend': 'ps',
+          'axes.labelsize': 12,
+          'text.fontsize': 12,
+          'legend.fontsize': 12,
+          'xtick.labelsize':10,
+          'ytick.labelsize':10,
+          'text.usetex': True,
+          'figure.figsize': fig_size}
[…]

### 2009-01-30 py/avgenhance.py

@@ -1,0 +1,112 @@
+#Make density plot of averaged Sommerfeld enhancement as a
+#function of m_V/m and alpha
+#argument1: log_10 beta
+#argument2: grid (default: 64)
+
+from math import *
+from numpy import *
+import matplotlib
+matplotlib.use('PS')
+from pylab import *
+from matplotlib.pyplot import *
+from enhance import avg_enhance
+import os, sys
+import pickle
+import matplotlib.cm as cm
+from matplotlib import rc
+
+
+#parameters
+if len(sys.argv) > 1:
+    logb= int(sys.argv[1])
+else:
+    print "No beta given, returning..."
+    sys.exit(-1)
+b=10**logb
+m=1.0
+if len(sys.argv) > 2:
+    nms= int(sys.argv[2])
+    nas= int(sys.argv[2])
+else:
+    nms= 64
+    nas= 64
+mv= linspace(-5,-2,nms)
+mv= 10**mv
+a= linspace(-3,-1,nas)
+a= 10**a
+    
+#construct filenames
+savefilename= 'avg'+str(logb)+'_'+str(nas)+'.sav'
+plotfilename= 'avg'+str(logb)+'_'+str(nas)+'.eps'
+
+
+if os.path.exists(savefilename):
+    savefile=open(savefilename,'r+')
+    data=pickle.load(savefile)
+    ii= data['ii']
+    jj= data['jj']
+    if data['nms'] != nms:
+        print "Warning: suppplied nms does not equal savefile nms"
+        sys.exit(-1)
+    if data['nas'] != nas:
+        print "Warning: suppplied nas does not equal savefile nas"
+        sys.exit(-1)
+    mv= data['mv']
+    a= data['a']
+    S= data['S']
+    print 'restarting at ii= '+str(ii)+', jj= '+str(jj)+', at iteration '+str(ii*nms+jj+1)+'/'+str(nms*nas)
+else:
+    S= zeros((nas,nms))
+    ii= 0
+    jj= 0
+    savefile=open(savefilename,'w')
+
+while ii < nms:
+    while jj < nas:
+        S[ii,jj]= log10(avg_enhance(mv[ii],m,a[jj],b))
+        sys.stdout.write('\r'+str(ii*nms+jj+1)+'/'+str(nms*nas))
+        sys.stdout.flush()
+        jj= jj+1
[…]

### 2009-01-30 py/boost.py

@@ -1,0 +1,114 @@
+from scipy import special
+from numpy import *
+from math import exp, log
+from enhance import avg_enhance, enhance
+from scipy import integrate
+
+def deriv(y,t,mv,a,gamma,einasto,n,q,A,C,K1,K2,vdisp_sat,somm_sat):
+    """Implements the right hand side of the boost differential equation
+
+    Input:
+       y    - current dependent variable  value
+       t    - current independent variable
+       mv      - force carrier mass (/mchi)
+       a       - new force fine structure constant
+       gamma   - GNFW gamma or Einasto gamma
+       einasto - if 0 GNFW, 1 Einasto
+       n       - mass function exponent
+       q       - mass fraction until which to integrate
+       A       - normalization of mass function
+       C       - normalization of sigma(m) relation
+       K1      - normalization of concentration-mass relation
+       K2      - exponent of concentratio-mass relation
+       vdisp_sat - velocity dispersion at which the Sommerfeld enhancement saturates
+       somm_sat  - saturated value of the Sommerfeld enhancement
+
+    Output:
+       derivative
+    """
+    #Diagnostics, print M and B(M)
+    #print t
+    #print y
+
+    #Calculate each factor/term in turn
+    #First calculate the concentration
+    conc= K1*pow(t,K2)/pow(10.,K2*12)
+    concq= K1*pow(q*t,K2)/pow(10,K2*12)
+    
+    #First up, L(qM)/L(M)
+    LL= pow(q,3.*K2+1.)
+    if einasto == 1:
+        fc= special.gammainc(3./gamma,2./gamma*pow(conc,gamma))
+        LL*= pow(fc,2)#we don't need to add gamma(3/gamma) since this would get divided out
+        LL/= pow(special.gammainc(3./gamma,2./gamma*pow(concq,gamma)),2)#we don't need to add gamma(3/gamma) since this would get divided out
+        fc*= special.gamma(3./gamma)
+        dlnLdlnM= 1.+3.*K2+2.*conc/fc*K2*(pow(2,3./gamma)*pow(gamma,1.-3./gamma)*pow(conc,2.)*exp(-2./gamma*pow(conc,gamma)))
+    else:
[…]

### 2009-01-30 py/enhance.py

@@ -1,0 +1,82 @@
+from math import exp, cos, sin, sqrt
+from scipy import integrate
+from numpy import *
+from pylab import plot, show
+from matplotlib.pyplot import *
+
+def deriv(y,t,p1,p2):
+    """Implements the right hand side of the Sommerfeld enhancement set of differential equations (dy/dt)
+
+    Input:
+       y    - current dependent variable  value
+       t    - current independent variable
+       p1   - first parameter = b**2/a**2
+       p2   - second parameter = mv/(a*m)
+
+    Output:
+       derivative
+    """
+    return array([y[1],
+                    (-p1-1.0/t*exp(-p2*t))*y[0],
+                    y[3],
+                    (-p1-1.0/t*exp(-p2*t))*y[2]])
+
+def enhance(mv,m,a,b):
+    """Calculates the Sommerfeld enhancement for a given set of parameters
+
+    Input:
+       mv    - mass of the force carrier
+       m     - DM mass
+       a     - new force fine structure constant
+       b     - v/c
+
+    Output:
+       Sommerfeld enhancement
+    """
+    #Integration parameters
+    par1= b**2/a**2
+    par2= mv/(a*m)
+
+    #Determine start
+    if (10/par1 < 1/par2):#Coulomb like
+        start= 10**2/par1
+    else:
+        start= 10**2/par2
+
+    end=10**-4
+    y0=array([cos(b/a*start),-b/a*sin(b/a*start),
+              sin(b/a*start),b/a*cos(b/a*start)])
+    time=[start,end]
+    y=integrate.odeint(deriv,y0,time,args=(par1,par2),mxstep=10**8,mxhnil=0)
+
+    return 1/(y[1,0]**2+y[1,2]**2)
+
+def avg_enhance(mv,m,a,b,nsamples=100,vesc=0):
+    """Calculates the Sommerfeld enhancement averaged over a truncated Maxwell-Boltzman distribution
+
+    Input:
+       mv       - mass of the force carrier
+       m        - DM mass
+       a        - new force fine structure constant
+       b        - v/c
+       nsamples - number of equidistant samples to use in the averaging; default 100
+       vesc     - escape velocity (truncation velocity); default sqrt(2)*b
+
+    Output:
+       Averaged Sommerfeld enhancement
+    """
+    avg= 0.
+    norm= 0.
+
+    if vesc == 0.:
[…]

### 2009-01-30 py/plot_norm.py

@@ -1,0 +1,36 @@
+#Plot the functional dependence of the lum normalization
+from math import *
+from numpy import *
+import matplotlib
+matplotlib.use('PS')
+from pylab import *
+from matplotlib.pyplot import *
+from matplotlib import rc
+
+nsamples= 100
+
+gammas= linspace(0,1.5,nsamples,False)
+Ns= zeros(nsamples)
+for ii in range(nsamples):
+    x= gammas[ii]
+    Ns[ii]= 4*pi*pow(2,2*x-5)*(-16+11*x-2*pow(x,2))/(-30+47*x-24*pow(x,2)+4*pow(x,3))
+
+#Plotting parameters
+fig_width = 3.25  # width in inches
+fig_height = 2.9      # height in inches
+fig_size =  [fig_width,fig_height]
+params = {'backend': 'ps',
+          'axes.labelsize': 12,
+          'text.fontsize': 12,
+          'legend.fontsize': 12,
+          'xtick.labelsize':10,
+          'ytick.labelsize':10,
+          'text.usetex': True,
+          'figure.figsize': fig_size}
+rcParams.update(params)
+plot(gammas,Ns,color='black')
+axis([0,1.5,0,25])
+xlabel(r'$\gamma$')#,fontsize=16)
+ylabel(r'$\mathcal{N}(\gamma)$')#,fontsize=16)
+
+savefig("normgamma.eps",format='eps')

### 2009-01-30 py/plot_normEinasto.py

@@ -1,0 +1,119 @@
+#Plot the functional dependence of the lum normalization, Einasto profile
+from math import *
+from numpy import *
+import matplotlib
+matplotlib.use('PS')
+from pylab import *
+from matplotlib.pyplot import *
+from matplotlib import rc
+from scipy.special import *
+
+nsamples= 100
+
+gammas= linspace(0.1,0.2,nsamples)
+Ns= zeros(nsamples)
+Fs= zeros(nsamples)
+
+for ii in range(nsamples):
+    x= gammas[ii]
+    Ns[ii]= 4*pi*pow(64,-1./x)*pow(x,-1.+3./x)*exp(4./x)*gammainc(3./x,4./x)*gamma(3./x)
+    Fs[ii]=4*pi*pow(64,-1./x)*pow(x,-1.+3./x)*exp(4./x)*gamma(3./x)
+    
+
+#gammas= array( [0.1, 0.101, 0.102, 0.103, 0.104, 0.105, 0.106, 0.107,
+#0.108, 0.109,  0.11, 0.111, 0.112, 0.113, 0.114, 0.115, 0.116, 0.117,
+#0.118, 0.119,  0.12, 0.121, 0.122, 0.123, 0.124, 0.125, 0.126, 0.127,
+#0.128, 0.129,  0.13, 0.131, 0.132, 0.133, 0.134, 0.135, 0.136, 0.137,
+#0.138, 0.139,  0.14, 0.141, 0.142, 0.143, 0.144, 0.145, 0.146, 0.147,
+#0.148, 0.149,  0.15, 0.151, 0.152, 0.153, 0.154, 0.155, 0.156, 0.157,
+#0.158, 0.159,  0.16, 0.161, 0.162, 0.163, 0.164, 0.165, 0.166, 0.167,
+#0.168, 0.169,  0.17, 0.171, 0.172, 0.173, 0.174, 0.175, 0.176, 0.177,
+#0.178, 0.179,  0.18, 0.181, 0.182, 0.183, 0.184, 0.185, 0.186, 0.187,
+#0.188, 0.189,  0.19, 0.191, 0.192, 0.193, 0.194, 0.195, 0.196, 0.197,
+#0.198, 0.199,  0.2] )
+
+#Ns= array( [ 217.038, 212.896, 208.898, 205.038, 201.31, 197.707,
+#194.224,  190.854, 187.593, 184.437, 181.379, 178.417, 175.546,
[…]

### 2009-01-30 py/plot_normNFW.py

@@ -1,0 +1,73 @@
+#Plot the functional dependence of the lum normalization
+from math import *
+from numpy import *
+import matplotlib
+matplotlib.use('PS')
+from pylab import *
+from matplotlib.pyplot import *
+from matplotlib import rc
+
+nsamples= 100
+
+gammas= linspace(0,1.5,nsamples,False)
+Ns= zeros(nsamples)
+Fs= zeros(nsamples)
+
+for ii in range(nsamples):
+    x= gammas[ii]
+    Ns[ii]= 4*pi*pow(2,2*x-5)*(-16+11*x-2*pow(x,2))/(-30+47*x-24*pow(x,2)+4*pow(x,3))
+    Fs[ii]=Ns[ii]/(4*pi)*(30-47*x+24*pow(x,2)-4*pow(x,3))
+    
+#Plotting parameters
+fig_width = 3.25  # width in inches
+fig_height = 2.9      # height in inches
+fig_size =  [fig_width,fig_height]
+params = {'backend': 'ps',
+          'axes.labelsize': 12,
+          'text.fontsize': 12,
+          'legend.fontsize': 12,
+          'xtick.labelsize':10,
+          'ytick.labelsize':10,
+          'text.usetex': True,
+          'figure.figsize': fig_size}
+rcParams.update(params)
+
+
+figBG   = 'w'        # the figure background color
+axesBG  = 'w'  # the axies background color
+
+def get_locator():
+        """
+            the axes cannot share the same locator, so this is a helper
+                function to generate locators that have identical functionality
+                    """
+
+        return IndexLocator(10, 1)
+
+left, width = 0.1, 0.8
+rect1 = [left, 0.3, width, 0.6]
+rect2 = [left, 0.1, width, 0.2]
+axUpper      = axes(rect1, axisbg=axesBG)  #left, bottom, width, height
+axLower     = axes(rect2, axisbg=axesBG)#, sharex=axUpper)
+
+axUpper.xaxis.set_major_locator( get_locator() )
+
+axUpper.plot(gammas,Ns,color='black')
+axUpper.set_ylabel(r'$\mathcal{N}_1(\gamma)$')#,fontsize=16)
+axUpper.set_xticks( (0,0.5,1,1.5) )
+axUpper.set_xticklabels( () )
+
+axLower.plot(gammas,Fs,color='black')
[…]

### 2009-01-30 py/rk.py

@@ -1,0 +1,52 @@
+#Runge-Kutta routines
+
+def rk4_step(y,t,step,deriv):
+    """rk4_step: one Runge_Kutta step
+
+    Input:
+       y   - initial value
+       t   - dependent variable value
+       step- stepsize
+       deriv(y,t)- callable function which evaluates to the derivatives of y at t
+
+    Output:
+       Updated value y
+    """
+    k1= step*deriv(y,t)
+    k2= step*deriv(y+0.5*k1,t+0.5*step)
+    k3= step*deriv(y+0.5*k2,t+0.5*step)
+    k4= step*deriv(y+k3,t+step)
+
+    #print k1, k2, k3, k4
+
+    return y+1.0/6.0*(k1+2*k2+2*k3+k4)
+
+def rk4(y,start,end,nsteps,deriv):
+    """rk4: Runge-Kutta ODE solver
+
+    Input:
+       y     - Initial value
+       start - dependent variable starting point
+       end   - endpoint
+       nsteps- number of steps to take
+       deriv(y,t) - callable function which evaluates the derivative of y at t
+       
+    Output:
+       Updated value of y at end, i.e., y(end)
+    """
+    #Determine the stepsize
+    step= (end-start)/nsteps
+    
+    #Run rk4_step
+    t=start
+    import time
+
+    i=0
+    while i <nsteps:
+        #print y
+        y= rk4_step(y,t,step,deriv)
+        t-= step
+        i+=1
+        #time.sleep(1)
+
+    return y

### 2009-01-30 py/testode.py

@@ -1,0 +1,53 @@
+from pylab import plot, show, semilogx, loglog
+from enhance import deriv
+from scipy import integrate
+from numpy import array,linspace
+from math import cos,sin, exp, log, log10, fabs
+from matplotlib.pyplot import *
+
+#parameters
+a=1/30.
+b=.0001
+mv=.00001#9
+m=68
+
+par1= b**2/a**2
+par2= mv/(a*m)
+
+
+#Determine start
+if (10/par1 < 1/par2):#Coulomb like
+    start= 10**2/par1
+    print "Coulomb-like"
+else:
+    start= 10**2/par2
+    print "Finite range"
+
+print start
+end=10**-4
+y0=array([cos(b/a*start),-b/a*sin(b/a*start),
+          sin(b/a*start),b/a*cos(b/a*start)])
+numsteps=10000
+time=linspace(log10(start),log10(end),numsteps)
+time=10**time
+
+y=integrate.odeint(deriv,y0,time,args=(par1,par2),mxstep=10**8)
+
+print log10(1/(y[numsteps-1,0]**2+y[numsteps-1,2]**2)), 1/(y[numsteps-1,0]**2+y[numsteps-1,2]**2)
+print log10(1/(y[numsteps-1,0]**2)), 1/(y[numsteps-1,0]**2)
+print log10(1/(y[numsteps-1,2]**2)), 1/(y[numsteps-1,2]**2)
+print (y[numsteps-1,3]**2)*a**2/b**2
+print (y[numsteps-1,1]**2)*a**2/b**2
+
+
+#extrapolate?
+print (2/(log(10.)*fabs(y[numsteps-1,0]))*y[numsteps-1,1]*end-2*log10(fabs(y[numsteps-1,0])))
+print (2/(log(10.)*fabs(y[numsteps-1,2]))*y[numsteps-1,3]*end-2*log10(fabs(y[numsteps-1,2])))
+
+
+loglog(time,(y[:,0])**2)
+loglog(time,(y[:,2])**2)
+loglog(time,(y[:,1])**2)
+loglog(time,(y[:,3])**2)
+axis([end,a**2/b**2,b**2,1.5])
+show()

### 2009-01-30 py/calc_2dboost.py
Calculating and plotting the 2d boost

@@ -1,0 +1,96 @@
+#Calculate the boost for a halo, by manually parallalizing the calculation
+#Makes matrix of total boost for a halo of mass Mh
+#function of m_V/m and alpha
+#argument1: log_10 Mh
+#argument2: grid (default: 64)
+#argument3: bin number (bins go vertical)
+
+from math import *
+from numpy import *
+#import matplotlib
+#matplotlib.use('PS')
+#from pylab import *
+#from matplotlib.pyplot import *
+from enhance import avg_enhance
+from boost import boost
+import os, sys
+import pickle
+#import matplotlib.cm as cm
+#from matplotlib import rc
+
+
+#parameters
+if len(sys.argv) > 1:
+    logMh= int(sys.argv[1])
+else:
+    print "No Mh given, returning..."
+    sys.exit(-1)
+Mh=10**logMh
+vdisp= 2.5*10**-2*pow(Mh,1./3.)/(3*10**5)
+m=1.0
+if len(sys.argv) > 2:
+    nms= int(sys.argv[2])
+    nas= int(sys.argv[2])
+else:
+    nms= 64
+    nas= 64
+nbins= 16#Hard-coded!!
+if len(sys.argv) > 3:
+    bin= int(sys.argv[3])
+else:
+    bin= 1
+mv= linspace(-5,-2,nms)
+mv= 10**mv
+a= linspace(-3,-1,nas)
+a= 10**a
+    
+#construct filenames
+savefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(bin)+'.sav'
+
+
+
+if os.path.exists(savefilename):
+    savefile=open(savefilename,'r+')
+    data=pickle.load(savefile)
+    ii= data['ii']
+    jj= data['jj']
+    if data['nms'] != nms:
+        print "Warning: suppplied nms does not equal savefile nms"
+        sys.exit(-1)
+    if data['nas'] != nas:
+        print "Warning: suppplied nas does not equal savefile nas"
+        sys.exit(-1)
+    mv= data['mv']
+    a= data['a']
+    S= data['S']
+    print 'restarting at ii= '+str(ii)+', jj= '+str(jj)+', at iteration '+str(ii*nms+jj+1)+'/'+str(nms*nas/nbins)
+else:
+    S= zeros((nas,nms))
+    ii= 0
+    jj= (bin-1)*nas/nbins
[…]

### 2009-01-30 py/plot_2dboost.py

@@ -1,0 +1,106 @@
+#Makes density plot of total boost for a halo of mass Mh
+#function of m_V/m and alpha
+#argument1: log_10 Mh
+#argument2: grid (default: 64)
+#First calculate this density matrix using calc_2dboost.py!!!
+
+from math import *
+from numpy import *
+import matplotlib
+matplotlib.use('PS')
+from pylab import *
+from matplotlib.pyplot import *
+from enhance import avg_enhance
+from boost import boost
+import os, sys
+import pickle
+import matplotlib.cm as cm
+from matplotlib import rc
+
+
+#parameters
+if len(sys.argv) > 1:
+    logMh= int(sys.argv[1])
+else:
+    print "No Mh given, returning..."
+    sys.exit(-1)
+Mh=10**logMh
+#vdisp= 2.5*10**-2*pow(Mh,1./3.)/(3*10**5)
+#m=1.0
+if len(sys.argv) > 2:
+    nms= int(sys.argv[2])
+    nas= int(sys.argv[2])
+else:
+    nms= 64
+    nas= 64
+nbins= 16#Hard-coded!!
+mv= linspace(-5,-2,nms)
+mv= 10**mv
+a= linspace(-3,-1,nas)
+a= 10**a
+
+boostS= zeros((nas,nms))
+
+#Loop over bins, restore and put in the right matrix
+for bb in range(16):
+    if bb==0:
+        continue
+    if bb>2:
+        break
+    #construct filenames
+    bin= bb+1
+    restorefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(bin)+'.sav'
+    restorefile=open(restorefilename,'r')
+    data=pickle.load(restorefile)
+    S= data['S']
+    #Load the boosts from S
+    ii= 0
+    jj= (bin-1)*nas/nbins
+    while jj < bin*nas/nbins:
+        while ii < nms:
+            boostS[ii,jj]= S[ii,jj]
+            ii= ii+1
+        jj= jj+1
+        ii= 0
+restorefile.close()
+
+
+plotfilename= '2dboost'+str(logMh)+'_'+str(nas)+'.eps'
+
+
+#Plotting parameters
+fig_width = 3.25  # width in inches
+fig_height = 2.9      # height in inches
+fig_size =  [fig_width,fig_height]
+params = {'backend': 'ps',
[…]

### 2009-01-30 py/plot_2dboost.py
Tweak

@@ -43,10 +43,6 @@ boostS= zeros((nas,nms))
 
 #Loop over bins, restore and put in the right matrix
 for bb in range(16):
-    if bb==0:
-        continue
-    if bb>2:
-        break
     #construct filenames
     bin= bb+1
     restorefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(bin)+'.sav'

### 2009-02-02 py/boost.py
Calculate the reduced boost + updated plot_ and calc_2dboost

@@ -62,7 +62,7 @@ def deriv(y,t,mv,a,gamma,einasto,n,q,A,C,K1,K2,vdisp_sat,somm_sat):
     return 1./(1.+(1.-q)*LL) * 1./t * (A*pow(q,n)*(somm+y)*LL - n*y-y*dlnLdlnM)  
 
 
-def boost(Mh,mv=10**-2,a=10**-1,gamma=1,einasto=0,n=-0.9,q=0.1,A=2.3*10**-2,C=2.5*10**-2,K1=10.5,K2=-.11,mo=10**-6):
+def boost(Mh,mv=10**-2,a=10**-1,nosomm=0,gamma=1,einasto=0,n=-0.9,q=0.1,A=2.3*10**-2,C=2.5*10**-2,K1=10.5,K2=-.11,mo=10**-6):
     """Calculates the boost factor for a given halo mass
 
     Input:
@@ -78,32 +78,32 @@ def boost(Mh,mv=10**-2,a=10**-1,gamma=1,einasto=0,n=-0.9,q=0.1,A=2.3*10**-2,C=2.
        K1      - normalization of concentration-mass relation
        K2      - exponent of concentration-mass relation
        mo      - free-streaming limit
+       nosomm  - DON'T include Sommerfeld enhancement
 
     Output:
        boost
     """
 
-    #To speed up the calculation, first find where the Sommerfeld enhancement saturates, and pass that value to the differential equation to use for smaller velocities.
-    vdisp_Mh= 2.5*10**-2*pow(Mh,1./3.)/(3*10**5)
-    vdisps=[vdisp_Mh,vdisp_Mh*10**-1,vdisp_Mh*10**-2,vdisp_Mh*10**-3,vdisp_Mh*10**-4,vdisp_Mh*10**-5,vdisp_Mh*10**-6,vdisp_Mh*10**-7,vdisp_Mh*10**-8,vdisp_Mh*10**-9,vdisp_Mh*10**-10]
-    #vdisps=[10**-2,10**-3,10**-4,10**-5,10**-6,10**-7,10**-8,10**-9,10**-10,10**-11,10**-12]
-    somms=zeros(11)
-    indx=0
-    for ii in range(11):
-        somms[ii]= avg_enhance(mv,1.,a,vdisps[ii])
-        if ii > 0:
-            if somms[ii] <= 1.1*somms[ii-1]:
-                break
-            else:
-                indx+= 1
+    if nosomm == 0:
[…]

### 2009-02-02 py/calc_2dboost.py

@@ -2,22 +2,15 @@
 #Makes matrix of total boost for a halo of mass Mh
 #function of m_V/m and alpha
 #argument1: log_10 Mh
-#argument2: grid (default: 64)
+#argument2: grid (default: 64), should be divisible by 16!
 #argument3: bin number (bins go vertical)
 
 from math import *
 from numpy import *
-#import matplotlib
-#matplotlib.use('PS')
-#from pylab import *
-#from matplotlib.pyplot import *
 from enhance import avg_enhance
 from boost import boost
 import os, sys
 import pickle
-#import matplotlib.cm as cm
-#from matplotlib import rc
-
 
 #parameters
 if len(sys.argv) > 1:
@@ -29,12 +22,12 @@ Mh=10**logMh
 vdisp= 2.5*10**-2*pow(Mh,1./3.)/(3*10**5)
 m=1.0
 if len(sys.argv) > 2:
-    nms= int(sys.argv[2])
-    nas= int(sys.argv[2])
+    nms= int(sys.argv[2])/2*3+1
+    nas= int(sys.argv[2])+1
 else:
-    nms= 64
-    nas= 64
-nbins= 16#Hard-coded!!
+    nms= 97
+    nas= 65
+nbins= 16#Hard-coded!!, actually +1
 if len(sys.argv) > 3:
     bin= int(sys.argv[3])
 else:
@@ -47,50 +40,60 @@ a= 10**a
 #construct filenames
 savefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(bin)+'.sav'
 
-
-
 if os.path.exists(savefilename):
     savefile=open(savefilename,'r+')
     data=pickle.load(savefile)
     ii= data['ii']
     jj= data['jj']
     if data['nms'] != nms:
-        print "Warning: suppplied nms does not equal savefile nms"
+        print "Warning: suplied nms does not equal savefile nms"
         sys.exit(-1)
     if data['nas'] != nas:
-        print "Warning: suppplied nas does not equal savefile nas"
+        print "Warning: suplied nas does not equal savefile nas"
         sys.exit(-1)
     mv= data['mv']
     a= data['a']
     S= data['S']
[…]

### 2009-02-02 py/calc_reducedboost.py

@@ -1,8 +1,8 @@
-#Makes density plot of total boost for a halo of mass Mh
+#Calculate the reduced boost for a halo, i.e., totalboost/(somm(Mh)*substboost)
+#Makes matrix of reduced boost for a halo of mass Mh, and plots it
 #function of m_V/m and alpha
 #argument1: log_10 Mh
 #argument2: grid (default: 64)
-#First calculate this density matrix using calc_2dboost.py!!!
 
 from math import *
 from numpy import *
@@ -25,15 +25,14 @@ else:
     print "No Mh given, returning..."
     sys.exit(-1)
 Mh=10**logMh
-#vdisp= 2.5*10**-2*pow(Mh,1./3.)/(3*10**5)
-#m=1.0
+vdisp= 2.5*10**-2*pow(Mh,1./3.)/(3*10**5)
+m=1.0
 if len(sys.argv) > 2:
     nms= int(sys.argv[2])/2*3+1
     nas= int(sys.argv[2])+1
 else:
     nms= 97
     nas= 65
-nbins= 16#Hard-coded!!
 mv= linspace(-5,-2,nms)
 mv= 10**mv
 a= linspace(-3,-1,nas)
@@ -41,10 +40,15 @@ a= 10**a
 
 boostS= zeros((nms,nas))
 
+#########################################################
+#
+# First get the total boosts from the right files
+#
+#########################################################
+
 #Loop over bins, restore and put in the right matrix
+nbins= 16
 for bb in range(nbins):
-#    if (bb > 0 and bb < 8):
-#        continue
     #construct filenames
     bin= bb+1
     restorefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(bin)+'.sav'
@@ -75,9 +79,53 @@ while ii < nms:
 restorefile.close()
 
 
-#Then plot
-plotfilename= '2dboost'+str(logMh)+'_'+str(nas)+'.eps'
+#Then calculate the substructure boost for this halo mass, without the Sommerfeld enhancement
+subboost= boost(Mh,1,1,1)
 
+#Then calculate the reduced Sommerfeld enhancement for each grid point
+#filename
+savefilename= '2dredboost'+str(logMh)+'_'+str(nas)+'.sav'
+
+if os.path.exists(savefilename):
+    savefile=open(savefilename,'r+')
[…]

### 2009-02-02 tex/somm.tex
Added table with dSph properties

@@
 \newcommand{\Msol}{\ensuremath{M_{\odot}}}
+\newcommand{\Msun}{\Msol}
 \newcommand{\Mvir}{\ensuremath{M_{\mbox{{\footnotesize vir}}}}}
@@
 
+\begin{figure*}[t]
+\includegraphics[clip=]{2dboost8_65.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost7_65.eps}\\%%BoundingBox: 190 291 418 474
+%\includegraphics[clip=]{avg-5_128.eps}%%BoundingBox: 178 291 400 474
+%\includegraphics[clip=]{avg-6_128.eps}%%BoundingBox: 190 291 418 474
+%\caption{Sommerfeld enhancement factor as a function of the particle
+%physics parameters averaged over a Maxwell-Boltzman velocity
+%distribution with velocity dispersion \sigv\ for four different orders
+%of magnitude of the velocity dispersion (all velocities are relative
+%to the speed of light).}%
+\label{fig:subboost}%
+\end{figure*}
+
+\begingroup
+\squeezetable
+\begin{table}[t]
+\begin{ruledtabular}
+\begin{tabular}{lrr@{.}lr@{.}lr@{.}lr@{ $\times$ }l}
+dSph & \dist\ & \multicolumn{2}{c}{\sigv}\ & \multicolumn{2}{c}{\rhos} & \multicolumn{2}{c}{\rs} & \multicolumn{2}{c}{$\rhos^2 \rs^3$}\\
+& [kpc] & \multicolumn{2}{c}{[km s$^{-1}$]} & \multicolumn{2}{c}{[$\Msun$ kpc$^{-3}$]} & \multicolumn{2}{c}{[kpc]} & \multicolumn{2}{c}{[$\Msun^2$ kpc$^{-3}$]}\\[4pt]
+\hline
+Ursa Minor.......... & 66 & 15&0 & 1&2 $\times$ $10^8$ & 0&6 & 3& 10$^{15}$\\
+Draco................... & 80 & 5&5 &&&&&&\\
+Sculptor............... & 79 & 8&5&&&&&&\\
+Fornax................. & 138 & 11&1&&&&&&\\
+Carina.................. & 101 & 6&8 &&&&&&\\
+Sextans................ & 86 & 5&8 &&&&&&\\
+Ursa Major II....... & 32 & && 3&0 $\times$ $10^8$& 0&3  & 8&10$^{14}$\\
+Coma Berenices.... & 44 &&& 3&0 $\times$ $10^8$ & 0&3 & 8&10$^{14}$\\
+Willman I............. & 38 &&& 5&5 $\times$ $10^8$ & 0&2 & 1&10$^{15}$\\
+Segue I.................. & 23 &&&&
+\end{tabular}
+\end{ruledtabular}
+\caption{Properties of the dwarf Spheroidals used in this study.}\label{tab:dsphs}
+\end{table}
+\endgroup
+
 \section{A typical distribution of signals in a halo}

### 2009-02-03 py/boost.py
Added Eiasto to calc_2dboost

@@ -26,10 +26,6 @@ def deriv(y,t,mv,a,gamma,einasto,n,q,A,C,K1,K2,vdisp_sat,somm_sat):
     Output:
        derivative
     """
-    #Diagnostics, print M and B(M)
-    #print t
-    #print y
-
     #Calculate each factor/term in turn
     #First calculate the concentration
     conc= K1*pow(t,K2)/pow(10.,K2*12)

### 2009-02-03 py/calc_2dboost.py

@@ -4,6 +4,7 @@
 #argument1: log_10 Mh
 #argument2: grid (default: 64), should be divisible by 16!
 #argument3: bin number (bins go vertical)
+#argument4: Einasto alpha, or gamma NFW (based on the value)
 
 from math import *
 from numpy import *
@@ -32,13 +33,25 @@ if len(sys.argv) > 3:
     bin= int(sys.argv[3])
 else:
     bin= 1
+if len(sys.argv) > 4:
+    gamma= double(sys.argv[4])
+    if gamma > 0.4:
+        einasto= 0
+    else:
+        einasto= 1
+else:
+    gamma= 1.
+    einasto= 0
 mv= linspace(-5,-2,nms)
 mv= 10**mv
 a= linspace(-3,-1,nas)
 a= 10**a
     
 #construct filenames
-savefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(bin)+'.sav'
+savefilename= '2dboost'+str(logMh)+'_'+str(nas)
+if len(sys.argv) > 4:
+    savefilename+= '_'+str(gamma)
+savefilename+='_'+str(bin)+'.sav'
 
 if os.path.exists(savefilename):
     savefile=open(savefilename,'r+')
@@ -67,7 +80,7 @@ else:
 if bin != 17:
     while jj < bin*(nas-1)/nbins:
         while ii < nms:
-            boo= boost(Mh,mv[ii],a[jj])
+            boo= boost(Mh,mv[ii],a[jj],0,gamma,einasto)
             so= avg_enhance(mv[ii],m,a[jj],vdisp)
             S[ii,jj]= log10(so+boo)
             sys.stdout.write('\r'+str(ii+(jj-(bin-1)*(nas-1)/nbins)*nms+1)+'/'+str(nms*(nas-1)/nbins))
@@ -82,7 +95,7 @@ if bin != 17:
     sys.stdout.write('\n')
 else:
     while ii < nms:
-        boo= boost(Mh,mv[ii],a[jj])
+        boo= boost(Mh,mv[ii],a[jj],0,gamma,einasto)
         so= avg_enhance(mv[ii],m,a[jj],vdisp)
         S[ii,jj]= log10(so+boo)
         sys.stdout.write('\r'+str(ii+(jj-(bin-1)*(nas-1)/nbins)*nms+1)+'/'+str(nms*(nas-1)/nbins))

### 2009-02-03 py/calc_reducedboost.py

@@ -100,7 +100,7 @@ if os.path.exists(savefilename):
     mv= data['mv']
     a= data['a']
     redboostS= data['redboostS']
-    print 'restarting at ii= '+str(ii)+', jj= '+str(jj)+', at iteration '+str(ii*nms+jj+1)+'/'+str(nms*nas)
+    print 'restarting at ii= '+str(ii)+', jj= '+str(jj)+', at iteration '+str(ii*nas+jj+1)+'/'+str(nms*nas)
 else:
     redboostS= zeros((nms,nas))
     ii= 0
@@ -112,7 +112,7 @@ while ii < nms:
         somm= avg_enhance(mv[ii],m,a[jj],vdisp)
         redboostS[ii,jj]= boostS[ii,jj]-log10(somm*(1.0+subboost))
         #print boostS[ii,jj], log10(somm*(1+subboost))
-        sys.stdout.write('\r'+str(ii*nms+jj+1)+'/'+str(nms*nas))
+        sys.stdout.write('\r'+str(ii*nas+jj+1)+'/'+str(nms*nas))
         sys.stdout.flush()
         jj= jj+1
         data={'mv':mv,'a':a,'redboostS':redboostS,'nms':nms,'nas':nas,'ii':ii,'jj':jj}
@@ -147,7 +147,7 @@ im= imshow(redboostS,origin='lower',#cmap=cm.gray,
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)
 xlabel(r'$\log_{10}\alpha$')#,fontsize=16)
-text(-2,-4.75,r'$M = 10^{'+str(logMh)+r'} M_{\odot}$')#,fontsize=16)
+text(-2,-4.75,r'$M = 10^{'+str(logMh)+r'} M_{\odot}$',color='w')#,fontsize=16)
 CB1= colorbar(im,shrink=0.87)
 levels=linspace(0,5,11)
 #cont= contour(data['S'],levels,origin='lower',linewidths=1,colors='k',
@@ -157,7 +157,7 @@ levels=linspace(0,5,11)
 #Move the 2 colorbars on top of each other
 #ll,bb,ww,hh = CB2.ax.get_position()
 #CB1.ax.set_position([ll, bb, ww, hh])
[…]

### 2009-02-13 py/2dboost.py

@@ -27,11 +27,11 @@ Mh=10**logMh
 vdisp= 2.5*10**-2*pow(Mh,1./3.)/(3*10**5)
 m=1.0
 if len(sys.argv) > 2:
-    nms= int(sys.argv[2])
-    nas= int(sys.argv[2])
+    nms= int(sys.argv[2])/2*3+1
+    nas= int(sys.argv[2])+1
 else:
-    nms= 64
-    nas= 64
+    nms= 97
+    nas= 65
 mv= linspace(-5,-2,nms)
 mv= 10**mv
 a= linspace(-3,-1,nas)
@@ -56,9 +56,9 @@ if os.path.exists(savefilename):
     mv= data['mv']
     a= data['a']
     S= data['S']
-    print 'restarting at ii= '+str(ii)+', jj= '+str(jj)+', at iteration '+str(ii*nms+jj+1)+'/'+str(nms*nas)
+    print 'restarting at ii= '+str(ii)+', jj= '+str(jj)+', at iteration '+str(ii*nas+jj+1)+'/'+str(nms*nas)
 else:
-    S= zeros((nas,nms))
+    S= zeros((nms,nas))
     ii= 0
     jj= 0
     savefile=open(savefilename,'w')
@@ -73,7 +73,7 @@ while ii < nms:
 #        print so
         S[ii,jj]= log10(so+boo)
 #        print S[ii,jj]
-        sys.stdout.write('\r'+str(ii*nms+jj+1)+'/'+str(nms*nas))
+        sys.stdout.write('\r'+str(ii*nas+jj+1)+'/'+str(nms*nas))
         sys.stdout.flush()
         jj= jj+1
         data={'mv':mv,'a':a,'S':S,'nms':nms,'nas':nas,'ii':ii,'jj':jj}

### 2009-02-13 py/2denhance.py

@@ -31,25 +31,25 @@ if os.path.exists(savefilename):
     savefile=open(savefilename,'r+')
     data=pickle.load(savefile)
 else:
-    b=10**-3
+    b=10**logb
     m=1.0
     if len(sys.argv) > 2:
-        nms= int(sys.argv[2])
-        nas= int(sys.argv[2])
+        nms= int(sys.argv[2])/2*3+1
+        nas= int(sys.argv[2])+1
     else:
-        nms= 64
-        nas= 64
+        nms= 97
+        nas= 65
     mv= linspace(-5,-2,nms)
     mv= 10**mv
     a= linspace(-3,-1,nas)
     a= 10**a
     
-    S= zeros((nas,nms))
+    S= zeros((nms,nas))
     
     for ii in range(nms):
         for jj in range(nas):
             S[ii,jj]= log10(enhance(mv[ii],m,a[jj],b))
-            sys.stdout.write('\r'+str(ii*nms+jj+1)+'/'+str(nms*nas))
+            sys.stdout.write('\r'+str(ii*nas+jj+1)+'/'+str(nms*nas))
             sys.stdout.flush()
     sys.stdout.write('\n')
 

### 2009-02-13 py/avgenhance.py

@@ -56,7 +56,7 @@ if os.path.exists(savefilename):
     S= data['S']
     print 'restarting at ii= '+str(ii)+', jj= '+str(jj)+', at iteration '+str(ii*nas+jj+1)+'/'+str(nms*nas)
 else:
-    S= zeros((nas,nms))
+    S= zeros((nms,nas))
     ii= 0
     jj= 0
     savefile=open(savefilename,'w')

### 2009-02-13 py/calc_dsph.py

@@ -63,6 +63,18 @@ if len(sys.argv) > 1:
         M= 1.23*10**8
         D= 82.
         vdisp= 9.5
+    elif sys.argv[1] == "coma":
+        rs= 0.3
+        rhos= 3.*10**8
+        M= 1.97*10**8
+        D= 44.
+        vdisp= 4.6
+    elif sys.argv[1] == "umaii":
+        rs= 0.3
+        rhos= 3*10**8
+        M= 1.97*10**8
+        D= 32.
+        vdisp= 6.7
 else:
     print "No dsph given, returning..."
     sys.exit(-1)

### 2009-02-13 py/hist_VL.py

@@ -11,6 +11,7 @@ import matplotlib
 matplotlib.use('PS')
 from pylab import *
 from matplotlib.pyplot import *
+from matplotlib.axes import *
 from matplotlib import rc
 from math import log, cos, sin, sqrt, log10
 from enhance import avg_enhance
@@ -48,6 +49,7 @@ if os.path.exists(savefilename):
     data=pickle.load(savefile)
     sublum=data['sublum']
     sublumobs=data['sublumobs']
+    totallum=data['totallum']
 else:
     #first calculate the masses of the subhalos and sort by ascending halo mass
     masses= zeros(VLdata.shape[0])
@@ -119,15 +121,95 @@ else:
             sublumobs[ii,jj,3]= log10(sublum[jj,4]/(sublum[jj,6]*(1.+sublum[jj,5])))
             sublumobs[ii,jj,0]= log10(sublumobs[ii,jj,0])
             sublumobs[ii,jj,4]= sublum[jj,7]
-            sublumobs[ii,jj,4]= sublum[jj,8]
-            sublumobs[ii,jj,4]= sublum[jj,9]
+            sublumobs[ii,jj,5]= sublum[jj,8]
+            sublumobs[ii,jj,6]= sublum[jj,9]
+    #Also generate 1,000 random observers and histogram the total luminosity
+    nobs2= 10000
+    totallum= zeros(nobs2)
+    for ii in range(nobs2):
+        #Generate random observer
+        phi= numpy.random.uniform(0.,2.*pi)
+        theta= numpy.random.uniform(0.,pi)
+        pos= array([obsR*sin(theta)*sin(phi), obsR*sin(theta)*cos(phi), obsR*cos(theta)])
+        for jj in range(nlums):
+            dist2= pow((sublum[jj,0]-pos[0]),2.)+pow((sublum[jj,1]-pos[1]),2.)+pow((sublum[jj,2]-pos[2]),2.)
+            totallum[ii]+= sublum[jj,3]*sublum[jj,4]/dist2
+        totallum[ii]= log10(totallum[ii])
     #Save
     savefile=open(savefilename,'w')
-    data={'sublum':sublum,'sublumobs':sublumobs}
[…]

### 2009-02-13 py/plot_dsph.py

@@ -21,6 +21,30 @@ from matplotlib import rc
 
 #parameters
 if len(sys.argv) > 1:
+    if sys.argv[1] == "fornax":
+        title="Fornax"
+    elif sys.argv[1] == "leoI":
+        title="Leo I"
+    elif sys.argv[1] == "sculptor":
+        title="Sculptor"
+    elif sys.argv[1] == "leoII":
+        title="Leo II"
+    elif sys.argv[1] == "sextans":
+        title="Sextans"
+    elif sys.argv[1] == "carina":
+        title="Carina"
+    elif sys.argv[1] == "ursaminor":
+        title="Ursa Minor"
+    elif sys.argv[1] == "draco":
+        title="Draco"
+    elif sys.argv[1] == "coma":
+        title="Coma Berenices"
+    elif sys.argv[1] == "umaii":
+        title="Ursa Major II"
+else:
+    print "No dsph given, returning..."
+    sys.exit(-1)
+if len(sys.argv) > 1:
     print "Plotting "+sys.argv[1]
 else:
     print "No Mh given, returning..."
@@ -124,6 +148,7 @@ levels=linspace(0,5,11)
 #ll,bb,ww,hh = CB2.ax.get_position()
 #CB1.ax.set_position([ll, bb, ww, hh])
 CB1.set_label(r'$\log_{10}\frac{\mathrm{d} N_{\gamma}}{\mathrm{d} \mathrm{A}\, \mathrm{d}t} [\mathrm{cm}^{-2}\mathrm{ s}^{-1}]$')#,fontsize=16)
+suptitle(title,x=0.6,y=0.93)
 
 savefig(plotfilename,format='eps')
 

### 2009-02-13 tex/somm.tex
Updated draft

--- 2009-01-30 tex/somm.tex
@@
 \newcommand{\jcap}{J.~Cosmol.~Astropart.~Phys.}
+\newcommand{\araa}{Ann.~Rev.~Astron.~Astrophys.}
 
@@
 \begin{abstract}
-We discuss the implications of the Sommerfeld enhancement of the dark
-matter annihilation cross section for the detection of dark matter
-annihilation from subhalos in the Galactic halo. A robust prediction
-of \CDM\ is that the Galactic halo is teeming with cold substructure,
-and that the subhalos themselves contain even smaller substructure,
-leading to an approximately self-similar distribution of substructure
-all the way down to earth-mass halos. In addition to the boost to the
-dark matter annihilation cross section from the high densities of
-these subhalos, an additional boost due to Sommerfeld enhancement
-results from the fact that they are kinematically cold. We take
-subhalos from the large numerical $N$-body simulation \VL\ to study
-the expected distribution of dark matter annihilation signals from
-subhalos at a Sun-like position in a large dark matter halo hosting a
-late-type galaxy. We also study the expected Sommerfeld enhanced
-annihilation signal for the smallest, most-dark-matter-dominated, and
-kinematically coldest dwarf spheroidal satellite galaxy Segue
-1. Finally, we discuss the prospects for detecting annihilation
-signals from microhalos.
+I discuss the implications of the recently introduced Sommerfeld
+enhancement of the dark matter annihilation cross section for the
+detection of subhalos in the Galactic halo.  In addition to the boost
+to the dark matter annihilation cross section from the high densities
+of these subhalos with respect to the main halo, an additional boost
+due to Sommerfeld enhancement results from the fact that they are
+kinematically colder than the Galactic halo. If we further believe the
+generic prediction of \CDM\ that in each subhalo there is an abundance
+of substructure which is approximately self-similar to that of the
+Galactic halo, then additional boosts coming from the density
+enhancements of these small substructures and their small velocity
+dispersions enhance the dark matter annihilation cross section even
+further. I show that the main effect of these substructure boosts is
+that the Sommerfeld enhancement is pushed to its saturation level in
+all the interesting regions of parameters space, and that very large
+boost factors ($10^6$ -- $10^9$) are obtained in a large class of
+models. The implications of these boost factors for the detection of
+dark matter annihilation from dwarf Speroidal galaxies in the Galactic
+halo are such that many of the models proposed to explain the positron
+excess seen by the PAMELA experiment are already marginally excluded.
 \end{abstract}
@@
 \section{Introduction}
-
-First two paragraphs straight from my paper with Glennys:
-
 In the standard flat, Gaussian, adiabatic, and scale-invariant \LCDM\
@@
 \rangle_{\mathrm{ann}} \approx 3 \times 10^{-26} $ cm$^3$ s$^{-1}$
-needed to lead to an observable annihilation flux. The boost factors
+needed to lead to an observable annihilation flux. The boost factor
 from the clumpiness of the dark matter halo is expected to be small
@@
 gamma-ray background today and the cosmic microwave background
-\cite{2008arXiv0810.3233K}.
+\cite{2008arXiv0810.3233K}. Since the typical velocities of dark
+matter particles in the Galactic halo today are of the order of the
+velocity dispersion of the halo, which is non-relativistic ($\beta\sim
+10^{-3}$), the cross section of dark matter particles annihilating in
+the halo could be significantly enhanced by this Sommerfeld enhancement.
 
-BOVY: Say something here about clumpiness of the halo: much more than
-the following: \cite{1998MNRAS.300..146G,johnston98a,2003MNRAS.339..834H,2008ApJ...688..277T,2002MNRAS.330..778W,2002MNRAS.330..792K,koposov,Koposov:2009ru}
+Both observations
+\cite{1994Natur.370..194I,1999Natur.402...53H,2005ApJ...626L..85W,2007ApJ...654..897B,koposov}
+as well as numerical simulations
+\cite{1998MNRAS.300..146G,johnston98a,1999ApJ...524L..19M,1999ApJ...522...82K,1999MNRAS.307..495H,2003MNRAS.339..834H,2008Natur.454..735D,2008MNRAS.391.1685S}
+have firmly established that dark matter halos in \CDM\ are not smooth
+structures, but that they are filled with a large amount of
+substructure. These substructures are both in the form of subhalos,
+self-bound objects orbiting in the Galactic halo, as well as in
+streams, relics of tidally disrupted subhalos. Since these
+substructures are overdense regions in the halo, the dark matter
+annihilation cross section will be correspondingly higher, since
+annihilation cross sections scale as the density squared. The effects
+of these kinds of substructure boosts have been studied extensively
+during the last decade
+\cite{2004PhRvD..70b3512B,2004PhRvD..69d3501K,2007ApJ...657..262D,2007PhRvD..75h3526S,2008A&A...479..427L,2008MNRAS.384.1627P,2008ApJ...678..614S,2008ApJ...686..262K,2008Natur.456...73S,2008arXiv0809.2781G}.
+However,
+these subhalos are not only denser than the Galactic halo in which
+they orbit, they are also kinematically colder than the Galactic halo,
+with velocity dispersion typically being two orders of magnitude
+smaller \cite{1998ARA&A..36..435M,2007ApJ...670..313S}. This means
+that the Sommerfeld enhancement in these subhalos can be many orders
+of magnitude larger than the Sommerfeld enhancement acting in the
+Galactic halo.
+
+In addition to this larger boost due to the small
+velocity dispersion in the subhalos, these subhalos themselves are
+filled with an abundance of substructure which is approximately
+self-similar to that of the main halo. Each of these even smaller
+substructure has a larger density than the subhalo they are orbiting
+in (quantified by the well established concentration-mass
+relationships for \CDM\ halos
+\cite{2001MNRAS.321..559B,2001ApJ...554..114E}) as well as a smaller
+velocity dispersion. This opens up the possibility that the dark
+matter annihilation signal in these already cold subhalos can be
+boosted even further by contribution of their substructure. In this
+paper we quantify the magnitude of this new substructure boost and
+discuss its implications for the detection of dark matter annihilation
+from subhalos in the Galactic halos.
 
@@
 \begin{figure*}[t]
-\includegraphics[clip=]{avg-3_64.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{avg-4_128.eps}\\%%BoundingBox: 190 291 418 474
-\includegraphics[clip=]{avg-5_128.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{avg-6_128.eps}%%BoundingBox: 190 291 418 474
+\includegraphics[clip=]{avg-3_65.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{avg-4_129.eps}\\%%BoundingBox: 190 291 418 474
+\includegraphics[clip=]{avg-5_129.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{avg-6_129.eps}%%BoundingBox: 190 291 418 474
 \caption{Sommerfeld enhancement factor as a function of the particle
@@
 \begin{equation}\label{eq:Mvirdef}
-\Mvir = 4\pi\rhos\rs^3 \left[ - (-\conc)^\gamma\,\conc^{-\gamma} B(-c;3-\gamma,\gamma-2)\right]\, ,
+\Mvir = 4\pi\rhos\rs^3 f(c)\, ,
 \end{equation}
-in which $B(z;a,b)$ is the incomplete Beta function, given by the
-familiar form $f(c) = \ln(1+c)-c/(1+c)$ in the special case of the
-\NFW\ profile. The virial mass for the Einasto profile is given by
+in which the function $f(c)$ is given by
+\begin{equation}
+f(c) = \frac{c^{3-\gamma}}{3-\gamma} \, {_2F_1}\left(3-\gamma,3-\gamma;4-\gamma;-c\right)\, .
+\end{equation}\label{eq:hypergeometric}
+The function $_2 F_1\left(a,b;c;z\right)$ is the Gauss hypergeometric
+function. In the $\gamma = 1$ case $f(c)$ is given by the familiar
+form $f(c) = \ln(1+c)-c/(1+c)$. The virial mass for the Einasto
+profile is given by
 \begin{equation}
@@
 \end{equation}
-in which $x \equiv r/\rminustwo$ and $f(x)$ is given by
-$B(-x;3-\gamma,\gamma-2)$ in the case of a \GNFW\ profile and by
+in which $x \equiv r/\rminustwo$ and $f(x)$ is defined in \eqnname\
+(\ref{eq:hypergeometric}) in the case of a \GNFW\ profile and is given
+by
 $\gamma\left[\frac{3}{\alphaEinasto},\frac{2}{\alphaEinasto}x^\alphaEinasto\right]$
-in the case of an Einasto profile. One can find \rmax\ by (numerically) solving the equation
+in the case of an Einasto profile. One can find \rmax\ by
+(numerically) solving the equation
 \begin{equation}
@@
 \cite{1998MNRAS.300..146G,2008ApJ...673..226P,2004MNRAS.355..794H,2005ApJ...635..931B,2004ApJ...608..663K}. This
-is a small effect, of order 10\%, and we will investigate later
-whether it significantly influences our results.
+is a small effect, of order 10\%, and it does not play a large role in
+the results we obtain below. This concentration-halo mass
+concentration has not just been establised in numerical simulations:
+statistical weak lensing analysis of luminous red galaxies and
+clusters from the Sloan Digital Sky Survey has shown that the halo
+profiles of these objects are consistent with the \NFW\ profile and
+mass-concentration relation given above \cite{2008JCAP...08..006M}.
 
@@
 Therefore, we will use (as a rough approximation) the scaling relation
-given in \eqnname\ (\ref{eq:vdispscaling}).
+given in \eqnname\ (\ref{eq:vdispscaling}). We calibrate this
+relationship at the mass scale of the dwarf spheroidals by using the
+velocity and mass of Canes Venatici I from Ref.~\cite{2007ApJ...670..313S}.
 
-\section{A typical distribution of signals in a halo}
+\subsection{The total and the ``reduced'' boost}
 
-\section{Segue 1}
+\begin{figure*}[t]
+\includegraphics[clip=]{2dboost8_65.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost7_65.eps}\\%%BoundingBox: 190 291 430 474
+\includegraphics[clip=]{2dredboost8_65.eps}%%BoundingBox: 170 277 398 487
+\includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 192 277 432 487
+\caption{}%
+\label{fig:subboost}%
+\end{figure*}
 
-\section{Microhalos}
+\begin{figure*}[t]
+\includegraphics[clip=]{2dboost8_65_0.8.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost8_65_1.4.eps}\\%%BoundingBox: 190 291 430 474
+%\includegraphics[clip=]{2dredboost8_65.eps}%%BoundingBox: 178 291 400 474
+%\includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 190 291 430 474
+\caption{}%
+\label{fig:subboostprofiles}%
+\end{figure*}
+
+\section{Predicted annihilation signals for dwarf Spheroidals}
+
+\begingroup
+%\squeezetable
+\begin{table*}[t]
+\begin{ruledtabular}
+\begin{tabular}{lrr@{.}lr@{}lr@{}lr@{.}lr@{.}lr@{.}l@{ $\times$ }lc}
+dSph & \dist\ & \multicolumn{2}{c}{\sigv}\ & \multicolumn{2}{c}{\rmax} & \multicolumn{2}{c}{\vmax} & \multicolumn{2}{c}{\rhos} & \multicolumn{2}{c}{\rs} & \multicolumn{3}{c}{$M(<\rs)$} & References\\
+& [kpc] & \multicolumn{2}{c}{[km s$^{-1}$]} & \multicolumn{2}{c}{[kpc]} & \multicolumn{2}{c}{[km s$^{-1}$]} & \multicolumn{2}{c}{[$\Msun$ kpc$^{-3}$]} & \multicolumn{2}{c}{[kpc]} & \multicolumn{3}{c}{[$\Msun$]}&\\[4pt]
+\hline
+Carina.......................................................................... & 101 & 6&8 & 3&\phantom{f}      & 16&$\!\!\!$      & 1&1 $\times$ $10^7$ & 1&4 & 7&4&$10^8$ & 1,3\\
+Coma Berenices...........................................................      & 44  & 4&6 & $\cdot$&& $\cdot$& & 3&0 $\times$ $10^8$ & 0&3 & 2&0&10$^7$ & 2,4\\
+Draco........................................................................... & 80  & 5&5 & 8&      & 40&      & 1&0 $\times$ $10^7$ & 3&7 & 1&2&$10^9$ & 1,3\\
+Fornax.........................................................................  & 138 & 11&1& 4&      & 20&      & 1&0 $\times$ $10^7$ & 1&9 & 1&5&$10^8$ & 1,3\\
+Leo I............................................................................  & 250 & 8&8 & 6&      & 30&      & 1&0 $\times$ $10^7$ & 2&8 & 5&2&$10^8$ & 1,3\\
+Leo II...........................................................................  & 205 & 6&7 & 4&      & 20&      & 1&0 $\times$ $10^7$ & 1&9 & 1&5&$10^8$ & 1,3\\
+Sculptor.......................................................................  & 79  & 8&5 & 6&      & 30&      & 1&0 $\times$ $10^7$ & 2&8 & 5&2&$10^8$ & 1,3\\
+Sextans........................................................................  & 86  & 5&8 & 2&      & 10&      & 1&0 $\times$ $10^7$ & 0&9 & 1&9&$10^7$ & 1,3\\
+Ursa Major II..............................................................   & 32  & 6&7 & $\cdot$&& $\cdot$& & 3&0 $\times$ $10^8$ & 0&3 & 2&0&$10^7$ & 2,4\\
+Ursa Minor..................................................................    & 66  & 15&0& 6&      & 30&      & 1&0 $\times$ $10^8$ & 2&8 & 5&2&$10^8$ & 1,3\\
+%Willman I............. & 38 &&& 5&5 $\times$ $10^8$ & 0&2 & 1&10$^{15}$\\
+%Segue I.................. & 23 &&&&
+\end{tabular}
+\end{ruledtabular}
+\caption{Properties of the dwarf Spheroidals used in this study.\\
+References: (1) \citet{2008ApJ...672..904P}, (2) \citet{2008ApJ...678..614S}, (3) \citet{1998ARA&A..36..435M}, (4) \citet{2007ApJ...670..313S}.}\label{tab:dsphs}
+\end{table*}
+\endgroup
+
+\begin{figure*}[t]
+\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 291 430 474
+\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{flux_leoI_65.eps}\\%%BoundingBox: 190 291 430 474
+\includegraphics[clip=]{flux_leoII_65.eps}\\%%BoundingBox: 178 291 400 474
+%\includegraphics[clip=]{flux_sculptor_65.eps}\\%%BoundingBox: 190 291 430 474
+\includegraphics[clip=]{flux_sextans_65.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 291 430 474
+\caption{}%
+\label{fig:dsph}%
+\end{figure*}
+
+\section{The annihilation signal from subhalos in \VL\ 2}
+
+\begin{figure}[t]
+\centering
+\includegraphics{hist_-2_-3_10_1.eps}
+\caption{Distances and masses in \VL\ 2.}%\
+\label{fig:VL23_1}%
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics{hist_-2_-3_10_2.eps}
+\caption{Photon count for individual subhalos and total boosts for \ldots.}%\
+\label{fig:VL23_2}%
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics{hist_-2_-3_10_3.eps}
+\caption{Distribution of total photon counts for 1,000 random observers at 8.5 kpc.}%\
+\label{fig:VL23_3}%
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics{hist_-5_-1_10_2.eps}
+\caption{}%\
+\label{fig:VL51_1}%
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics{hist_-5_-1_10_3.eps}
+\caption{}%\
+\label{fig:VL51_2}%
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics{hist_-3.84375_-2.625_10_2.eps}
+\caption{}%\
+\label{fig:VLresonance1}%
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics{hist_-3.84375_-2.625_10_3.eps}
+\caption{}%\
+\label{fig:VLresonance2}%
+\end{figure}
+
+\section{Discussion}
+
+Don't forget to mention the fact that the whole halo might be build up
+of distinct streams with velocity dispersion of order 1 km/s, such
+that the annihilation signal we might expect from the halo could be
+much larger than that people have suggested.
 
 {\bf Acknowledgments:}
-NASA 07-ADP07-0099.
+Hogg. NASA NNX08AJ48G.
 

### 2009-02-15 tex/somm.tex
Expanded Sommerfeld enhancement section, added referencesto dSphss and VL-2

@@
 \newcommand{\eg}{e.g.}
+\newcommand{\ie}{i.e.}
 \newcommand{\Fermi}{\emph{Fermi}}
@@
 \newcommand{\boost}{\ensuremath{B}}
+\newcommand{\redboost}{\ensuremath{\mathcal{B}_{\mbox{{\footnotesize red}}}}}
+\newcommand{\totalboost}{\ensuremath{\mathcal{B}}}
 \newcommand{\redr}{\ensuremath{\tilde{r}}}
@@
 \newcommand{\rmax}{\ensuremath{r_{\mbox{{\footnotesize max}}}}}
+\newcommand{\dsph}{dSph}
+\newcommand{\dsphs}{dSphs}
+
+\def\urltilda{\kern -.15em\lower .7ex\hbox{\~{}}\kern .04em}
 
@@
 all the interesting regions of parameters space, and that very large
-boost factors ($10^6$ -- $10^9$) are obtained in a large class of
+boost factors ($10^5$ -- $10^9$) are obtained in a large class of
 models. The implications of these boost factors for the detection of
@@
 
+\begin{figure}[t]
+\centering
+\includegraphics{2d-3.eps}
+\\
+\includegraphics{2d-4.eps}
+\caption{Sommerfeld enhancement as a function of the particle physics
+parameters for two different values of the relative velocity
+$\beta$. Because of the scaling properties of \eqnname\
+(\ref{eq:sommdiff2}) the behavior of the Sommerfeld enhancement as a
+function of $\log_{10}\beta$ can actually be read off along diagonals
+in the plane.}%\
+\label{fig:sommb34}%
+\end{figure}
+
 \Eqnname\ (\ref{eq:sommdiff2}) can also be numerically integrated
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics{2d-3.eps}
-\\
-\includegraphics{2d-4.eps}
-\caption{Sommerfeld enhancement as a function of the particle physics
-parameters for two different values of the relative velocity
-$\beta$. Because of the scaling properties of \eqnname\
-(\ref{eq:sommdiff2}) the behavior of the Sommerfeld enhancement as a
-function of $\log_{10}\beta$ can actually be read off along diagonals
-in the plane.}%\
-\label{fig:sommb34}%
-\end{figure}
-
 Of course, when considering the Sommerfeld enhancement of a halo of a
 certain velocity dispersion $\sigv$ we need to average the Sommerfeld
-enhancement over the velocity distribution of that halo. We will
-approximate the velocity distribution of the halo as a single Gaussian
-distribution with velocity dispersion $\sigv$, in which case we need
-to average the enhancement over a truncated Maxwell-Boltzmann
-distribution
+enhancement the distribution of relative velocities of that halo. We
+will approximate the one-particle velocity distribution of the halo as
+a single truncated Maxwell-Boltzmann distribution with velocity
+dispersion $\sigv$
 \begin{equation}\label{eq:maxboltz}
@@
 will tend to sustain the $1/v$ behavior up to a factor $\order(1)$,
-since then
-\begin{equation}
-\int_0^\infty \dd v \, v^2\, e^{-v^2/2\sigv^2}
-\somm\left/\int_0^\infty \dd v \, v^2\, e^{-v^2/2\sigv^2} \right.\propto
+as we have that
+\begin{equation}
+\int_0^\infty \dd v \, v^2\, e^{-v^2/4\sigv^2}
+\somm\left/\int_0^\infty \dd v \, v^2\, e^{-v^2/4\sigv^2} \right.\propto
 \frac{1}{\sigv}\, ,
@@
 \end{figure*}
+
+In Figure \ref{fig:sommavg} we show this averaged Sommerfeld
+enhancement for four different decades of velocity dispersion. The
+overall behavior of the Sommerfeld enhancement seen in Figure
+\ref{fig:sommb34} is indeed conserved well by the averaging: this has
+to do with the fact that the most probable relative velocity is given
+by the velocity dispersion for a Maxwell-Boltzmann relative velocity
+distribution. The Sommerfeld enhancement attains very large values for
+subhalos with very small velocity dispersions, moslty, but not
+exclusively, in the resonance regions. This means that the smallest
+subhalos that exists in \CDM\ will get large boosts to their \DM\
+annihilation from Sommerfeld enhancement.
 
@@
 \includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 192 277 432 487
-\caption{}%
+\caption{The total (defined in Eq.~[\ref{eq:totalboost}])and reduced
+boost (defined in Eq.~[\ref{eq:redboost}]) for subhalos of two
+different mass scales. The dark matter profile of all the subhalos is
+assumed to be well-described by an \NFW\ profile.}%
 \label{fig:subboost}%
@@
 %\includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 190 291 430 474
-\caption{}%
+\caption{The total boost for a $M=10^8$ \Msun\ subhalo assuming a
+\GNFW\ dark matter density profile for the subhalos.}%
 \label{fig:subboostprofiles}%
@@
 
+We define the total boost for a given subhalo mass as
+\begin{equation}\label{eq:totalboost}
+\totalboost(M) = \somm(M) + \boost(M)\, ,
+\end{equation}
+in which $\somm(M)$ is the Sommerfeld enhancement for a halo of mass
+$M$ and $\boost(M)$ is the substructure boost for a halo of that mass,
+as defined in Eq.~(\ref{eq:boost}). This total boost for subhalos of
+mass $M$ = $10^8$ \Msun\ and $M$ = $10^7$ \Msun\ calculated by
+following the procedure outlined in the previous subsection is shown
+in Figure \ref{fig:subboost} (\emph{top panels}. It is immediately
+obvious that the total boost factors can get very large in certain
+regions of the parameter space spanned by \mv/\mdm and $\alpha$:
+boosts vary between $10^2$ to $10^9$. These masses correspond to the
+typical masses of the classical and newly-discovered dwarf Spheroidals
+(\dsphs) and as such their velocity dispersions are of the order
+$10^{-5}$. The pattern of resonance regions and non-resonance regions
+that appears for the pure Sommerfeld enhancement (as shown in Figure
+\ref{fig:sommavg}) is also apparent in these plots of the total
+boosts, however, it is noteworthy that the resonances seems to be more
+highly peaked than they were in the pure Sommerfeld enhancement case,
+and that the boost the lower half of the parameter space, \ie, for
+smaller values of \mv/\mdm, seems to be higher than one would expect
+from the pure Sommerfeld enhancement combined with a boost factor from
+pure density enhancements.
+
+This is confirmed by looking at the ``reduced boost'' defined as
+\begin{equation}\label{eq:redboost}
+\redboost(M) \equiv \frac{\totalboost(M)}{\somm(M)\left(1+\boost_0(M)\right)}\, ,
+\end{equation}
+in which $\boost_0(M)$ is the boost we would find if we turned off the
+Sommerfeld enhancement in the calculation of the total boost, \ie, the
+boost we would get if we set $\somm(m)$ equal to one in \eqnname\
+(\ref{eq:boost}). This $\boost_0$ corresponds to the boost factor that
+has been used before to estimate the total boost from density
+enhancements in \dsphs\ and it is generally of order \order(10). This
+reduced boost therefore divides out the contribution of the Sommerfeld
+enhancement of the top-level subhalo and the boost from the
+overdensities in lower-level halos, and it would be equal to one if
+there were no extra boost coming from the fact that the substructures
+are kinematically colder than the main subhalo. As we can see in
+Figure \ref{fig:subboost} (\emph{lower panels}), in various parts of
+the Yukawa coupling parameter space this reduced boost is equal to one
+such that there is no extra boosts due to the lower velocity
+dispersion in the substructures. This is mainly the case for large
+values of the parameter \mv/\mdm, since the Sommerfeld enhancement, in
+the 1/$v$ regime, levels off around $\somm \sim \mv/\mdm$, and since
+the top-level subhalo has a velocity dispersion of $\sim 10^{-5}$ this
+means that the Sommerfeld enhancement has mostly saturated at these
+large values of \mv/\mdm, even in regions of parameter space in which
+$\alpha$ is small. Indeed, in the 1/$v$ regime the Sommerfeld
+enhancement is $\sim \alpha/\beta$ $\approx$ $10^{-2}$ for $\alpha =
+10^{-3}$, such that the Sommerfeld enhancement has fully saturated
+over the whole $\alpha$ range for large values of \mv/\mdm\ ($\mv/\mdm
+\sim 10^{-2}$).
+
+However, the Sommerfeld enhancement has not saturated in all regions
+of parameter space at this mass scale, and for smaller values of
+\mv/\mdm and especially in resonance regions, the reduced boost can
+reach high values, signalling a large extra boost due to the
+kinematically coldness of the substructure.  For small values of
+\mv/\mdm\ we see that we can get boosts that are an order of magnitude
+to two orders of magnitude larger than what we would expect from
+naively combining the Sommerfeld enhancement due to the velocity
+dispersion of the subhalo with the boost from density enhancements
+alone. This can be easily understood as being due to the fact that the
+Sommerfeld enhancement has not saturated yet at the mass scale of the
+top-level subhalo, such that it can be boosted up to it saturation
+point by the contribution from the kinematically colder substructure.
+
+The largest values of the reduced boost are found in the resonance
+regions. In these regions the Sommerfeld enhancement both grows faster
+with decreasing relative velocity as well as saturates at a much
+larger value of the enhancement. Both of these effects conspire to
+give very large extra boosts. The resonance regions are clearly
+distinguishable in this plot of the reduced boost as 45$\deg$
+lines. Because the resonance regions are so narrow the apparent extra
+boosts in different resonances seems to vary greatly, but this is
+mostly due to the finite resolution of this figure. Extra boosts up to
+$\sim 10^5$ are obtained in these resonances, and these extra boosts
+increase along the resonances when decreasing $\alpha$, again due to
+the higher saturation level for smaller $\mv/\mdm$ values.
+
+The total boost factors obtained by assuming an \NFW\ dark matter
+density profile do not depend much on the inner slope of the density
+profile. This is mainly because the calculation of the boost only
+considers \emph{relative} luminosities, for which the difference in
+normalization of the $\lum \propto \rhos^2 \rs^3/\dist^2$ relation for
+different inner slopes is irrelevant, and because the function $f(c)$
+is only a weak function of the inner slope. This is confirmed by
+calculating the total boost for a $M = 10^8$ \Msun subhalo for a
+subhalo with a very shallow inner cusp as well as for a subhalo with a
+very steep inner slope. These are shown in Figure
+\ref{fig:subboostprofiles}, and, from comparing this with the total
+boost shown in Figure \ref{fig:subboost}, it is clear that the inner
+slope does not affect the total boost due to substructure. The main
+difference between the different dark matter density profiles is
+therefore given by the normalization of the $\lum \propto \rhos^2
+\rs^3/\dist^2$ relation, shown in Figure \ref{fig:normgammaNFW}.
+
+BOVY: describe different slopes $n$ here.
+
+BOVY: Also describe different thermal free streaming limits here?
+
 \section{Predicted annihilation signals for dwarf Spheroidals}
+
+dSphs good candidates for \DM\ annihilation:
+2000PhRvD..61b3514B,%Baltz/SUSY
+2002PhRvD..66b3509T,%Tyler,Draco
+2004PhRvD..70b3512B,%Baltz,general
+2004PhRvD..69l3501E,%Evans,general
+2004PhRvD..70d3503P,%Peirani, Draco, Sagit, M31, M87.
+2004PhRvL..93p1302H,%Hooper: Sagit: 511 keV
+2006PhRvD..73f3510B,%Bergstrom,Draco
+2006JCAP...03..003P,%Profumo, Draco, SUSY
+2007PhRvD..75b3513C,%Colafrancesco,Draco
+2007PhRvD..76l3509S,%Sanchez-Conde,Draco
+2007PhRvD..75h3526S,%StrigariI
+2008ApJ...678..614S,%StrigariII
+2008arXiv0809.2781G,%Geha
+2008arXiv0812.1494P,%Pieri, no Draco, Umi, Carina or Sextans with Fermi
+
+Limits:
+2008ApJ...678..594W,%Wood: Whipple: Draco and Umi
+2008ApJ...679..428A,%Albert et al, MAGIC: <10^-11
+2008APh....29...55A,%HESS: Sagg: 
+2008PhRvD..78h7101D,%Driscoll: Draco with STACEE: limit on the spectrum
+2009ApJ...691..175A,%Hess: Canis Major
+2008arXiv0810.3561M,%MAGIC: Willman 1 < 10^-12
+
+dSph modeling:
+2007PhRvD..75h3526S,%StrigariI
+2007ApJ...667L..53W,%Walker: vdisp profiles
+2008ApJ...672..904P,%Penarubia+CDM
+2009AJ....137.3109W,%Walker, EM contamination
 
@@
 
+\VL 2 data \cite{2008Natur.454..735D} \footnote{Available at
+\texttt{http://www.ucolick.org/\urltilda$\!$diemand/vl/data.html}}.
+
 \begin{figure}[t]

### 2009-02-15 py/boost.py
Added n parameter to 2dboost, fixed boost calculation for Einasto

@@ -37,13 +37,13 @@ def deriv(y,t,mv,a,gamma,einasto,n,q,A,C,K1,K2,vdisp_sat,somm_sat):
         fc= special.gammainc(3./gamma,2./gamma*pow(conc,gamma))
         LL*= pow(fc,2)#we don't need to add gamma(3/gamma) since this would get divided out
         LL/= pow(special.gammainc(3./gamma,2./gamma*pow(concq,gamma)),2)#we don't need to add gamma(3/gamma) since this would get divided out
-        fc*= special.gamma(3./gamma)
+        fc*= special.gamma(3./gamma)*1./gamma*exp((3.*log(gamma)+2.-log(8.))/gamma)
         #dlnLdlnM= 1.+3.*K2+2.*conc/fc*K2*(pow(2,3./gamma)*pow(gamma,1.-3./gamma)*pow(conc,2.)*exp(-2./gamma*pow(conc,gamma)))
         dlnLdlnM= 1.+3.*K2-2.*conc/fc*K2*(pow(conc,2.)*exp(2./gamma*(1.-pow(conc,gamma))))
         print -2.*conc/fc*K2*(pow(conc,2.)*exp(2./gamma*(1.-pow(conc,gamma))))
-        print special.gammainc(3./gamma,2./gamma*pow(conc,gamma))
-        print fc
-        print (pow(conc,2.)*exp(2./gamma*(1.-pow(conc,gamma)))), (pow(conc,2.-gamma)*pow(1.+conc,gamma-3.))
+#        print special.gammainc(3./gamma,2./gamma*pow(conc,gamma))
+#        print fc
+#        print (pow(conc,2.)*exp(2./gamma*(1.-pow(conc,gamma)))), (pow(conc,2.-gamma)*pow(1.+conc,gamma-3.))
         print dlnLdlnM
     else:
         if gamma == 1:
@@ -56,10 +56,10 @@ def deriv(y,t,mv,a,gamma,einasto,n,q,A,C,K1,K2,vdisp_sat,somm_sat):
             LL/= pow(special.hyp2f1(3.-gamma,3.-gamma,4.-gamma,-concq)*pow(concq,3.-gamma)/(3.-gamma),2)
         dlnLdlnM= 1.+3.*K2-2.*conc/fc*K2*(pow(conc,2.-gamma)*pow(1.+conc,gamma-3.))
[…]

### 2009-02-15 py/calc_2dboost.py

@@ -5,6 +5,7 @@
 #argument2: grid (default: 64), should be divisible by 16!
 #argument3: bin number (bins go vertical)
 #argument4: Einasto alpha, or gamma NFW (based on the value)
+#argument5: slope of the subhalo abundance power-law (default -0.9)
 
 from math import *
 from numpy import *
@@ -42,6 +43,20 @@ if len(sys.argv) > 4:
 else:
     gamma= 1.
     einasto= 0
+if len(sys.argv) > 5:
+    n= double(sys.argv[5])
+    #Also recalculate the normalization of the subhalo abundance function
+    q= 0.1
+    g= 10**-5
+    if n != -1:
+        A= 0.1*(n+1.)/(pow(q,n+1.)-pow(g,n+1.))
+    else:
+        A= 0.1/log(q/g)
+else:
+    n= -0.9
+    q= 0.1
+    g= 10**-5
+    A= 0.1*(n+1.)/(pow(q,n+1.)-pow(g,n+1.))
 mv= linspace(-5,-2,nms)
 mv= 10**mv
 a= linspace(-3,-1,nas)
@@ -51,6 +66,8 @@ a= 10**a
 savefilename= '2dboost'+str(logMh)+'_'+str(nas)
 if len(sys.argv) > 4:
     savefilename+= '_'+str(gamma)
+if len(sys.argv) > 5:
+    savefilename+= '_'+str(n)
 savefilename+='_'+str(bin)+'.sav'
 
 if os.path.exists(savefilename):
@@ -80,7 +97,7 @@ else:
 if bin != 17:
     while jj < bin*(nas-1)/nbins:
         while ii < nms:
-            boo= boost(Mh,mv[ii],a[jj],0,gamma,einasto)
+            boo= boost(Mh,mv[ii],a[jj],0,gamma,einasto,10**-6,0.,n,0.1,A)
             so= avg_enhance(mv[ii],m,a[jj],vdisp)
             S[ii,jj]= log10(so+boo)
             sys.stdout.write('\r'+str(ii+(jj-(bin-1)*(nas-1)/nbins)*nms+1)+'/'+str(nms*(nas-1)/nbins))
@@ -95,7 +112,7 @@ if bin != 17:
     sys.stdout.write('\n')
 else:
     while ii < nms:
-        boo= boost(Mh,mv[ii],a[jj],0,gamma,einasto)
[…]

### 2009-02-16 tex/somm.tex
dSph intro

@@
 \newcommand{\rmax}{\ensuremath{r_{\mbox{{\footnotesize max}}}}}
-\newcommand{\dsph}{dSph}
-\newcommand{\dsphs}{dSphs}
+\newcommand{\dSph}{dSph}
+\newcommand{\dSphs}{dSphs}
+\newcommand{\SDSS}{SDSS}
 
@@
 
-\subsection{The Dark Matter profile}
+\subsection{The Dark Matter profile}\label{sec:DMprofile}
 
@@
 
+\begin{figure*}[t]
+\includegraphics[clip=]{2dboost8_65.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost7_65.eps}\\%%BoundingBox: 190 291 430 474 
+\includegraphics[clip=]{2dredboost8_65.eps}%%BoundingBox: 170 277 398 487 
+\includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 192 277 432 487
+\caption{The total (defined in Eq.~[\ref{eq:totalboost}])and reduced
+boost (defined in Eq.~[\ref{eq:redboost}]) for subhalos of two
+different mass scales. The dark matter profile of all the subhalos is
+assumed to be well-described by an \NFW\ profile.}%
+\label{fig:subboost}%
+\end{figure*}
+
 In the previous section we calculated $\lumsmoothc$ for the different
@@
 statistical weak lensing analysis of luminous red galaxies and
-clusters from the Sloan Digital Sky Survey has shown that the halo
-profiles of these objects are consistent with the \NFW\ profile and
-mass-concentration relation given above \cite{2008JCAP...08..006M}.
+clusters from the Sloan Digital Sky Survey (\SDSS) has shown that the
+halo profiles of these objects are consistent with the \NFW\ profile
+and mass-concentration relation given above
+\cite{2008JCAP...08..006M}.
 
@@
 \begin{figure*}[t]
-\includegraphics[clip=]{2dboost8_65.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{2dboost7_65.eps}\\%%BoundingBox: 190 291 430 474 
-\includegraphics[clip=]{2dredboost8_65.eps}%%BoundingBox: 170 277 398 487 
-\includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 192 277 432 487
-\caption{The total (defined in Eq.~[\ref{eq:totalboost}])and reduced
-boost (defined in Eq.~[\ref{eq:redboost}]) for subhalos of two
-different mass scales. The dark matter profile of all the subhalos is
-assumed to be well-described by an \NFW\ profile.}%
-\label{fig:subboost}%
-\end{figure*}
-
-\begin{figure*}[t]
 \includegraphics[clip=]{2dboost8_65_0.8.eps}%%BoundingBox: 178 291 400 474
 \includegraphics[clip=]{2dboost8_65_1.4.eps}\\%%BoundingBox: 190 291 430 474
-%\includegraphics[clip=]{2dredboost8_65.eps}%%BoundingBox: 178 291 400 474
-%\includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 190 291 430 474
+\includegraphics[clip=]{2dboost8_65_-0.7.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost8_65_-1.0.eps}%%BoundingBox: 190 291 430 474
 \caption{The total boost for a $M=10^8$ \Msun\ subhalo assuming a
@@
 typical masses of the classical and newly-discovered dwarf Spheroidals
-(\dsphs) and as such their velocity dispersions are of the order
+(\dSphs) and as such their velocity dispersions are of the order
 $10^{-5}$. The pattern of resonance regions and non-resonance regions
@@
 has been used before to estimate the total boost from density
-enhancements in \dsphs\ and it is generally of order \order(10). This
+enhancements in \dSphs\ and it is generally of order \order(10). This
 reduced boost therefore divides out the contribution of the Sommerfeld
@@
 
-dSphs good candidates for \DM\ annihilation:
-2000PhRvD..61b3514B,%Baltz/SUSY
-2002PhRvD..66b3509T,%Tyler,Draco
-2004PhRvD..70b3512B,%Baltz,general
-2004PhRvD..69l3501E,%Evans,general
-2004PhRvD..70d3503P,%Peirani, Draco, Sagit, M31, M87.
-2004PhRvL..93p1302H,%Hooper: Sagit: 511 keV
-2006PhRvD..73f3510B,%Bergstrom,Draco
-2006JCAP...03..003P,%Profumo, Draco, SUSY
-2007PhRvD..75b3513C,%Colafrancesco,Draco
-2007PhRvD..76l3509S,%Sanchez-Conde,Draco
-2007PhRvD..75h3526S,%StrigariI
-2008ApJ...678..614S,%StrigariII
-2008arXiv0809.2781G,%Geha
-2008arXiv0812.1494P,%Pieri, no Draco, Umi, Carina or Sextans with Fermi
-
-Limits:
-2008ApJ...678..594W,%Wood: Whipple: Draco and Umi
-2008ApJ...679..428A,%Albert et al, MAGIC: <10^-11
-2008APh....29...55A,%HESS: Sagg: 
-2008PhRvD..78h7101D,%Driscoll: Draco with STACEE: limit on the spectrum
-2009ApJ...691..175A,%Hess: Canis Major
-2008arXiv0810.3561M,%MAGIC: Willman 1 < 10^-12
-
-dSph modeling:
-2007PhRvD..75h3526S,%StrigariI
-2007ApJ...667L..53W,%Walker: vdisp profiles
-2008ApJ...672..904P,%Penarubia+CDM
-2009AJ....137.3109W,%Walker, EM contamination
+Besides detection of \DM\ annihilation from the Galactic Center or the
+diffuse component from the Galactic halo, there idea exists that the
+best candidates for \DM\ annihilation detection might be the dwarf
+Spheroidal satellite galaxies (\dSphs) of our Galaxy, since the dwarf
+galaxies are believed to be the most dark matter dominated structures
+in the Universe, and the \DM\ annihilation signal coming from them
+would be much easier to distinguish from background sources than, \eg,
+would be the case for \DM\ annihilation photons coming from the
+Galactic Center. That \dSphs\ make good candidates for \DM\
+annihilation detection has been much discussed before, especially for
+the ``classical'', or pre-\SDSS, \dSphs\
+\cite{2000PhRvD..61b3514B,2002PhRvD..66b3509T,2004PhRvD..70b3512B,2004PhRvD..69l3501E,2004PhRvD..70d3503P,2004PhRvL..93p1302H}. Especially
+the annihilation signal coming from Draco has been the subject of much
+research in recent years because of its relative proximity to the
+Earth
+\cite{2002PhRvD..66b3509T,2004PhRvD..70d3503P,2006PhRvD..73f3510B,2006JCAP...03..003P,2007PhRvD..75b3513C,2007PhRvD..76l3509S,2007PhRvD..75h3526S,2008arXiv0812.1494P}. The
+prospects for \DM\ annihilation detection from \dSphs\ has in general
+been found to be rather poor, with detection with \Fermi\ being
+unlikely \cite{2008arXiv0812.1494P}. Some of the \dSphs\ have already
+been the subject of observational searches for \DM\ annihilation and
+some limits on the \DM\ annihilation signal have already been set
+\cite{2008ApJ...678..594W,2008ApJ...679..428A,2008APh....29...55A,2008PhRvD..78h7101D,2009ApJ...691..175A,2008arXiv0810.3561M}.
+
+The \dSphs\ discovered by \SDSS\ have also been suggested to make good
+candidates for \DM\ annihilation detection, because of their very
+large mass-to-light ratios ($\sim\!10^3$)
+\cite{2008ApJ...678..614S,2008arXiv0809.2781G}. In the context of the
+Sommerfeld enhancement, these recently discovered \dSphs\ are
+especially interesting, since they also have the smallest velocity
+dispersion of all the satellites of our Galaxy. Predicting the \DM\
+annihilation signal from these low-surface brightness galaxies is
+complicated by the dearth of kinematical tracers of the \DM\
+distribution and their contamination with background stars
+\cite{2009AJ....137.3109W}. Nevertheless, detailed studies have
+constrained the \DM\ density and velocity dispersion profiles of some
+of these \dSphs\
+\cite{2007PhRvD..75h3526S,2007ApJ...667L..53W,2008ApJ...672..904P}. We
+will use some of the results on the \DM\ density and velocity profiles
+found in these studies to show the extra effect of the Sommerfeld
+enhancement on the predicted \DM\ annihilation signal from some of
+these \dSphs.
+
+We consider the \dSphs\ given in Table \ref{tab:dsphs} here. Eight of
+the \dSphs\ are classical dwarf galaxies and we use the \DM\ density
+profiles fitted to them in Ref.~\cite{2008ApJ...672..904P}. In this
+the \DM\ density profiles of the \dSphs\ were not completely
+determined by the available data ---a degeneracy between the
+parameters of the \NFW\ profile used to describe the \DM\ density and
+the King profile used to fit the luminous component of the galaxies
+was found--- and the concentration-mass relation from \CDM\
+\cite{2001MNRAS.321..559B,2001ApJ...554..114E} was used to break the
+degeneracy. Since these authors expressed their results in the form of
+the \rmax\ and \vmax\ parameters of the \DM\ density profiles, we use
+the procedure outlined in Section \ref{sec:DMprofile} to convert these
+to \rhos\ and \rs\ values. As these \vmax\ and \rmax\ parameters were
+fit to an \NFW\ \DM\ profile, we will only consider an \NFW\ profile
+for the \dSphs\ here, but the general conclusions about the importance
+of the exact profile shape also hold in these particular cases. The
+velocity dispersions of these \dSphs\ are taken from
+Ref.~\cite{1998ARA&A..36..435M} and we again model these subhalos as
+having a constant velocity dispersion.
+
+We also consider some of the newly discovered \dSphs. In particular we
+discuss the prospects for \DM\ annihilation detection from Coma
+Berenices and Ursa Major II, and we use the best fit values to the
+available data from Ref.~\cite{2008ApJ...678..614S}. Since these
+authors directly quote \rhos\ and \rs\ values we do not include \rmax\
+and \vmax\ values for these \dSphs. We use the velocity dispersions
+determined for these galaxies in Ref.~\cite{2007ApJ...670..313S}.
+
+%dSphs good candidates for \DM\ annihilation:
+%2000PhRvD..61b3514B,%Baltz/SUSY 2002PhRvD..66b3509T,%Tyler,Draco
+%2004PhRvD..70b3512B,%Baltz,general 
+%2004PhRvD..69l3501E,%Evans,general
+%2004PhRvD..70d3503P,%Peirani, Draco, Sagit, M31, M87.
+%2004PhRvL..93p1302H,%Hooper: Sagit: 511 keV
+%2006PhRvD..73f3510B,%Bergstrom,Draco 
+%2006JCAP...03..003P,%Profumo,Draco, SUSY 
+%2007PhRvD..75b3513C,%Colafrancesco,Draco
+%2007PhRvD..76l3509S,%Sanchez-Conde,Draco
+%2007PhRvD..75h3526S,%StrigariI 
+%2008ApJ...678..614S,%StrigariII
+%2008arXiv0809.2781G,%Geha 
+%2008arXiv0812.1494P,%Pieri, no Draco, Umi,Carina or Sextans with Fermi
+
+%Limits:
+%2008ApJ...678..594W,%Wood: Whipple: Draco and Umi
+%2008ApJ...679..428A,%Albert et al, MAGIC: <10^-11
+%2008APh....29...55A,%HESS: Sagg: 
+%2008PhRvD..78h7101D,%Driscoll: Draco with STACEE: limit on the spectrum
+%2009ApJ...691..175A,%Hess: Canis Major
+%2008arXiv0810.3561M,%MAGIC: Willman 1 < 10^-12
+
+%dSph modeling:
+%2007PhRvD..75h3526S,%StrigariI
+%2007ApJ...667L..53W,%Walker: vdisp profiles
+%2008ApJ...672..904P,%Penarubia+CDM
+%2009AJ....137.3109W,%Walker, EM contamination
 

### 2009-02-17 tex/somm.tex
Added to total and reduced boost section and dSph section

@@
 
+%Some voodoo
 \def\urltilda{\kern -.15em\lower .7ex\hbox{\~{}}\kern .04em}
+\renewcommand{\topfraction}{0.95}
+\renewcommand{\bottomfraction}{0.95}
+\renewcommand{\dbltopfraction}{0.95}
 
@@
 
-\subsection{Some details of the particle physics that do not particularly concern us here}
+\subsection{Some details of the particle physics that do not particularly concern us here}\label{sec:partphys}
 
@@
 strongly on the value of the logarithmic slope of the inner density
-profile, as can be seen in Figure \ref{fig:normgammaNFW}. The lower
+profile, as can be seen in Figure \ref{fig:normgammaNFW}. The bottom
 panel of Fig.~\ref{fig:normgammaNFW} shows the fraction of the \DM\
@@
 free-streaming scale, a \DM\ model-dependent quantity, which in many
-\DM\ models lies in the range 10$^{-6}$ -- 10$^{-12}$ \Msol
+\DM\ models lies in the range 10$^{-6}$ -- 10$^{-12}$ \Msol\
 \cite{1999PhRvD..59d3517S,2001PhRvD..64h3507H,2004MNRAS.353L..23G,2005PhRvD..71j3520L,2006PhRvL..97c1301P}. The
@@
 \includegraphics[clip=]{2dboost8_65_1.4.eps}\\%%BoundingBox: 190 291 430 474
-\includegraphics[clip=]{2dboost8_65_-0.7.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{2dboost8_65_-1.0.eps}%%BoundingBox: 190 291 430 474
-\caption{The total boost for a $M=10^8$ \Msun\ subhalo assuming a
-\GNFW\ dark matter density profile for the subhalos.}%
+\includegraphics[clip=]{2dboost8_65_0.1.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost8_65_0.2.eps}%%BoundingBox: 190 291 430 474
+\caption{The total boost for a $M=10^8$ \Msun\ subhalo for different shapes of the \DM\ density profile: a generalized \NFW\ profile described by an inner slope $\gamma$ (\emph{top panels}); an Einasto profile described by a shape parameter \alphaEinasto\ (\emph{bottom panels}).}%
 \label{fig:subboostprofiles}%
@@
 are kinematically colder than the main subhalo. As we can see in
-Figure \ref{fig:subboost} (\emph{lower panels}), in various parts of
+Figure \ref{fig:subboost} (\emph{bottom panels}), in various parts of
 the Yukawa coupling parameter space this reduced boost is equal to one
@@
 
+\begin{figure*}[t]
+\includegraphics[clip=]{2dboost8_65_-0.7.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost8_65_-1.0.eps}\\%%BoundingBox: 190 291 430 474
+%\includegraphics[clip=]{2dboost8_65_-3.eps}%%BoundingBox: 178 291 400 474
+%\includegraphics[clip=]{2dboost8_65_-12.eps}%%BoundingBox: 190 291 430 474
+\caption{The total boost for a $M=10^8$ \Msun\ for different slopes
+$n$ for the subhalo mass function (\emph{top panels}) and for
+different values of the dark matter thermal free streaming limit
+$m_0$ (\emph{bottom panels}).}%
+\label{fig:subboostslope}%
+\end{figure*}
+
 However, the Sommerfeld enhancement has not saturated in all regions
@@
 normalization of the $\lum \propto \rhos^2 \rs^3/\dist^2$ relation for
-different inner slopes is irrelevant, and because the function $f(c)$
-is only a weak function of the inner slope. This is confirmed by
-calculating the total boost for a $M = 10^8$ \Msun subhalo for a
-subhalo with a very shallow inner cusp as well as for a subhalo with a
-very steep inner slope. These are shown in Figure
-\ref{fig:subboostprofiles}, and, from comparing this with the total
-boost shown in Figure \ref{fig:subboost}, it is clear that the inner
-slope does not affect the total boost due to substructure. The main
-difference between the different dark matter density profiles is
-therefore given by the normalization of the $\lum \propto \rhos^2
-\rs^3/\dist^2$ relation, shown in Figure \ref{fig:normgammaNFW}.
-
-BOVY: describe different slopes $n$ here.
-
-BOVY: Also describe different thermal free streaming limits here?
-
-\section{Predicted annihilation signals for dwarf Spheroidals}
-
-Besides detection of \DM\ annihilation from the Galactic Center or the
-diffuse component from the Galactic halo, there idea exists that the
-best candidates for \DM\ annihilation detection might be the dwarf
-Spheroidal satellite galaxies (\dSphs) of our Galaxy, since the dwarf
-galaxies are believed to be the most dark matter dominated structures
-in the Universe, and the \DM\ annihilation signal coming from them
-would be much easier to distinguish from background sources than, \eg,
-would be the case for \DM\ annihilation photons coming from the
-Galactic Center. That \dSphs\ make good candidates for \DM\
-annihilation detection has been much discussed before, especially for
-the ``classical'', or pre-\SDSS, \dSphs\
-\cite{2000PhRvD..61b3514B,2002PhRvD..66b3509T,2004PhRvD..70b3512B,2004PhRvD..69l3501E,2004PhRvD..70d3503P,2004PhRvL..93p1302H}. Especially
-the annihilation signal coming from Draco has been the subject of much
-research in recent years because of its relative proximity to the
-Earth
-\cite{2002PhRvD..66b3509T,2004PhRvD..70d3503P,2006PhRvD..73f3510B,2006JCAP...03..003P,2007PhRvD..75b3513C,2007PhRvD..76l3509S,2007PhRvD..75h3526S,2008arXiv0812.1494P}. The
-prospects for \DM\ annihilation detection from \dSphs\ has in general
-been found to be rather poor, with detection with \Fermi\ being
-unlikely \cite{2008arXiv0812.1494P}. Some of the \dSphs\ have already
-been the subject of observational searches for \DM\ annihilation and
-some limits on the \DM\ annihilation signal have already been set
-\cite{2008ApJ...678..594W,2008ApJ...679..428A,2008APh....29...55A,2008PhRvD..78h7101D,2009ApJ...691..175A,2008arXiv0810.3561M}.
-
-The \dSphs\ discovered by \SDSS\ have also been suggested to make good
-candidates for \DM\ annihilation detection, because of their very
-large mass-to-light ratios ($\sim\!10^3$)
-\cite{2008ApJ...678..614S,2008arXiv0809.2781G}. In the context of the
-Sommerfeld enhancement, these recently discovered \dSphs\ are
-especially interesting, since they also have the smallest velocity
-dispersion of all the satellites of our Galaxy. Predicting the \DM\
-annihilation signal from these low-surface brightness galaxies is
-complicated by the dearth of kinematical tracers of the \DM\
-distribution and their contamination with background stars
-\cite{2009AJ....137.3109W}. Nevertheless, detailed studies have
-constrained the \DM\ density and velocity dispersion profiles of some
-of these \dSphs\
-\cite{2007PhRvD..75h3526S,2007ApJ...667L..53W,2008ApJ...672..904P}. We
-will use some of the results on the \DM\ density and velocity profiles
-found in these studies to show the extra effect of the Sommerfeld
-enhancement on the predicted \DM\ annihilation signal from some of
-these \dSphs.
-
-We consider the \dSphs\ given in Table \ref{tab:dsphs} here. Eight of
-the \dSphs\ are classical dwarf galaxies and we use the \DM\ density
-profiles fitted to them in Ref.~\cite{2008ApJ...672..904P}. In this
-the \DM\ density profiles of the \dSphs\ were not completely
-determined by the available data ---a degeneracy between the
-parameters of the \NFW\ profile used to describe the \DM\ density and
-the King profile used to fit the luminous component of the galaxies
-was found--- and the concentration-mass relation from \CDM\
-\cite{2001MNRAS.321..559B,2001ApJ...554..114E} was used to break the
-degeneracy. Since these authors expressed their results in the form of
-the \rmax\ and \vmax\ parameters of the \DM\ density profiles, we use
-the procedure outlined in Section \ref{sec:DMprofile} to convert these
-to \rhos\ and \rs\ values. As these \vmax\ and \rmax\ parameters were
-fit to an \NFW\ \DM\ profile, we will only consider an \NFW\ profile
-for the \dSphs\ here, but the general conclusions about the importance
-of the exact profile shape also hold in these particular cases. The
-velocity dispersions of these \dSphs\ are taken from
-Ref.~\cite{1998ARA&A..36..435M} and we again model these subhalos as
-having a constant velocity dispersion.
-
-We also consider some of the newly discovered \dSphs. In particular we
-discuss the prospects for \DM\ annihilation detection from Coma
-Berenices and Ursa Major II, and we use the best fit values to the
-available data from Ref.~\cite{2008ApJ...678..614S}. Since these
-authors directly quote \rhos\ and \rs\ values we do not include \rmax\
-and \vmax\ values for these \dSphs. We use the velocity dispersions
-determined for these galaxies in Ref.~\cite{2007ApJ...670..313S}.
-
-%dSphs good candidates for \DM\ annihilation:
-%2000PhRvD..61b3514B,%Baltz/SUSY 2002PhRvD..66b3509T,%Tyler,Draco
-%2004PhRvD..70b3512B,%Baltz,general 
-%2004PhRvD..69l3501E,%Evans,general
-%2004PhRvD..70d3503P,%Peirani, Draco, Sagit, M31, M87.
-%2004PhRvL..93p1302H,%Hooper: Sagit: 511 keV
-%2006PhRvD..73f3510B,%Bergstrom,Draco 
-%2006JCAP...03..003P,%Profumo,Draco, SUSY 
-%2007PhRvD..75b3513C,%Colafrancesco,Draco
-%2007PhRvD..76l3509S,%Sanchez-Conde,Draco
-%2007PhRvD..75h3526S,%StrigariI 
-%2008ApJ...678..614S,%StrigariII
-%2008arXiv0809.2781G,%Geha 
-%2008arXiv0812.1494P,%Pieri, no Draco, Umi,Carina or Sextans with Fermi
-
-%Limits:
-%2008ApJ...678..594W,%Wood: Whipple: Draco and Umi
-%2008ApJ...679..428A,%Albert et al, MAGIC: <10^-11
-%2008APh....29...55A,%HESS: Sagg: 
-%2008PhRvD..78h7101D,%Driscoll: Draco with STACEE: limit on the spectrum
-%2009ApJ...691..175A,%Hess: Canis Major
-%2008arXiv0810.3561M,%MAGIC: Willman 1 < 10^-12
-
-%dSph modeling:
-%2007PhRvD..75h3526S,%StrigariI
-%2007ApJ...667L..53W,%Walker: vdisp profiles
-%2008ApJ...672..904P,%Penarubia+CDM
-%2009AJ....137.3109W,%Walker, EM contamination
+different inner slopes is irrelevant, and because the mass of the
+subhalo is only weakly affected by the value of the inner slope ---the
+function $f(c)$ is only a weak function of the inner slope. This is
+confirmed by calculating the total boost for a $M = 10^8$ \Msun\
+subhalo for a subhalo with a very shallow inner cusp as well as for a
+subhalo with a very steep inner slope. These are shown in Figure
+\ref{fig:subboostprofiles} (\emph{top panels}), and, from comparing
+this with the total boost shown in Figure \ref{fig:subboost}, it is
+clear that the inner slope does not affect the total boost due to
+substructure. The main difference between the different dark matter
+density profiles is therefore given by the normalization of the $\lum
+\propto \rhos^2 \rs^3/\dist^2$ relation, shown in Figure
+\ref{fig:normgammaNFW}.
+
+If the \DM\ density profiles of halos is better described by an
+Einasto profile then the substructure boost is affected more strongly
+by the shape parameter \alphaEinasto\, which basically describes how
+cored the \DM\ density profile is. Changing \alphaEinasto\ has a
+greater impact on changing the overall \DM\ density profileand
+therefore also plays a larger role in the relation between the
+characteristic density, the concentration, and the mass of the
+subhalo. The total boosts for a $M$ = $10^8$ \Msun\ subhalo is shown
+in Figure \ref{fig:subboostprofiles} (\emph{bottom panels}) for two
+different values of the shape parameter \alphaEinasto\, which span the
+range found in numerical simulations (for values of \alphaEinasto\
+greater than 0.2 the behavior basically saturates at that of
+\alphaEinasto\ = 0.2). For \alphaEinasto\ = 0.1 the total boost is
+about the same as for the \GNFW\ profile shape, which is
+understandable given that the Einasto profile is more cuspy for
+smaller values of \alphaEinasto. The same dependence on the parameter
+\mv/\mdm\ as is seen for the \GNFW\ profile is displayed for the
+Einasto profile. In the case that \alphaEinasto\ is larger the
+substructure boost reaches slightly larger values than for the \GNFW\
+density profile. Thus, as was the case for the \GNFW\ profile, the
+impact of the different normalization of the
+$\rhominustwo^2\rminustwo^3/\dist^2$ for different values of
+\alphaEinasto\ will play a leading role in the difference between the
+different profile shapes.
 
@@
 %\squeezetable
-\begin{table*}[t]
+\begin{table*}
 \begin{ruledtabular}
@@
 \includegraphics[clip=]{flux_leoI_65.eps}\\%%BoundingBox: 190 291 430 474
-\includegraphics[clip=]{flux_leoII_65.eps}\\%%BoundingBox: 178 291 400 474
-%\includegraphics[clip=]{flux_sculptor_65.eps}\\%%BoundingBox: 190 291 430 474
-\includegraphics[clip=]{flux_sextans_65.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{flux_leoII_65.eps}%%BoundingBox: 178 291 400 474
 \includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 291 430 474
@@
 
+In Figure \ref{fig:subboostslope} the effect of varying two more of
+the many parameters introduced in the previous section is shown. The
+slope of the subhalo mass function $n$ was set to -0.9 before,
+corresponding to the results from the latest high-resolution numerical
+simulations. In Figure \ref{fig:subboostslope} (\emph{top panels}) we
+show the effect of both a shallower as well as a steeper subhalo mass
+function. The value of $n = -0.7$ corresponds to the shallow slope
+found in Ref.~\cite{2002PhRvD..66f3502H}. We see that the effect of
+such a shallow slope is to seriously reduce the substructure boost,
+which is simply due to the fact that for a shallower slope less
+substructure is found in smaller and more concentrated clumps. A
+steeper slope such as $n = -1$, as was found in the recent \VL\
+simulation \cite{2008ApJ...686..262K} leads to a substructure boost
+which is about an order of magnitude larger than for the fiducial
+$n=-0.9$ value.
+
+The effect of the thermal free-streaming scale of \DM\ on the
+substructure boost is shown in Figure \ref{fig:subboostslope}
+(\emph{bottom panels}). This effect is relatively large since the
+free-streaming scale sets the scale of the smallest and most
+concentrated \DM\ halos and raising or lowering this scale therefore
+changes an important part of the contribution to the substructure
+boost. For very large values of the free-streaming scale such as $m_0
+= 10^{-3}$ \Msun, the substructure boost becomes relatively small
+BOVY: BE MORE PRECISE HERE. When the free-streaming scale becomes very
+small, much of the substructure is in very small, cold clumps, such
+that the total boost becomes very large. BOVY: BE MORE PRECISE HERE.
+
+BOVY: Also describe different thermal free streaming limits here?
+
+\section{Predicted annihilation signals for dwarf Spheroidals}
+
+Besides detection of \DM\ annihilation from the Galactic Center or the
+diffuse component from the Galactic halo, there idea exists that the
+best candidates for \DM\ annihilation detection might be the dwarf
+Spheroidal satellite galaxies (\dSphs) of our Galaxy, since the dwarf
+galaxies are believed to be the most dark matter dominated structures
+in the Universe, and the \DM\ annihilation signal coming from them
+would be much easier to distinguish from background sources than, \eg,
+would be the case for \DM\ annihilation photons coming from the
+Galactic Center. That \dSphs\ make good candidates for \DM\
+annihilation detection has been much discussed before, especially for
+the ``classical'', or pre-\SDSS, \dSphs\
+\cite{2000PhRvD..61b3514B,2002PhRvD..66b3509T,2004PhRvD..70b3512B,2004PhRvD..69l3501E,2004PhRvD..70d3503P,2004PhRvL..93p1302H}. Especially
+the annihilation signal coming from Draco has been the subject of much
+research in recent years because of its relative proximity to the
+Earth
+\cite{2002PhRvD..66b3509T,2004PhRvD..70d3503P,2006PhRvD..73f3510B,2006JCAP...03..003P,2007PhRvD..75b3513C,2007PhRvD..76l3509S,2007PhRvD..75h3526S,2008arXiv0812.1494P}. The
+prospects for \DM\ annihilation detection from \dSphs\ has in general
+been found to be rather poor, with detection with \Fermi\ being
+unlikely \cite{2008arXiv0812.1494P}. Some of the \dSphs\ have already
+been the subject of observational searches for \DM\ annihilation and
+some limits on the \DM\ annihilation signal have already been set
+\cite{2008ApJ...678..594W,2008ApJ...679..428A,2008APh....29...55A,2008PhRvD..78h7101D,2009ApJ...691..175A,2008arXiv0810.3561M}.
+
+The \dSphs\ discovered by \SDSS\ have also been suggested to make good
+candidates for \DM\ annihilation detection, because of their very
+large mass-to-light ratios ($\sim\!10^3$)
+\cite{2008ApJ...678..614S,2008arXiv0809.2781G}. In the context of the
+Sommerfeld enhancement, these recently discovered \dSphs\ are
+especially interesting, since they also have the smallest velocity
+dispersion of all the satellites of our Galaxy. Predicting the \DM\
+annihilation signal from these low-surface brightness galaxies is
+complicated by the dearth of kinematical tracers of the \DM\
+distribution and their contamination with background stars
+\cite{2009AJ....137.3109W}. Nevertheless, detailed studies have
+constrained the \DM\ density and velocity dispersion profiles of some
+of these \dSphs\
+\cite{2007PhRvD..75h3526S,2007ApJ...667L..53W,2008ApJ...672..904P}. We
+will use some of the results on the \DM\ density and velocity profiles
+found in these studies to show the extra effect of the Sommerfeld
+enhancement on the predicted \DM\ annihilation signal from some of
+these \dSphs.
+
+We consider the \dSphs\ given in Table \ref{tab:dsphs} here. Eight of
+the \dSphs\ are classical dwarf galaxies and we use the \DM\ density
+profiles fitted to them in Ref.~\cite{2008ApJ...672..904P}. In this
+the \DM\ density profiles of the \dSphs\ were not completely
+determined by the available data ---a degeneracy between the
+parameters of the \NFW\ profile used to describe the \DM\ density and
+the King profile used to fit the luminous component of the galaxies
+was found--- and the concentration-mass relation from \CDM\
+\cite{2001MNRAS.321..559B,2001ApJ...554..114E} was used to break the
+degeneracy. Since these authors expressed their results in the form of
+the \rmax\ and \vmax\ parameters of the \DM\ density profiles, we use
+the procedure outlined in Section \ref{sec:DMprofile} to convert these
+to \rhos\ and \rs\ values. As these \vmax\ and \rmax\ parameters were
+fit to an \NFW\ \DM\ profile, we will only consider an \NFW\ profile
+for the \dSphs\ here, but the general conclusions about the importance
+of the exact profile shape also hold in these particular cases. The
+velocity dispersions of these \dSphs\ are taken from
+Ref.~\cite{1998ARA&A..36..435M} and we again model these subhalos as
+having a constant velocity dispersion.
+
+We also consider some of the newly discovered \dSphs. In particular we
+discuss the prospects for \DM\ annihilation detection from Coma
+Berenices and Ursa Major II, and we use the best fit values to the
+available data from Ref.~\cite{2008ApJ...678..614S}. Since these
+authors directly quote \rhos\ and \rs\ values we do not include \rmax\
+and \vmax\ values for these \dSphs. We use the velocity dispersions
+determined for these galaxies in Ref.~\cite{2007ApJ...670..313S}.
+
+In order to calculate the substructure boost to the \DM\ annihilation
+cross section we need the mass of the \dSphs. Many definitions of the
+mass of the galaxy could be used here, among the more appropriate ones
+is the mass contained inside the tidal radius, \ie, the radius inside
+of which material in the \dSph\ is safe from being tidally stripped by
+the Galaxy. However, since we are considering the \DM\ annihilation
+flux coming from within the radius \rminustwo\ we will use the mass
+contained within this radius to calculate the substructure boost. This
+is more appropriate since (1) it can be calculated straightforwardly
+from the \DM\ density profile of the \dSph\ and (2) the mass contained
+within \rminustwo\ is really the more relevant mass to calculate the
+substructure boost, since we are only interested in contributions to
+the \DM\ annihilation flux coming from within the radius \rminustwo.
+
+Because of the similarities between some of the \DM\ density profiles
+of the \dSphs\ in Table \ref{tab:dsphs} we do not show the predicted
+annihilation fluxes for all of the \dSphs\ in Figure \ref{fig:dsph},
+instead focusing on a representative sample of them. Inspection of
+Table \ref{tab:dsphs} shows that annihilation flux of the \dSphs\ not
+shown here will be similar to that of the ones shown in Figure
+\ref{fig:dsph}.
+
+It is clear from Figure \ref{fig:dsph} that the \DM\ annihilation flux
+predicted to be coming from the \dSphs\ is very large in a large class
+of models. For example, the flux predicted for Draco exceeds the upper
+limit set on the \DM\ annihilation flux set by
+\cite{2008ApJ...679..428A} for all models parametrized by \mv/\mdm\
+and $\alpha$ by many orders of magnitude. We have to stress here that
+our predicted fluxes are all for the particle physics specific
+parameters given in Section \ref{sec:partphys}, \ie, \sigmaannv\ = $3
+\times 10^{-26}$ cm$^3$ s$^{-1}$, \mdm\ = 700 GeV and a particular
+form for the \DM\ annihilation spectrum. To predict the gamma-ray flux
+for a particular model with different values of these particle physics
+parameters one is required to calculate the particle physics parameter
+specific factor to the \DM\ annihilation flux and appropriately
+rescale the values given here. Nevertheless, any model that predicts
+very large annihilation cross sections to account for the PAMELA/ATIC
+excesses, and we are thinking here specifically about models that try
+to fit the PAMELA data with very large boost factors with respect to
+the fiducial annihilation cross section, give rise to these very large
+predicted gamma-ray fluxes coming from the \dSphs. Therefore, unless
+the boosts from substructure are somehow suppressed or the assumptions
+about the subhalo mass function, concentration-mass relation for
+subhalos, and/or any of the other asumptions about subhalos beyond the
+resolution limits of the state-of-the-art numerical simulations are
+violated for physical subhalos, these models are in tension with the
+experimental limits on the gamma-ray flux from \dSphs.
+
+%dSphs good candidates for \DM\ annihilation:
+%2000PhRvD..61b3514B,%Baltz/SUSY 2002PhRvD..66b3509T,%Tyler,Draco
+%2004PhRvD..70b3512B,%Baltz,general 
+%2004PhRvD..69l3501E,%Evans,general
+%2004PhRvD..70d3503P,%Peirani, Draco, Sagit, M31, M87.
+%2004PhRvL..93p1302H,%Hooper: Sagit: 511 keV
+%2006PhRvD..73f3510B,%Bergstrom,Draco 
+%2006JCAP...03..003P,%Profumo,Draco, SUSY 
+%2007PhRvD..75b3513C,%Colafrancesco,Draco
+%2007PhRvD..76l3509S,%Sanchez-Conde,Draco
+%2007PhRvD..75h3526S,%StrigariI 
+%2008ApJ...678..614S,%StrigariII
+%2008arXiv0809.2781G,%Geha 
+%2008arXiv0812.1494P,%Pieri, no Draco, Umi,Carina or Sextans with Fermi
+
+%Limits:
+%2008ApJ...678..594W,%Wood: Whipple: Draco and Umi
+%2008ApJ...679..428A,%Albert et al, MAGIC: <10^-11
+%2008APh....29...55A,%HESS: Sagg: 
+%2008PhRvD..78h7101D,%Driscoll: Draco with STACEE: limit on the spectrum
+%2009ApJ...691..175A,%Hess: Canis Major
+%2008arXiv0810.3561M,%MAGIC: Willman 1 < 10^-12
+
+%dSph modeling:
+%2007PhRvD..75h3526S,%StrigariI
+%2007ApJ...667L..53W,%Walker: vdisp profiles
+%2008ApJ...672..904P,%Penarubia+CDM
+%2009AJ....137.3109W,%Walker, EM contamination
+
 \section{The annihilation signal from subhalos in \VL\ 2}
@@
 
-\begin{figure}[t]
+\begin{figure}
 \centering
 \includegraphics{hist_-2_-3_10_1.eps}
-\caption{Distances and masses in \VL\ 2.}%\
+\caption{The distances to and masses of subhalos within the tidal
+radius of the main halo in \VL\ 2. The distances are for three random
+observers at a distance of 8.5 kpc from the galactic center of the
+main halo.}%\
 \label{fig:VL23_1}%
@@
 
-\begin{figure}[t]
+\begin{figure}
 \centering
 \includegraphics{hist_-2_-3_10_2.eps}
-\caption{Photon count for individual subhalos and total boosts for \ldots.}%\
+\caption{Left: Histogram of the predicted gamma-ray flux coming from
+indivual subhalos contained within the tidal radius of the main halo
+in \VL\ 2 for three random observer placed at a distance of 8.5 kpc
+from the center of the main \VL\ halo for \mv/\mdm\ = 10$^{-2}$ and
+$\alpha = 10^{-3}$. Right: Histogram of the total boosts of the
+subhalos for the same values of \mv/\mdm\ and $\alpha$.}%\
 \label{fig:VL23_2}%
@@
 
-\begin{figure}[t]
+\begin{figure}
 \centering
 \includegraphics{hist_-2_-3_10_3.eps}
-\caption{Distribution of total photon counts for 1,000 random observers at 8.5 kpc.}%\
+\caption{Total gamma-ray flux from subhalos in \VL\ 2 for 1,000 random
+observers at 8.5 kpc from the center of the main \VL\ halo for
+\mv/\mdm\ = 10$^{-2}$ and $\alpha = 10^{-3}$.}%\
 \label{fig:VL23_3}%
@@
 
-\begin{figure}[t]
+\begin{figure}
 \centering
 \includegraphics{hist_-5_-1_10_2.eps}
-\caption{}%\
+\caption{Same as Figure \ref{fig:VL23_2} but for \mv/\mdm\ = 10$^{-5}$
+and $\alpha = 10^{-1}$.}%
 \label{fig:VL51_1}%
@@
 
-\begin{figure}[t]
+\begin{figure}
 \centering
 \includegraphics{hist_-5_-1_10_3.eps}
-\caption{}%\
+\caption{Same as Figure \ref{fig:VL23_3} but for \mv/\mdm\ = 10$^{-5}$
+and $\alpha = 10^{-1}$.}%
 \label{fig:VL51_2}%
@@
 
-\begin{figure}[t]
+\begin{figure}
 \centering
 \includegraphics{hist_-3.84375_-2.625_10_2.eps}
-\caption{}%\
+\caption{Same as Figure \ref{fig:VL23_2} but for a resonance:
+\mv/\mdm\ = 10$^{-3.84375}$ and $\alpha = 10^{-2.625}$.}%
 \label{fig:VLresonance1}%
@@
 
-\begin{figure}[t]
+\begin{figure}
 \centering
 \includegraphics{hist_-3.84375_-2.625_10_3.eps}
-\caption{}%\
+\caption{Same as Figure \ref{fig:VL23_3} but for a resonance:
+\mv/\mdm\ = 10$^{-3.84375}$ and $\alpha = 10^{-2.625}$.}%
 \label{fig:VLresonance2}%
@@
 
-{\bf Acknowledgments:}
-Hogg. NASA NNX08AJ48G.
+{\bf Acknowledgments:} It is a pleasure to thank Mulin Ding and David
+Hogg for comments and assistance. This research was partially
+supported by NASA (grant NNX08AJ48G).
 

### 2009-02-19 tex/somm.tex
Added VL section

@@
 \usepackage{amsmath,amssymb}
-\usepackage{subfigure,epsfig}
-%\usepackage{feynmp}
-\usepackage{float}
+%\usepackage{float}
 %\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
@@
 \includegraphics[clip=]{2dboost8_65_0.2.eps}%%BoundingBox: 190 291 430 474
-\caption{The total boost for a $M=10^8$ \Msun\ subhalo for different shapes of the \DM\ density profile: a generalized \NFW\ profile described by an inner slope $\gamma$ (\emph{top panels}); an Einasto profile described by a shape parameter \alphaEinasto\ (\emph{bottom panels}).}%
+\caption{The total boost for a $M=10^8$ \Msun\ subhalo for different
+shapes of the \DM\ density profile: a generalized \NFW\ profile
+described by an inner slope $\gamma$ (\emph{top panels}); an Einasto
+profile described by a shape parameter \alphaEinasto\ (\emph{bottom
+panels}).}%
 \label{fig:subboostprofiles}%
@@
 \includegraphics[clip=]{2dboost8_65_-1.0.eps}\\%%BoundingBox: 190 291 430 474
-%\includegraphics[clip=]{2dboost8_65_-3.eps}%%BoundingBox: 178 291 400 474
-%\includegraphics[clip=]{2dboost8_65_-12.eps}%%BoundingBox: 190 291 430 474
+\includegraphics[clip=]{2dboost8_65_-3.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost8_65_-12.eps}%%BoundingBox: 190 291 430 474
 \caption{The total boost for a $M=10^8$ \Msun\ for different slopes
@@
 the higher saturation level for smaller $\mv/\mdm$ values.
-
-The total boost factors obtained by assuming an \NFW\ dark matter
-density profile do not depend much on the inner slope of the density
-profile. This is mainly because the calculation of the boost only
-considers \emph{relative} luminosities, for which the difference in
-normalization of the $\lum \propto \rhos^2 \rs^3/\dist^2$ relation for
-different inner slopes is irrelevant, and because the mass of the
-subhalo is only weakly affected by the value of the inner slope ---the
-function $f(c)$ is only a weak function of the inner slope. This is
-confirmed by calculating the total boost for a $M = 10^8$ \Msun\
-subhalo for a subhalo with a very shallow inner cusp as well as for a
-subhalo with a very steep inner slope. These are shown in Figure
-\ref{fig:subboostprofiles} (\emph{top panels}), and, from comparing
-this with the total boost shown in Figure \ref{fig:subboost}, it is
-clear that the inner slope does not affect the total boost due to
-substructure. The main difference between the different dark matter
-density profiles is therefore given by the normalization of the $\lum
-\propto \rhos^2 \rs^3/\dist^2$ relation, shown in Figure
-\ref{fig:normgammaNFW}.
-
-If the \DM\ density profiles of halos is better described by an
-Einasto profile then the substructure boost is affected more strongly
-by the shape parameter \alphaEinasto\, which basically describes how
-cored the \DM\ density profile is. Changing \alphaEinasto\ has a
-greater impact on changing the overall \DM\ density profileand
-therefore also plays a larger role in the relation between the
-characteristic density, the concentration, and the mass of the
-subhalo. The total boosts for a $M$ = $10^8$ \Msun\ subhalo is shown
-in Figure \ref{fig:subboostprofiles} (\emph{bottom panels}) for two
-different values of the shape parameter \alphaEinasto\, which span the
-range found in numerical simulations (for values of \alphaEinasto\
-greater than 0.2 the behavior basically saturates at that of
-\alphaEinasto\ = 0.2). For \alphaEinasto\ = 0.1 the total boost is
-about the same as for the \GNFW\ profile shape, which is
-understandable given that the Einasto profile is more cuspy for
-smaller values of \alphaEinasto. The same dependence on the parameter
-\mv/\mdm\ as is seen for the \GNFW\ profile is displayed for the
-Einasto profile. In the case that \alphaEinasto\ is larger the
-substructure boost reaches slightly larger values than for the \GNFW\
-density profile. Thus, as was the case for the \GNFW\ profile, the
-impact of the different normalization of the
-$\rhominustwo^2\rminustwo^3/\dist^2$ for different values of
-\alphaEinasto\ will play a leading role in the difference between the
-different profile shapes.
 
@@
 
+The total boost factors obtained by assuming an \NFW\ dark matter
+density profile do not depend much on the inner slope of the density
+profile. This is mainly because the calculation of the boost only
+considers \emph{relative} luminosities, for which the difference in
+normalization of the $\lum \propto \rhos^2 \rs^3/\dist^2$ relation for
+different inner slopes is irrelevant, and because the mass of the
+subhalo is only weakly affected by the value of the inner slope ---the
+function $f(c)$ is only a weak function of the inner slope. This is
+confirmed by calculating the total boost for a $M = 10^8$ \Msun\
+subhalo for a subhalo with a very shallow inner cusp as well as for a
+subhalo with a very steep inner slope. These are shown in Figure
+\ref{fig:subboostprofiles} (\emph{top panels}), and, from comparing
+this with the total boost shown in Figure \ref{fig:subboost}, it is
+clear that the inner slope does not affect the total boost due to
+substructure. The main difference between the different dark matter
+density profiles is therefore given by the normalization of the $\lum
+\propto \rhos^2 \rs^3/\dist^2$ relation, shown in Figure
+\ref{fig:normgammaNFW}.
+
+If the \DM\ density profiles of halos is better described by an
+Einasto profile then the substructure boost is affected more strongly
+by the shape parameter \alphaEinasto\, which basically describes how
+cored the \DM\ density profile is. Changing \alphaEinasto\ has a
+greater impact on changing the overall \DM\ density profileand
+therefore also plays a larger role in the relation between the
+characteristic density, the concentration, and the mass of the
+subhalo. The total boosts for a $M$ = $10^8$ \Msun\ subhalo is shown
+in Figure \ref{fig:subboostprofiles} (\emph{bottom panels}) for two
+different values of the shape parameter \alphaEinasto\, which span the
+range found in numerical simulations (for values of \alphaEinasto\
+greater than 0.2 the behavior basically saturates at that of
+\alphaEinasto\ = 0.2). For \alphaEinasto\ = 0.1 the total boost is
+about the same as for the \GNFW\ profile shape, which is
+understandable given that the Einasto profile is more cuspy for
+smaller values of \alphaEinasto. The same dependence on the parameter
+\mv/\mdm\ as is seen for the \GNFW\ profile is displayed for the
+Einasto profile. In the case that \alphaEinasto\ is larger the
+substructure boost reaches slightly larger values than for the \GNFW\
+density profile. Thus, as was the case for the \GNFW\ profile, the
+impact of the different normalization of the
+$\rhominustwo^2\rminustwo^3/\dist^2$ for different values of
+\alphaEinasto\ will play a leading role in the difference between the
+different profile shapes.
+
 \begin{figure*}[t]
@@
 boost. For very large values of the free-streaming scale such as $m_0
-= 10^{-3}$ \Msun, the substructure boost becomes relatively small
-BOVY: BE MORE PRECISE HERE. When the free-streaming scale becomes very
+= 10^{-3}$ \Msun, the substructure boost becomes about an order of
+magnitude smaller. When the free-streaming scale becomes very
 small, much of the substructure is in very small, cold clumps, such
-that the total boost becomes very large. BOVY: BE MORE PRECISE HERE.
-
-BOVY: Also describe different thermal free streaming limits here?
+that the total boost becomes about an order of magnitude larger.
 
@@
 
-\section{The annihilation signal from subhalos in \VL\ 2}
-
-\VL 2 data \cite{2008Natur.454..735D} \footnote{Available at
-\texttt{http://www.ucolick.org/\urltilda$\!$diemand/vl/data.html}}.
-
 \begin{figure}
 \centering
-\includegraphics{hist_-2_-3_10_1.eps}
+\includegraphics[clip=]{hist_-2_-3_10_1.eps}
 \caption{The distances to and masses of subhalos within the tidal
-radius of the main halo in \VL\ 2. The distances are for three random
+radius of the main halo in \VL\ II. The distances are for three random
 observers at a distance of 8.5 kpc from the galactic center of the
@@
 indivual subhalos contained within the tidal radius of the main halo
-in \VL\ 2 for three random observer placed at a distance of 8.5 kpc
+in \VL\ II for three random observer placed at a distance of 8.5 kpc
 from the center of the main \VL\ halo for \mv/\mdm\ = 10$^{-2}$ and
@@
 \includegraphics{hist_-2_-3_10_3.eps}
-\caption{Total gamma-ray flux from subhalos in \VL\ 2 for 1,000 random
+\caption{Total gamma-ray flux from subhalos in \VL\ II for 1,000 random
 observers at 8.5 kpc from the center of the main \VL\ halo for
@@
 
+\section{The annihilation signal from subhalos in \VL\ II}
+
+Besides the satellite galaxies that have been observed in the Galactic
+halo \CDM\ predicts that the halo is filled with hundreds of
+individual subhalos ---a fact which gives rise to the so-called
+``missing sattelites problem''
+\cite{1999ApJ...522...82K,1999ApJ...524L..19M}. Each of these subhalos
+has a mass of the order of the masses of the observed \dSphs\ or
+smaller, and thus will be kinematically cold with respect to the main
+halo. Thus, the total \DM\ annihilation signal coming from all of the
+substructure will be boosted by the Sommerfeld enhancement described
+above, such that an appreciable \DM\ annihilation signal could be
+expected from all of the substructure. In this section we get a
+feeling for the expected signal from \DM\ substructures in a typical
+Galaxy-sized halo. If the \DM\ annihilation signal from subhalos
+without star formation could be detected this could give a clear
+signal that the \DM\ exists and would open up a whole new arena for
+testing \CDM.
+
 \begin{figure}
@@
 
+We use the distribution of \DM\ subhalos from the \VL\ II numerical
+$N$-body simulation\footnote{Available at
+\texttt{http://www.ucolick.org/\urltilda$\!$diemand/vl/data.html}}
+\cite{2008Natur.454..735D}. \VL\ II is a state-of-the-art $N$-body
+simulation of a Galaxy-sized halo since redshift 104.3. It sampled the
+region of the main halo with 1,1 $\times$ 10$^9$ particles, each with
+a mass of 4,100 \Msun. Over 40,000 subhalos can be resolved within 402
+kpc of the center of the main hala and 20,047 subhalos are detected
+which had a peak circular velocity larger than 4 km s$^{-1}$ at some
+time. The properties of the subhalos in this latter sample can be
+accurately established, \ie, they are unaffected by resolution
+effects. We select from this sample the 9,830 subhalos which lie
+within the tidal radius of the main halo, which is approximately 462
+kpc. We show the distribution of these subhalos in mass and distance
+from the Sun in Figure \ref{fig:VL23_1}. The have placed the Sun at
+random locations at a distance of 8.5 kpc of the center of the main
+halo, corresponding to a earth-based-observer-like vantage point. The
+distribution of the distances in Figure \ref{fig:VL23_1} is given for
+three of these random observers and the distribution is very similar
+for all three observers. We see that most of the subhalos have masses
+in the range 10$^4$ -- 10$^6$ \Msun, with subhalos with masses of the
+order of the \dSphs\ being relatively rare. The lower limit is of
+course set by the finite numerical resolution of the simulation and
+doesn't correspond to a physical cut-off ---indeed, the physical
+cut-off lies at the thermal free-streaming scale. The distribution of
+the distance to the subhalos shows that the distribution of the
+subhalos is strongly radially anti-biased, with a sharp cut-off due to
+our cut at the tidal radius of the main halo. This radial anti-bias is
+due to the fact that there is much more space far away from the center
+of the galaxy, and because of the strong tidal forces near the center
+of the galaxy, which strongly affects halos which pass near the center
+of the galaxy.
+
+For each halo we can estimate the velocity dispersion using the
+mass-velocity dispersion relation given in \eqnname\
+\ref{eq:vdispscaling} and we can calculate the characteristic density
+\rhos\ and the characteristic length-scale \rs\ from the values of
+\vmax\ and \rmax\ given for each subhalo, by the procedure outlined at
+the end of Sec.~\ref{sec:DMprofile}. For a given set of particle
+physics parameters \mv/\mdm\ and $\alpha$ this allows us to calculate
+the total boost to the \DM\ annihilation cross section for each of the
+subhalos, and by using the assumptions about the particle physics
+given in Sec.~\ref{sec:partphys} we can also calculate the photon flux
+from \DM\ annihilation coming from each subhalo, and the total
+gamma-ray flux coming from all of the subhalos combined.
+
+We will only cover some typical cases for the Sommerfeld enhancement
+here which suffice to show the general trend. We consider the particle
+physics parameters \mv/\mdm\ and $\alpha$ which give rise to the
+smallest Sommerfeld enahcements in Figures \ref{fig:VL23_2} and
+\ref{fig:VL23_3}, \ie, \mv/\mdm\ = 10$^{-2}$ and $\alpha$ = 0.001. The
+left panel of Figure \ref{fig:VL23_2} shows the distribution of
+gamma-ray fluxes from \DM\ annihilation for the subhalos, while the
+right panel shows the distribution of the total boost factors. As can
+be seen, the total boost factors are quite small: the typical boost
+factor is only about 30 with the highest boost factors being about
+100. The total gamma-ray flux coming from subhalos is then also quite
+small in this case, as is shown in Figure \ref{fig:VL23_3}, in which
+the total gamma-ray flux is shown for 1,000 random observer at a
+distance of 8.5 kpc from the galactic center. At most observation
+locations this total flux is about 10$^{-7}$ cm$^2$ s$^{-1}$, with the
+high-end tail not extending further than about half an order of
+magnitude from this. This is still a gamma-ray flux which is quite
+large, showing that the cumulative signal from all of the subhalos is
+dominated by the contribution from the subhalos which give rise to the
+largest gamma-ray fluxes.
+
+Figures \ref{fig:VL51_1} and \ref{fig:VL51_2} show the same for a
+different extremum of the Sommerfeld enhancement: the case in which
+\mv/\mdm\ = 10$^{-5}$ and $\alpha$ = 0.1. As we saw before, in this
+case the Sommerfeld enhancement can reach very high values, and thus
+we see that the typical \DM\ annihilation signal coming from \DM\
+subhalos is large: the typical flux for one subhalo is equal to the
+total flux in the scenario described in the previous paragraph. The
+total boost factors for most of the halos are very large, of the order
+of 10$^7$, which gives rise to this very large photon-flux. The total
+predicted flux in this case is many orders of magnitude larger than
+the total flux in the other extremum of the Sommerfeld
+enhancement. Thus, the exact value of the parameters \mv/\mdm\ and
+$\alpha$, upon which the Sommerfeld enhancement depends, matters a
+great deal for the predicted gamma-ray flux.
+
+Finally, we look at the expected signal coming from subhalos when the
+Sommerfeld enhancement is at a resonance, the same distribution as
+before are shown in Figures \ref{fig:VLresonance1} and
+\ref{fig:VLresonance2}. Again we find large boost factors and
+individual and total gamma-ray fluxes, although they are smaller than
+the values found for the \mv/\mdm\ = 10$^{-5}$ and $\alpha$ = 0.1
+parameter set. However, since these values are near a resonance in a
+region of parameter space around which the predicted signal looks more
+like that for the parameter set \mv/\mdm\ = 10$^{-2}$ and $\alpha$ =
+0.001, this again shows that whether dark halos, \ie, subhalos without
+star formation, can be detected through \DM\ annihilation products
+depends very sensitively on the particle physics parameters upon which
+the Sommerfeld enhancement depends.
+
+In conclusion we can say that in a scenario in which there is
+Sommerfeld enhancement there is much more hope to detect dark subhalos
+than in more conventional scenarios, but that the details of the
+prediction depend strongly on the underlying particle physics.
+
 \section{Discussion}

### 2009-02-19 py/boost.py
Plotting tools + updates

@@ -40,11 +40,11 @@ def deriv(y,t,mv,a,gamma,einasto,n,q,A,C,K1,K2,vdisp_sat,somm_sat):
         fc*= special.gamma(3./gamma)*1./gamma*exp((3.*log(gamma)+2.-log(8.))/gamma)
         #dlnLdlnM= 1.+3.*K2+2.*conc/fc*K2*(pow(2,3./gamma)*pow(gamma,1.-3./gamma)*pow(conc,2.)*exp(-2./gamma*pow(conc,gamma)))
         dlnLdlnM= 1.+3.*K2-2.*conc/fc*K2*(pow(conc,2.)*exp(2./gamma*(1.-pow(conc,gamma))))
-        print -2.*conc/fc*K2*(pow(conc,2.)*exp(2./gamma*(1.-pow(conc,gamma))))
+#        print -2.*conc/fc*K2*(pow(conc,2.)*exp(2./gamma*(1.-pow(conc,gamma))))
 #        print special.gammainc(3./gamma,2./gamma*pow(conc,gamma))
 #        print fc
 #        print (pow(conc,2.)*exp(2./gamma*(1.-pow(conc,gamma)))), (pow(conc,2.-gamma)*pow(1.+conc,gamma-3.))
-        print dlnLdlnM
+#        print dlnLdlnM
     else:
         if gamma == 1:
             fc= log(1.+conc)-conc/(1.+conc)
@@ -56,8 +56,8 @@ def deriv(y,t,mv,a,gamma,einasto,n,q,A,C,K1,K2,vdisp_sat,somm_sat):
             LL/= pow(special.hyp2f1(3.-gamma,3.-gamma,4.-gamma,-concq)*pow(concq,3.-gamma)/(3.-gamma),2)
         dlnLdlnM= 1.+3.*K2-2.*conc/fc*K2*(pow(conc,2.-gamma)*pow(1.+conc,gamma-3.))
 #        print (pow(conc,2.)*exp(2./gamma*(1.-pow(conc,gamma)))), (pow(conc,2.-gamma)*pow(1.+conc,gamma-3.))
-        print -2.*conc/fc*K2*(pow(conc,2.-gamma)*pow(1.+conc,gamma-3.))
-        print dlnLdlnM
+#        print -2.*conc/fc*K2*(pow(conc,2.-gamma)*pow(1.+conc,gamma-3.))
+#        print dlnLdlnM
 
     vdisp= C*pow(t,1./3.)/(3*10**5)
     if vdisp <= vdisp_sat:

### 2009-02-19 py/calc_2dboost.py

@@ -6,6 +6,7 @@
 #argument3: bin number (bins go vertical)
 #argument4: Einasto alpha, or gamma NFW (based on the value)
 #argument5: slope of the subhalo abundance power-law (default -0.9)
+#argument6: DM thermal free streaming limit (log10)
 
 from math import *
 from numpy import *
@@ -57,6 +58,10 @@ else:
     q= 0.1
     g= 10**-5
     A= 0.1*(n+1.)/(pow(q,n+1.)-pow(g,n+1.))
+if len(sys.argv) > 6:
+    mo= 10**double(sys.argv[6])
+else:
+    mo= 10**-6
 mv= linspace(-5,-2,nms)
 mv= 10**mv
 a= linspace(-3,-1,nas)
@@ -68,6 +73,8 @@ if len(sys.argv) > 4:
     savefilename+= '_'+str(gamma)
 if len(sys.argv) > 5:
     savefilename+= '_'+str(n)
+if len(sys.argv) > 6:
+    savefilename+= '_'+sys.argv[6]
 savefilename+='_'+str(bin)+'.sav'
 
 if os.path.exists(savefilename):
@@ -97,7 +104,7 @@ else:
 if bin != 17:
     while jj < bin*(nas-1)/nbins:
         while ii < nms:
-            boo= boost(Mh,mv[ii],a[jj],0,gamma,einasto,10**-6,0.,n,0.1,A)
+            boo= boost(Mh,mv[ii],a[jj],0,gamma,einasto,mo,0.,n,0.1,A)
             so= avg_enhance(mv[ii],m,a[jj],vdisp)
             S[ii,jj]= log10(so+boo)
             sys.stdout.write('\r'+str(ii+(jj-(bin-1)*(nas-1)/nbins)*nms+1)+'/'+str(nms*(nas-1)/nbins))
@@ -112,7 +119,7 @@ if bin != 17:
     sys.stdout.write('\n')
 else:
     while ii < nms:
-        boo= boost(Mh,mv[ii],a[jj],0,gamma,einasto,10**-6,0.,n)
+        boo= boost(Mh,mv[ii],a[jj],0,gamma,einasto,mo,0.,n)
         so= avg_enhance(mv[ii],m,a[jj],vdisp)
         S[ii,jj]= log10(so+boo)
         sys.stdout.write('\r'+str(ii+(jj-(bin-1)*(nas-1)/nbins)*nms+1)+'/'+str(nms*(nas-1)/nbins))

### 2009-02-19 py/plot_2dboostabundance.py

@@ -2,6 +2,8 @@
 #function of m_V/m and alpha
 #argument1: log_10 Mh
 #argument2: grid (default: 64)
+#argument3: DM density profile inner slope (GNFW or Einasto based on value)
+#argument4: substructure abundance power-law exponent (default -0.9)
 #First calculate this density matrix using calc_2dboost.py!!!
 
 from math import *
@@ -33,6 +35,19 @@ if len(sys.argv) > 2:
 else:
     nms= 97
     nas= 65
+if len(sys.argv) > 3:
+    gamma= double(sys.argv[3])
+    if gamma > 0.75:
+        einasto= 0
+    else:
+        einasto= 1
+else:
+    gamma= 1.
+    einasto= 0
+if len(sys.argv) > 4:
+    n= double(sys.argv[4])
+else:
+    n= -0.9
 nbins= 16#Hard-coded!!
 mv= linspace(-5,-2,nms)
 mv= 10**mv
@@ -47,7 +62,12 @@ for bb in range(nbins):
 #        continue
     #construct filenames
     bin= bb+1
-    restorefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(bin)+'.sav'
+    restorefilename= '2dboost'+str(logMh)+'_'+str(nas)
+    if len(sys.argv) > 3:
+        restorefilename+= '_'+str(gamma)
+    if len(sys.argv) > 4:
+        restorefilename+= '_'+str(n)
+    restorefilename+='_'+str(bin)+'.sav'
     restorefile=open(restorefilename,'r')
     data=pickle.load(restorefile)
     S= data['S']
@@ -62,7 +82,12 @@ for bb in range(nbins):
         ii= 0
     restorefile.close()
 #also take the last bin
-restorefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(17)+'.sav'
+restorefilename= '2dboost'+str(logMh)+'_'+str(nas)
+if len(sys.argv) > 3:
+    restorefilename+= '_'+str(gamma)
+if len(sys.argv) > 4:
+    restorefilename+= '_'+str(n)
+restorefilename+='_'+str(nbins+1)+'.sav'
 restorefile=open(restorefilename,'r')
 data=pickle.load(restorefile)
 S= data['S']
[…]

### 2009-02-19 py/plot_2dboostfree.py

@@ -2,6 +2,9 @@
 #function of m_V/m and alpha
 #argument1: log_10 Mh
 #argument2: grid (default: 64)
+#argument3: DM density profile inner slope (GNFW or Einasto based on value)
+#argument4: slope of the subhalo abundance power-law (default -0.9)
+#argument5: DM thermal free streaming limit (log10)
 #First calculate this density matrix using calc_2dboost.py!!!
 
 from math import *
@@ -33,6 +36,23 @@ if len(sys.argv) > 2:
 else:
     nms= 97
     nas= 65
+if len(sys.argv) > 3:
+    gamma= double(sys.argv[3])
+    if gamma > 0.75:
+        einasto= 0
+    else:
+        einasto= 1
+else:
+    gamma= 1.
+    einasto= 0
+if len(sys.argv) > 4:
+    n= double(sys.argv[4])
+else:
+    n= -0.9
+if len(sys.argv) > 5:
+    mo= 10**double(sys.argv[5])
+else:
+    mo= 10**-6
 nbins= 16#Hard-coded!!
 mv= linspace(-5,-2,nms)
 mv= 10**mv
@@ -47,7 +67,14 @@ for bb in range(nbins):
 #        continue
     #construct filenames
     bin= bb+1
-    restorefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(bin)+'.sav'
+    restorefilename= '2dboost'+str(logMh)+'_'+str(nas)
+    if len(sys.argv) > 3:
+        restorefilename+= '_'+str(gamma)
+    if len(sys.argv) > 4:
+        restorefilename+= '_'+str(n)
+    if len(sys.argv) > 5:
+        restorefilename+= '_'+sys.argv[5]
+    restorefilename+='_'+str(bin)+'.sav'
     restorefile=open(restorefilename,'r')
     data=pickle.load(restorefile)
     S= data['S']
@@ -62,7 +89,14 @@ for bb in range(nbins):
         ii= 0
     restorefile.close()
 #also take the last bin
-restorefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(17)+'.sav'
+restorefilename= '2dboost'+str(logMh)+'_'+str(nas)
+if len(sys.argv) > 3:
[…]

### 2009-02-19 py/plot_2dboostprofiles.py

@@ -2,6 +2,7 @@
 #function of m_V/m and alpha
 #argument1: log_10 Mh
 #argument2: grid (default: 64)
+#argument3: Einasto alpha, or gamma NFW (based on the value)
 #First calculate this density matrix using calc_2dboost.py!!!
 
 from math import *
@@ -33,6 +34,15 @@ if len(sys.argv) > 2:
 else:
     nms= 97
     nas= 65
+if len(sys.argv) > 3:
+    gamma= double(sys.argv[3])
+    if gamma > 0.75:
+        einasto= 0
+    else:
+        einasto= 1
+else:
+    gamma= 1.
+    einasto= 0
 nbins= 16#Hard-coded!!
 mv= linspace(-5,-2,nms)
 mv= 10**mv
@@ -47,7 +57,10 @@ for bb in range(nbins):
 #        continue
     #construct filenames
     bin= bb+1
-    restorefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(bin)+'.sav'
+    restorefilename= '2dboost'+str(logMh)+'_'+str(nas)
+    if len(sys.argv) > 3:
+        restorefilename+= '_'+str(gamma)
+    restorefilename+='_'+str(bin)+'.sav'
     restorefile=open(restorefilename,'r')
     data=pickle.load(restorefile)
     S= data['S']
@@ -62,7 +75,10 @@ for bb in range(nbins):
         ii= 0
     restorefile.close()
 #also take the last bin
-restorefilename= '2dboost'+str(logMh)+'_'+str(nas)+'_'+str(17)+'.sav'
+restorefilename= '2dboost'+str(logMh)+'_'+str(nas)
+if len(sys.argv) > 3:
+    restorefilename+= '_'+str(gamma)
+restorefilename+='_'+str(nbins+1)+'.sav'
 restorefile=open(restorefilename,'r')
 data=pickle.load(restorefile)
 S= data['S']
@@ -74,9 +90,12 @@ while ii < nms:
     ii= ii+1
 restorefile.close()
 
-#Then plot
-plotfilename= '2dboost'+str(logMh)+'_'+str(nas)+'.eps'
 
+#Then plot
+plotfilename= '2dboost'+str(logMh)+'_'+str(nas)
+if len(sys.argv) > 3:
+    plotfilename+= '_'+str(gamma)
[…]

### 2009-02-22 tex/somm.tex
Wrote the conclusion

@@
 in which we set the escape velocity to $v_{\mathrm{esc}} =
-\sqrt{2}\,\sigv$. Note that in the pure $1/v$ regime this averaging
-will tend to sustain the $1/v$ behavior up to a factor $\order(1)$,
-as we have that
+\sqrt{2}\,\sigv$. This approximation of a single Maxwell-Boltzmann
+distribution for the velocity distribution means that we assume that
+the halo has a constant velocity dispersion over the halo, \ie, that
+the halo is isothermal. This is not necessarily the case ---indeed, it
+is not if we believe that the density profiles of \DM\ halos are well
+fit by Einasto or \NFW\ profiles--- but it is not a bad assumption for
+intermediate distances from the halo's center, \ie, between the
+innermost and outermost parts of the halo
+\cite{2007ApJ...667L..53W}. A more detailed treatment of the
+Sommerfeld enhancement in subhalos should take the non-uniform
+velocity dispersion in subhalos into account. Note that in the pure
+$1/v$ regime this averaging will tend to sustain the $1/v$ behavior up
+to a factor $\order(1)$, as we have that
 \begin{equation}
@@
 \includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 291 430 474
-\caption{}%
+\caption{Predicted \DM\ annihilation fluxes for various dwarf
+Spheroidal satellite galaxies. The properties of the \dSphs\ used are
+given in table \ref{tab:dsphs}.}%
 \label{fig:dsph}%
@@
 
-\section{Predicted annihilation signals for dwarf Spheroidals}
+\section{Predicted annihilation signals for dwarf Spheroidals}\label{sec:dsphs}
 
@@
 \centering
-\includegraphics[clip=]{hist_-2_-3_10_1.eps}
+\includegraphics{hist_-2_-3_10_1.eps}
 \caption{The distances to and masses of subhalos within the tidal
@@
 
-\section{Discussion}
-
-Don't forget to mention the fact that the whole halo might be build up
-of distinct streams with velocity dispersion of order 1 km/s, such
-that the annihilation signal we might expect from the halo could be
-much larger than that people have suggested.
+\section{Conclusion}
+
+The recently proposed Sommerfeld enhancement to the \DM\ annihilation
+cross section forces us to reconsider statements made in the past
+about the expected signal of \DM\ anihilation from cold substructures
+in the halo of the Galaxy. Since the Sommerfeld enhancement depends
+criticically on the velocity dispersion of the substructure in
+question, as well as on the details of the particle physics model used
+to describe the attractive (or, possibly, repulsive) force that gives
+rise to the Sommerfeld enhancement, the contributions from the
+specific particle physics model and the details of the \DM\
+distribution in the substructure can no longer be as cleanly separated
+as they had been before. This means that in a scenario in which there
+is a significant Sommerfeld enhancement, previous claims of ``particle
+physics independent'' conclusions about the \DM\ annihilation signal
+from substructure (\eg,
+\cite{2007PhRvD..75h3526S,2008Natur.456...73S}) need to be
+reconsidered in this new framework. This complicates the analysis but
+it opens up new possibilities for detecting \DM\ annihilation and
+learning about the properties of the dark sector. Given a detailed
+knowledge of the density structure and velocity dispersion profile of
+the dwarf Spheroidal galaxies, much can be learned if we are able to
+detect the \DM\ annihilation signal coming from these \dSphs, by
+considering the relative signal and using the results of
+Sec.~\ref{sec:dsphs}, or by using the detailed velocity dispersion
+profile of the \dSphs\ \cite{2009arXiv0902.0362R}.
+
+We have focused on the consequences of the Sommerfeld enhancement for
+\DM\ subhalos in a \LCDM\ cosmology, using the latest results from
+high-resolution numerical $N$-body simulations to inform our
+assumptions about the abundance and properties of the
+substructure. Under these assumptions we were able to show that there
+are very large boosts to the \DM\ annihilation cross section from the
+combined effect of substructure boosts and Sommerfeld enhancement in
+large parts of the parameter space consisting of the parameters of the
+attractive Yukawa type force between the \DM\ particles which is
+responsible for the Sommerfeld enhancement. These regions of parameter
+space are in no way unrealistic. Indeed, in one of the most
+straightforward interpretations of the ATIC/PAMELA results the sharp
+cut-off around 1 TeV seen in the ATIC spectrum is interpreted as being
+due to the fact that the dark matter mass is about 700 GeV. The
+absence of an excess in anti-protons similar to the excess in
+positrons seen by PAMELA can be explained by positing that the dark
+matter particles annihilate to an intermediate boson $\phi$, which is
+lighter than a proton (or even, lighter than the effective quark mass
+$\Lambda_{\mbox{\footnotesize QCD}}$) in order to kinematically forbid
+the annihilation into protons \cite{2008arXiv0811.3641C}. This
+naturally leads to a \mv/\mdm $\lesssim$ $10^{-3}$. The coupling
+constant of the new force in the dark sector is also generally $\alpha
+\gtrsim 10^{-2}$ \cite{ArkaniHamed:2008qn,2008arXiv0812.0360L}. Thus,
+from Figure \ref{fig:subboost}, we expect there to be a rather large
+boost to the \DM\ annihilation cross section in these models. This
+leads to particularly large \DM\ annihilation signals predicted for
+the \dSphs\ (see Figure \ref{fig:dsph}).
+
+We have not discussed the substructure boost to the \DM\ annihilation
+signal from unresolved substructure in the Galactic halo. This is
+however, interesting in the light of the Sommerfeld enhancement, since
+it might lead to a different prediction about the relative magnitudes
+of the annihilation signal coming from the resolved Galactic halo and
+unresolved halo, as well from the resolved substructures in the
+Galactic halos. It is already clear that in the generic 1/$v$
+Sommerfeld enhancement the main halo, with its typical velocity
+dispersion of $\sim\!150$ km s$^{-1}$, will receive a much smaller
+enhancement than the \dSphs, with their typical velocity dispersions
+of 10 km s$^{-1}$. This, however, ignores the effects of saturation
+and resonances, which might be important in the specific particle
+physics model adopted for the Sommerfeld enhancement. Additionally,
+while the velocity dispersion of the Galactic halo might be of the
+order of 100 km s$^{-1}$ the actual velocity dispersion relevant to
+the Sommerfeld enhancement might be much smaller for halo \DM\
+particles. If a large fraction of the \DM\ in the Galactic halo is
+part of a cold stream ---and various numerical simulations predict
+that large parts of the halo, if not all of the halo, is made up of
+such cold streams with typical velocity dispersions of the order of 1
+km s$^{-1}$ \cite{1999MNRAS.307..495H,2003MNRAS.339..834H}--- then the
+relevant relative velocity of \DM\ particles is set by the velocity
+dispersion of particles in the stream, and not by the velocity
+dispersion of the streams in the halo, which is what the 100 km
+s$^{-1}$ Galactic halo velocity dispersion corresponds to in these
+scenarios. Thus, the expected \DM\ annihilation signal from the main
+halo could be much larger than that naively expected for a halo with a
+velocity dispersion of a 100 km s$^{-1}$.
 

### 2009-02-24 tex/somm.tex
Added dSph figures
Fixed various small things+figure

@@
 \newcommand{\SDSS}{SDSS}
+\newcommand{\degree}{^\circ}
 
@@
 dispersions enhance the dark matter annihilation cross section even
-further. I show that the main effect of these substructure boosts is
-that the Sommerfeld enhancement is pushed to its saturation level in
-all the interesting regions of parameters space, and that very large
-boost factors ($10^5$ -- $10^9$) are obtained in a large class of
-models. The implications of these boost factors for the detection of
-dark matter annihilation from dwarf Speroidal galaxies in the Galactic
-halo are such that many of the models proposed to explain the positron
-excess seen by the PAMELA experiment are already marginally excluded.
+further. I show that very large boost factors ($10^5$ -- $10^9$) are
+obtained in a large class of models. The implications of these boost
+factors for the detection of dark matter annihilation from dwarf
+Speroidal galaxies in the Galactic halo are such that many of the
+models proposed to explain the positron excess seen by the PAMELA
+experiment are already marginally excluded.
 \end{abstract}
@@
 \cite{ArkaniHamed:2008qn}. The main effect of this attractive force is
-to enhance the annihilation cross section with a factor proportional
-to $\beta^{-1}$ ($\beta \equiv v/c$), this is known as ``$1/v$''
-enhancement. There are regions in the parameter space made up of the
-dark matter mass \mdm, the force carrier mass \mv, and the force
-``fine-structure constant'' $\alpha \equiv$ coupling$^2$/$4 \pi$ in
-which the enhancement is near a resonance, where it approximately
+to enhance the annihilation cross section $\sigma v$ with a factor
+proportional to $\beta^{-1}$ ($\beta \equiv v/c$), this is known as
+``$1/v$'' enhancement. There are regions in the parameter space made
+up of the dark matter mass \mdm, the force carrier mass \mv, and the
+force ``fine-structure constant'' $\alpha \equiv$ coupling$^2$/$4 \pi$
+in which the enhancement is near a resonance, where it approximately
 behaves as $\beta^{-2}$. However, the enhancement levels off at a
@@
 10^{-3}$), the cross section of dark matter particles annihilating in
-the halo could be significantly enhanced by this Sommerfeld enhancement.
+the halo could be significantly enhanced by this Sommerfeld
+enhancement.
 
@@
 they orbit, they are also kinematically colder than the Galactic halo,
-with velocity dispersion typically being two orders of magnitude
+with velocity dispersions that are typically two orders of magnitude
 smaller \cite{1998ARA&A..36..435M,2007ApJ...670..313S}. This means
@@
 certain velocity dispersion $\sigv$ we need to average the Sommerfeld
-enhancement the distribution of relative velocities of that halo. We
-will approximate the one-particle velocity distribution of the halo as
-a single truncated Maxwell-Boltzmann distribution with velocity
-dispersion $\sigv$
+enhancement over the distribution of relative velocities of that
+halo. We will approximate the one-particle velocity distribution of
+the halo as a single truncated Maxwell-Boltzmann distribution with
+velocity dispersion $\sigv$
 \begin{equation}\label{eq:maxboltz}
@@
 velocity dispersion in subhalos into account. Note that in the pure
-$1/v$ regime this averaging will tend to sustain the $1/v$ behavior up
+$1/v$ regime the averaging will tend to sustain the $1/v$ behavior up
 to a factor $\order(1)$, as we have that
@@
 \int_0^\infty \dd v \, v^2\, e^{-v^2/4\sigv^2}
-\somm\left/\int_0^\infty \dd v \, v^2\, e^{-v^2/4\sigv^2} \right.\propto
+\somm(v)\left/\int_0^\infty \dd v \, v^2\, e^{-v^2/4\sigv^2} \right.\propto
 \frac{1}{\sigv}\, ,
@@
 \caption{Sommerfeld enhancement factor as a function of the particle
-physics parameters averaged over a Maxwell-Boltzman velocity
+physics parameters assuming a one-particle Maxwell-Boltzman velocity
 distribution with velocity dispersion \sigv\ for four different orders
@@
 distribution. The Sommerfeld enhancement attains very large values for
-subhalos with very small velocity dispersions, moslty, but not
+subhalos with very small velocity dispersions, mostly, but not
 exclusively, in the resonance regions. This means that the smallest
@@
 (possibly) different final states of the \DM\ annihilation, $\frac{\dd
-\Ngammai}{\dd E}$ is the spectrum from the annihilation to that state,
+\Ngammai}{\dd E}$ is the spectrum of the annihilation to that state,
 $\sigmaannv_i$ is the annihilation cross section to state $i$ (without
@@
 
-We will focus our analysis on the quantity on the ``structure quantity''
-\lum($M$), which we define here as
+We will focus our analysis on the ``structure quantity'' \lum($M$),
+which we define here as
 \begin{equation}\label{eq:structquant}
@@
 ruled out the steepest cusps, such as Moore's profile, which has an
-inner cusp with logarithmic slope of -1.5, however steeper inner cusps
-than the original \NFW\ profile are still observed in numerical
+inner cusp a with logarithmic slope of -1.5, however steeper inner
+cusps than the original \NFW\ profile are still observed in numerical
 simulations \cite{2005MNRAS.364..665D}. Other recent simulations find
 inner slopes that are shallower than an \NFW\ profile
-\cite{2008MNRAS.385..545K,2008arXiv0808.2981S,2008arXiv0810.1522N},
-although there is a growing consensus that the logarithmic slope of
-the inner density profile is a quantity that hasn't fully converged
-yet in any numerical simulation. It also has become apparent in recent
-years that the inner density profile of \DM\ halos might not be
-described by a power-law divergence, but that, on the contrary, the
-logarithmic slope becomes ever shallower, leading to a cored density
-profile
+\cite{2008MNRAS.385..545K,2008arXiv0808.2981S,2008arXiv0810.1522N}. It
+also has become apparent in recent years that the inner density
+profile of \DM\ halos might not be described by a power-law
+divergence, but that, on the contrary, the logarithmic slope becomes
+ever shallower, leading to a cored density profile
 \cite{2004MNRAS.349.1039N,2005ApJ...624L..85M,2006AJ....132.2685M,2006AJ....132.2701G,2006MNRAS.365..147S,2008arXiv0808.2981S,2008MNRAS.391.1685S,2008arXiv0810.1522N}. Therefore,
@@
 = 1 \cite{1997ApJ...490..493N}, which we will use as the fiducial
-cuspy model in what follows. Varying the value of the inner
-logarithmic slope $\gamma$ will allow us to describe the effect of
-steeper and shallower inner slopes than the \NFW\ model. Note that
-this functional form does \emph{not} include the exact form of Moore's
-profile; this need not concern us since inner cusps as steep as
-Moore's profile have been as good as excluded by the latest numerical
-simulations.
+model in what follows. Varying the value of the inner logarithmic
+slope $\gamma$ will allow us to describe the effect of steeper and
+shallower inner slopes than the \NFW\ model. Note that this functional
+form does \emph{not} include the exact form of Moore's profile; this
+need not concern us since inner cusps as steep as Moore's profile have
+been as good as excluded by the latest numerical simulations.
 
@@
 \lum\ can be simply evaluated in the context of any density model. As
-in Ref.~\cite{2007PhRvD..75h3526S} we assume that the distance
-\dist\ is much larger than the scale-radius of the \DM\ halo. For an
-\NFW\ profile 90\% of the flux originates from the region with one
-scale-radius, and we will, conservatively, only take contributions to
-the \DM\ annihilation flux coming rom this region for all the
-\GNFW\ profiles. Therefore, we find that 
+in Ref.~\cite{2007PhRvD..75h3526S} we assume that the distance to the
+subhalo \dist\ is much larger than the scale-radius of the \DM\
+halo. For an \NFW\ profile 90\% of the flux originates from the region
+within one scale-radius, and we will, conservatively, only take
+contributions to the \DM\ annihilation flux coming from this region
+for all of the different profiles. Therefore, we find that
 \begin{equation}\label{eq:lumrhosrsNFW}
@@
 Here, \rhominustwo\ is the density at the radius \rminustwo\ where the
-local slope is -2 and \alphaEinasto\ is a parameter often fixed at a
-value $\sim\!0.17$ \cite{2008MNRAS.391.1685S}. As above, assuming
+local slope is -2 and \alphaEinasto\ is a shape parameter often fixed
+at a value $\sim\!0.17$ \cite{2008MNRAS.391.1685S}. As above, assuming
 $\dist \gg \rminustwo$, calculating the flux originating from within
@@
 \ref{fig:normgammaEinasto}. In order to compare the Einasto profile as
-described here, we can relate the parameters $(\rhos,\rs)$ of the
-profile in \eqnname\ (\ref{eq:NFW}) to the parameters
+described here to the \GNFW\ profile, we can relate the parameters
+$(\rhos,\rs)$ of the \GNFW\ profile to the parameters
 $(\rhominustwo,\rminustwo)$ of the Einasto model by calculating the
@@
 A characteristic mass is given by the virial mass \Mvir. The virial
-mass of the \GNFW\ profile is given by
+mass of the \GNFW\ profile is
 \begin{equation}\label{eq:Mvirdef}
@@
 in which the function $f(c)$ is given by
-\begin{equation}
+\begin{equation}\label{eq:hypergeometric}
 f(c) = \frac{c^{3-\gamma}}{3-\gamma} \, {_2F_1}\left(3-\gamma,3-\gamma;4-\gamma;-c\right)\, .
-\end{equation}\label{eq:hypergeometric}
+\end{equation}
 The function $_2 F_1\left(a,b;c;z\right)$ is the Gauss hypergeometric
@@
 form $f(c) = \ln(1+c)-c/(1+c)$. The virial mass for the Einasto
-profile is given by
+profile is
 \begin{equation}
@@
 \end{equation}
-and \vmax\ follows then immediatly.
+and \vmax\ follows then immediately.
 
@@
 in which \lumsmoothc\ is the luminosity that would come from the
-smooth component of the halo in the absence of Sommerfeld
-enhancememnt.
+smooth component of the halo in the absence of Sommerfeld enhancement.
 
@@
 $M$, since the abundance of substructure does not continue up to this
-mass scale.
+mass scale. In the following we will set $q$ equal to 0.1.
 
@@
 With the exponent $n$ fixed, we turn to the normalization of the
-relation given in \eqnname\ (\ref{eq:massfraction}). As one might
-expect by now, different high-resolution numerical simulations give
-different answers to this question as it relates to the abundance of
-subhalos of Galactic subhalos. The problem of the normalization is
-complicated by the fact that numerical simulations have only recently
-begun to be able to resolve substructure in subhalos in a Galaxy-size
-halo. While the simulations generally agree that the resolved
-substructure abundance of subhalos in the host halo adds up to
-$\sim\!10$\% of the total mass of the halo
+relation given in \eqnname\ (\ref{eq:massfraction}). Again, different
+high-resolution numerical simulations give different answers to this
+question as it relates to the abundance of subhalos of Galactic
+subhalos. The problem of the normalization is complicated by the fact
+that numerical simulations have only recently begun to be able to
+resolve substructure in subhalos in a Galaxy-size halo. While the
+simulations generally agree that the resolved substructure abundance
+of subhalos in the host halo adds up to $\sim\!10$\% of the total mass
+of the halo
 \cite{1998MNRAS.300..146G,1999ApJ...522...82K,2004MNRAS.353..624D,2003ApJ...598...49Z,2007ApJ...657..262D,2007ApJ...667..859D},
@@
 \includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 192 277 432 487
-\caption{The total (defined in Eq.~[\ref{eq:totalboost}])and reduced
+\caption{The total (defined in Eq.~[\ref{eq:totalboost}]) and reduced
 boost (defined in Eq.~[\ref{eq:redboost}]) for subhalos of two
@@
 need to relate the velocity dispersion of a \DM\ subhalo to its
-velocity dispersion. It is important to note that all we really need
-is an approximate relationship between the velocity dispersion and the
-halo mass which fits well the range of Galactic size halos to a scale
-not much smaller than the mass scale of the smallest dwarf Spheroidals
+mass. It is important to note that all we really need is an
+approximate relationship between the velocity dispersion and the halo
+mass which fits well the range of Galactic size halos to a scale not
+much smaller than the mass scale of the smallest dwarf Spheroidals
 known today, since in most cases the Sommerfeld enhancement will
 saturate at velocity dispersions of the order of the velocity
-dispersion of the dwarf Spheroidals.
+dispersion of the dwarf Spheroidals. This means that the specific form
+that we adopt below for the velocity dispersion-mass scaling relation
+does not matter greatly.
 
@@
 \end{equation}
-from which \eqnname\ (\ref{eq:vdispscaling}) immediatly follows. This
+from which \eqnname\ (\ref{eq:vdispscaling}) immediately follows. This
 scaling relation has been found to describe numerical simulations of
 cluster-sized halos extremely well
-\cite{1991ApJ...383...95E,1995MNRAS.275..720N,1995AJ....110...21C,1996ApJ...469..494E,1996MNRAS.281..716C,1998ApJ...495...80B}.From
+\cite{1991ApJ...383...95E,1995MNRAS.275..720N,1995AJ....110...21C,1996ApJ...469..494E,1996MNRAS.281..716C,1998ApJ...495...80B}. From
 recent numerical simulations it seems that we can use this relation
@@
 By a simple estimate of the velocity dispersion using the virial
-theorem, we can use the maximum circular velocity as a measure of the
-halos velocity dispersion. This would give $\sigv \propto
+theorem, we can use the circular velocity at \rminustwo\ as a measure
+of the halos velocity dispersion. This would give $\sigv \propto
 m^{1/3.5}$. Alternatively, we could use the circular velocity at the
@@
 following the procedure outlined in the previous subsection is shown
-in Figure \ref{fig:subboost} (\emph{top panels}. It is immediately
+in Figure \ref{fig:subboost} (\emph{top panels}). It is immediately
 obvious that the total boost factors can get very large in certain
@@
 and that the boost the lower half of the parameter space, \ie, for
-smaller values of \mv/\mdm, seems to be higher than one would expect
-from the pure Sommerfeld enhancement combined with a boost factor from
-pure density enhancements.
+smaller values of \mv/\mdm, are higher than one would expect from the
+pure Sommerfeld enhancement combined with a boost factor from pure
+density enhancements.
 
@@
 has been used before to estimate the total boost from density
-enhancements in \dSphs\ and it is generally of order \order(10). This
-reduced boost therefore divides out the contribution of the Sommerfeld
+enhancements in \dSphs\ and it is generally \order(10). This reduced
+boost therefore divides out the contribution of the Sommerfeld
 enhancement of the top-level subhalo and the boost from the
@@
 give very large extra boosts. The resonance regions are clearly
-distinguishable in this plot of the reduced boost as 45$\deg$
+distinguishable in this plot of the reduced boost as 45$\degree$
 lines. Because the resonance regions are so narrow the apparent extra
@@
 clear that the inner slope does not affect the total boost due to
-substructure. The main difference between the different dark matter
-density profiles is therefore given by the normalization of the $\lum
-\propto \rhos^2 \rs^3/\dist^2$ relation, shown in Figure
+substructure, neither the overall magnitude of the effect nor the
+structure in the (\mv/\mdm,$\alpha$) plane. The main difference in
+\DM\ annihilation flux between the different dark matter density
+profiles is therefore given by the normalization of the $\lum \propto
+\rhos^2 \rs^3/\dist^2$ relation, shown in Figure
 \ref{fig:normgammaNFW}.
@@
 cored the \DM\ density profile is. Changing \alphaEinasto\ has a
-greater impact on changing the overall \DM\ density profileand
+greater impact on changing the overall \DM\ density profile and
 therefore also plays a larger role in the relation between the
@@
 \begin{figure*}[t]
-\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 291 430 474
-\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{flux_leoI_65.eps}\\%%BoundingBox: 190 291 430 474
-\includegraphics[clip=]{flux_leoII_65.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 291 430 474
+\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox:  171 290 412 488
+\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 290 433 488
+\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox:  171 290 412 488
+\includegraphics[clip=]{flux_sextans_65.eps}\\%%BoundingBox: 190 290 433 488
+\includegraphics[clip=]{flux_umaii_65.eps}%%BoundingBox:  171 290 412 488
+\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 290 433 488
 \caption{Predicted \DM\ annihilation fluxes for various dwarf
 Spheroidal satellite galaxies. The properties of the \dSphs\ used are
-given in table \ref{tab:dsphs}.}%
+given in Table \ref{tab:dsphs}.}%
 \label{fig:dsph}%
@@
 the ``classical'', or pre-\SDSS, \dSphs\
-\cite{2000PhRvD..61b3514B,2002PhRvD..66b3509T,2004PhRvD..70b3512B,2004PhRvD..69l3501E,2004PhRvD..70d3503P,2004PhRvL..93p1302H}. Especially
-the annihilation signal coming from Draco has been the subject of much
+\cite{2000PhRvD..61b3514B,2002PhRvD..66b3509T,2004PhRvD..70b3512B,2004PhRvD..69l3501E,2004PhRvD..70d3503P,2004PhRvL..93p1302H}. The
+annihilation signal coming from Draco has been the subject of much
 research in recent years because of its relative proximity to the
@@
 especially interesting, since they also have the smallest velocity
-dispersion of all the satellites of our Galaxy. Predicting the \DM\
-annihilation signal from these low-surface brightness galaxies is
-complicated by the dearth of kinematical tracers of the \DM\
-distribution and their contamination with background stars
+dispersion of all of the known satellites of our Galaxy. Predicting
+the \DM\ annihilation signal from these low-surface brightness
+galaxies is complicated by the dearth of kinematical tracers of the
+\DM\ distribution and their contamination with background stars
 \cite{2009AJ....137.3109W}. Nevertheless, detailed studies have
@@
 limit set on the \DM\ annihilation flux set by
-\cite{2008ApJ...679..428A} for all models parametrized by \mv/\mdm\
-and $\alpha$ by many orders of magnitude. We have to stress here that
-our predicted fluxes are all for the particle physics specific
-parameters given in Section \ref{sec:partphys}, \ie, \sigmaannv\ = $3
-\times 10^{-26}$ cm$^3$ s$^{-1}$, \mdm\ = 700 GeV and a particular
-form for the \DM\ annihilation spectrum. To predict the gamma-ray flux
-for a particular model with different values of these particle physics
-parameters one is required to calculate the particle physics parameter
-specific factor to the \DM\ annihilation flux and appropriately
-rescale the values given here. Nevertheless, any model that predicts
-very large annihilation cross sections to account for the PAMELA/ATIC
-excesses, and we are thinking here specifically about models that try
-to fit the PAMELA data with very large boost factors with respect to
-the fiducial annihilation cross section, give rise to these very large
-predicted gamma-ray fluxes coming from the \dSphs. Therefore, unless
-the boosts from substructure are somehow suppressed or the assumptions
-about the subhalo mass function, concentration-mass relation for
-subhalos, and/or any of the other asumptions about subhalos beyond the
-resolution limits of the state-of-the-art numerical simulations are
-violated for physical subhalos, these models are in tension with the
-experimental limits on the gamma-ray flux from \dSphs.
+Ref.~\cite{2008ApJ...679..428A} for all models parametrized by
+\mv/\mdm\ and $\alpha$ by many orders of magnitude. We have to stress
+here that our predicted fluxes are all for the particle physics
+specific parameters given in Section \ref{sec:partphys}, \ie,
+\sigmaannv\ = $3 \times 10^{-26}$ cm$^3$ s$^{-1}$, \mdm\ = 700 GeV and
+a particular form for the \DM\ annihilation spectrum. To predict the
+gamma-ray flux for a particular model with different values of these
+particle physics parameters one is required to calculate the particle
+physics parameter specific factor to the \DM\ annihilation flux and
+appropriately rescale the values given here. Nevertheless, any model
+that employs the Sommerfeld enhancement mechanism to predict the very
+large annihilation cross sections necessary to account for the
+PAMELA/ATIC excesses, and we are thinking here specifically about
+models that try to fit the PAMELA data with very large boost factors
+with respect to the fiducial annihilation cross section, give rise to
+these very large predicted gamma-ray fluxes coming from the
+\dSphs. Therefore, unless the boosts from substructure are somehow
+suppressed or the assumptions about the subhalo mass function,
+concentration-mass relation for subhalos, and/or any of the other
+asumptions about subhalos beyond the resolution limits of the
+state-of-the-art numerical simulations are violated for physical
+subhalos, these models are in tension with the experimental limits on
+the gamma-ray flux from \dSphs.
 
@@
 individual subhalos ---a fact which gives rise to the so-called
-``missing sattelites problem''
+``missing satellites problem''
 \cite{1999ApJ...522...82K,1999ApJ...524L..19M}. Each of these subhalos
@@
 simulation of a Galaxy-sized halo since redshift 104.3. It sampled the
-region of the main halo with 1,1 $\times$ 10$^9$ particles, each with
+region of the main halo with 1.1 $\times$ 10$^9$ particles, each with
 a mass of 4,100 \Msun. Over 40,000 subhalos can be resolved within 402
-kpc of the center of the main hala and 20,047 subhalos are detected
+kpc of the center of the main halo and 20,047 subhalos are detected
 which had a peak circular velocity larger than 4 km s$^{-1}$ at some
@@
 kpc. We show the distribution of these subhalos in mass and distance
-from the Sun in Figure \ref{fig:VL23_1}. The have placed the Sun at
+from the Sun in Figure \ref{fig:VL23_1}. We have placed the Sun at
 random locations at a distance of 8.5 kpc of the center of the main
-halo, corresponding to a earth-based-observer-like vantage point. The
+halo, corresponding to an earth-based-observer-like vantage point. The
 distribution of the distances in Figure \ref{fig:VL23_1} is given for
@@
 mass-velocity dispersion relation given in \eqnname\
-\ref{eq:vdispscaling} and we can calculate the characteristic density
-\rhos\ and the characteristic length-scale \rs\ from the values of
-\vmax\ and \rmax\ given for each subhalo, by the procedure outlined at
-the end of Sec.~\ref{sec:DMprofile}. For a given set of particle
-physics parameters \mv/\mdm\ and $\alpha$ this allows us to calculate
-the total boost to the \DM\ annihilation cross section for each of the
-subhalos, and by using the assumptions about the particle physics
-given in Sec.~\ref{sec:partphys} we can also calculate the photon flux
-from \DM\ annihilation coming from each subhalo, and the total
-gamma-ray flux coming from all of the subhalos combined.
-
-We will only cover some typical cases for the Sommerfeld enhancement
-here which suffice to show the general trend. We consider the particle
+(\ref{eq:vdispscaling}) and we can calculate the characteristic
+density \rhos\ and the characteristic length-scale \rs\ from the
+values of \vmax\ and \rmax\ given for each subhalo, by the procedure
+outlined at the end of Sec.~\ref{sec:DMprofile}. For a given set of
+particle physics parameters \mv/\mdm\ and $\alpha$ this allows us to
+calculate the total boost to the \DM\ annihilation cross section for
+each of the subhalos, and by using the assumptions about the particle
+physics given in Sec.~\ref{sec:partphys} we can also calculate the
+photon flux from \DM\ annihilation coming from each subhalo, and the
+total gamma-ray flux coming from all of the subhalos combined.
+
+We only cover some typical cases for the Sommerfeld enhancement here
+which suffices to show the general trend. We consider the particle
 physics parameters \mv/\mdm\ and $\alpha$ which give rise to the
@@
 magnitude from this. This is still a gamma-ray flux which is quite
-large, showing that the cumulative signal from all of the subhalos is
-dominated by the contribution from the subhalos which give rise to the
-largest gamma-ray fluxes.
+large, showing that the cumulative signal gives rise to the largest
+gamma-ray fluxes.
 
@@
 before are shown in Figures \ref{fig:VLresonance1} and
-\ref{fig:VLresonance2}. Again we find large boost factors and
+\ref{fig:VLresonance2}. Again we find large boost factors and large
 individual and total gamma-ray fluxes, although they are smaller than
@@
 distribution in the substructure can no longer be as cleanly separated
-as they had been before. This means that in a scenario in which there
+as they have been before. This means that in a scenario in which there
 is a significant Sommerfeld enhancement, previous claims of ``particle
@@
 We have not discussed the substructure boost to the \DM\ annihilation
-signal from unresolved substructure in the Galactic halo. This is
+signal from unresolved substructure in the Galactic halo. This is,
 however, interesting in the light of the Sommerfeld enhancement, since
@@
 halo could be much larger than that naively expected for a halo with a
-velocity dispersion of a 100 km s$^{-1}$.
-
-{\bf Acknowledgments:} It is a pleasure to thank Mulin Ding and David
+velocity dispersion of 100 km s$^{-1}$.
+
+{\bf Acknowledgements:} It is a pleasure to thank Mulin Ding and David
 Hogg for comments and assistance. This research was partially

### 2009-02-26 tex/somm.tex
Improvemenets after Neal's comments

@@
 v^2\, e^{-v^2/2\sigv^2}  & \qquad v \leq v_{\mathrm{esc}}\\
-0 & \qquad v > v_{\mathrm{esc}}\, ,\\
+0 & \qquad v > v_{\mathrm{esc}}\, .\\
 \end{array} \right.
 \end{equation}
-in which we set the escape velocity to $v_{\mathrm{esc}} =
-\sqrt{2}\,\sigv$. This approximation of a single Maxwell-Boltzmann
+This approximation of a single Maxwell-Boltzmann
 distribution for the velocity distribution means that we assume that
@@
 \begin{figure*}[t]
-\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox:  171 290 412 488
-\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 290 433 488
-\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox:  171 290 412 488
-\includegraphics[clip=]{flux_sextans_65.eps}\\%%BoundingBox: 190 290 433 488
-\includegraphics[clip=]{flux_umaii_65.eps}%%BoundingBox:  171 290 412 488
-\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 290 433 488
+\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox:  171 290 412 500
+\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 290 433 500
+\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox:  171 290 412 500
+\includegraphics[clip=]{flux_sextans_65.eps}\\%%BoundingBox: 190 290 433 500
+\includegraphics[clip=]{flux_umaii_65.eps}%%BoundingBox:  171 290 412 500
+\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 290 433 500
 \caption{Predicted \DM\ annihilation fluxes for various dwarf
@@
 
-{\bf Acknowledgements:} It is a pleasure to thank Mulin Ding and David
-Hogg for comments and assistance. This research was partially
+{\bf Acknowledgements:} It is a pleasure to thank Mulin Ding, David
+Hogg, and Neal Weiner for comments and assistance. This research was partially
 supported by NASA (grant NNX08AJ48G).

### 2009-02-26 py/2denhance.py
Changed escape velocity + VL + dsphs

@@ -80,7 +80,7 @@ im= imshow(data['S'],origin='lower',#cmap=cm.gray,
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)
 xlabel(r'$\log_{10}\alpha$')#,fontsize=16)
-text(-1.75,-4.75,r'$\beta = 10^{-3}$')#,fontsize=16)
+text(-1.75,-4.75,r'$\beta = 10^{-4}$')#,fontsize=16)
 CB1= colorbar(im,shrink=0.87)
 levels=linspace(0,5,11)
 #cont= contour(data['S'],levels,origin='lower',linewidths=1,colors='k',

### 2009-02-26 py/enhance.py

@@ -69,7 +69,7 @@ def avg_enhance(mv,m,a,b,nsamples=100,vesc=0):
     norm= 0.
 
     if vesc == 0.:
-        vesc= 2.*sqrt(2.)*b
+        vesc= 8.*b#2.*sqrt(2.)*b
 
     bs= linspace(vesc/(nsamples-1.),vesc,nsamples-1)
 

### 2009-02-26 py/hist_VL.py

@@ -144,6 +144,48 @@ else:
 for ii in range(sublum.shape[0]):
     sublum[ii,9]= log10(sublum[ii,9])
 
+
+
+
+
+
+savefilename2= 'hist_'+sys.argv[1]+'_'+sys.argv[2]+'_'+sys.argv[3]+'_2.sav'
+
+if os.path.exists(savefilename2):
+    savefile2=open(savefilename2,'r+')
+    data=pickle.load(savefile2)
+    totallumnosomm=data['totallumnosomm']
+else:
+    #calculate the signal for the 1,000 observers without Sommerfeld enhancement
+    nobs2= 1001
+    totallumnosomm= zeros(nobs2)
+    for ii in range(nobs2):
+        #Generate random observer
+        phi= numpy.random.uniform(0.,2.*pi)
+        theta= numpy.random.uniform(0.,pi)
+        pos= array([obsR*sin(theta)*sin(phi), obsR*sin(theta)*cos(phi), obsR*cos(theta)])
+        for jj in range(sublum.shape[0]):
+            dist2= pow((sublum[jj,0]-pos[0]),2.)+pow((sublum[jj,1]-pos[1]),2.)+pow((sublum[jj,2]-pos[2]),2.)
+            totallumnosomm[ii]+= sublum[jj,3]*(1.+sublum[jj,6])/dist2
+        totallumnosomm[ii]= log10(totallumnosomm[ii]) +log10(0.5/17.)
+    #Save
+    savefile2=open(savefilename2,'w')
+    data={'totallumnosomm':totallumnosomm}
+    pickle.dump(data,savefile2)
+
+print numpy.sum(totallumnosomm)/numpy.sum(totallum)
+
+
+#Post-process the fluxes to incorporate the new spectrum
+
+#First do totallum
+for ii in range(totallum.shape[0]):
+    totallum[ii]= totallum[ii]+log10(0.5/17.)
+#Then do the sublumobs
+for ii in range(sublumobs.shape[0]):
+    for jj in range(sublumobs.shape[1]):
+        sublumobs[ii,jj,1]= sublumobs[ii,jj,1]+log10(0.5/30)
+
     
[…]

### 2009-02-26 py/plot_dsph.py

@@ -11,6 +11,7 @@ import matplotlib
 matplotlib.use('PS')
 from pylab import *
 from matplotlib.pyplot import *
+from matplotlib.pyplot import title as pytitle
 from enhance import avg_enhance
 from boost import boost
 import os, sys
@@ -22,24 +23,74 @@ from matplotlib import rc
 #parameters
 if len(sys.argv) > 1:
     if sys.argv[1] == "fornax":
+        rs= 1.85
+        rhos= 10**7
+        M= 1.54*10**8
+        D= 138.
+        vdisp= 10.5
         title="Fornax"
     elif sys.argv[1] == "leoI":
+        rs= 2.77
+        rhos= 10**7
+        M= 5.16*10**8
+        D= 250.
+        vdisp= 8.8
         title="Leo I"
     elif sys.argv[1] == "sculptor":
+        rs= 2.77
+        rhos= 10**7
+        M= 5.16*10**8
+        D= 79.
+        vdisp= 6.6
         title="Sculptor"
     elif sys.argv[1] == "leoII":
+        rs= 1.85
+        rhos= 10**7
+        M= 1.54*10**8
+        D= 205.
+        vdisp= 6.7
         title="Leo II"
     elif sys.argv[1] == "sextans":
+        rs= 0.92
+        rhos= 10**7
+        M= 1.9*10**7
+        D= 86.
+        vdisp= 6.6
         title="Sextans"
     elif sys.argv[1] == "carina":
+        rs= 1.39
+        rhos= 1.13*10**7
+        M= 7.37*10**7
+        D= 101.
+        vdisp= 6.8
         title="Carina"
     elif sys.argv[1] == "ursaminor":
+        rs= 2.77
+        rhos= 10**7
+        M= 5.15*10**8
+        D= 66.
+        vdisp= 9.3
         title="Ursa Minor"
     elif sys.argv[1] == "draco":
+        rs= 3.7
+        rhos= 10**7
+        M= 1.23*10**8
+        D= 82.
+        vdisp= 9.5
         title="Draco"
     elif sys.argv[1] == "coma":
+        rs= 0.3
+        rhos= 3.*10**8
+        M= 1.97*10**8
+        D= 44.
+        vdisp= 4.6
         title="Coma Berenices"
[…]

### 2009-02-26 tex/somm.tex
Various Weiner sggestions incorporated

@@
 factors for the detection of dark matter annihilation from dwarf
-Speroidal galaxies in the Galactic halo are such that many of the
-models proposed to explain the positron excess seen by the PAMELA
-experiment are already marginally excluded.
+Speroidal galaxies in the Galactic halo are such that they outshine
+the extragalactic background gamma-ray flux and should be detectable
+by \Fermi.
 \end{abstract}
@@
 matter annihilation signal in these already cold subhalos can be
-boosted even further by contribution of their substructure. In this
-paper we quantify the magnitude of this new substructure boost and
-discuss its implications for the detection of dark matter annihilation
-from subhalos in the Galactic halos.
+boosted even further by contribution of their substructure.
+
+In this paper we quantify the magnitude of this new substructure boost
+and discuss its implications for the detection of dark matter
+annihilation from subhalos in the Galactic halos. As we will show, the
+total combined boost from Sommerfeld enhancement and substructure can
+reach very large values (10$^5$ -- 10$^9$), and these boosts are only
+weakly dependent on the assumptions that we have to make about the
+abundance of substructure and the \DM\ density profile. We use these
+total boosts together with published \DM\ density profiles of some of
+the dwarf Spheroidal satellite galaxies of the Milky Way to predict
+the total flux of gamma-rays from \DM\ annihilation that would be
+detectable by Fermi and find that in most of the models that posit the
+Sommerfeld enhancement the \DM\ annihilation signal from the dwarf
+Spheroidals shines above the diffuse extragalactic gamma-ray
+background, which is the only background signal at the high Galactic
+lattitudes of these satellite galaxies. Finally, we also consider the
+prospects for detecting smaller subhalos of the Galaxy by computing
+annihilation fluxes of subhalos in the \VL\-2 simulation. We will see
+that whether these subhalos will be detectable in the near future is
+highly dependent on the exact details of the Sommerfeld enhancement.
 
@@
 
-\subsection{Some details of the particle physics that do not particularly concern us here}\label{sec:partphys}
+\subsection{Some details of the particle physics}\label{sec:partphys}
 
@@
 Sommerfeld enhancement we will adopt the fiducial value $\sigmaannv =
-3 \times 10^{-26} $ cm$^3$ s$^{-1}$. The energy spectrum of the \DM\
+3 \times 10^{-26} $ cm$^3$ s$^{-1}$, and we choose a fiducial mass for
+the \DM\ particle of \mdm\ = 700 GeV. The energy spectrum of the \DM\
 annihilation depends on the model used to describe the \DM\ and its
@@
 detailed treatment of this for different \DM\ models. Therefore, we
-use an annihilation spectrum appropriate for neutralino dark matter,
-which provides a reasonable fit to a variety of different annihilation
-products. Since the energy spectrum contributes a multiplicative
-factor to the \DM\ annihilation flux, the effect of different
-annihilation spectra can be incorporated by simply multiplying our
-results with the appropriate factor. The neutralino dark matter
-annihilation spectrum we use has the form
-\cite{1998APh.....9..137B,2004PhRvD..70j3529F}
+will adopt an annihilation spectrum for \DM\ annihilation into photons
+appropriate for the newer models that were proposed to explain the
+PAMELA/ATIC results. Specifically, we use a gamma ray spectrum from a
+cascade annihilation model. In these models the \DM\ annihilates into
+a new light degree of freedom which then in turn decays into leptons
+and photons. The injection spectrum per \DM\ annihilation is given by
+\cite{2009arXiv0901.2926M}
 \begin{equation}\label{eq:annspectrum}
-\frac{\dd \Ngamma}{\dd E} = \frac{a_1}{\mdm} \left(\frac{E}{\mdm}\right)^{-3/2}\, \exp\left[-a_2\frac{E}{\mdm}\right]\, ,
-\end{equation}
-with parameters $a_1\sim 1$ and $a_2 \sim 10$, depending on the
-details of the annihilation process
-\cite{2004PhRvD..70j3529F}. Integrating this spectrum using a
-threshold energy appropriate for \Fermi\, $\Eth = 4$ GeV, and choosing
-a \DM\ mass, $\mdm = 700$ GeV, we find that
-\begin{equation}
-\int_{\Eth}^{\mdm}\dd E\, \sum_i \frac{\dd
-\Ngammai}{\dd E} \approx 17\, ,
-\end{equation}
-with any differences due to variations in $(a_1,a_2)$ and $\mdm$
-between different models leading to changes of a factor $\order(10)$.
+\begin{split}
+\frac{\dd \Ngamma}{\dd E} = &\frac{\alpha_{\mbox{{\footnotesize EM}}}}{\pi} \frac{1+(1-E/\mdm)^2}{E}\\
+&\times\left\{-1 + \ln\left(4(1-E/\mdm)\right) -2\ln\left(\frac{m_l}{\mdm}\right)\right\}\, ,
+\end{split}
+\end{equation}
+in which $\alpha_{\mbox{{\footnotesize EM}}}$ is the electromagnetic
+coupling constant and $m_l$ is the mass of the leptons created in the
+process $\chi\chi \rightarrow l^+l^-\gamma$. Integrating this spectrum
+using an energy range appropriate for \Fermi\ we find that
+\begin{equation} 
+\int_{4 \mbox{{\footnotesize GeV}}}^{250 \mbox{{\footnotesize GeV}}}\dd E\, \frac{\dd
+\Ngamma}{\dd E} \approx 0.5\, .
+\end{equation}
+We find a similar number if we integrate the spectrum due to one
+cascade step. It is interesting to note that these photon spectra have
+fewer photons in them, about an order of magnitude less, than the
+annihilation spectra due to neutralino annihilation which are often
+used \cite{1998APh.....9..137B,2004PhRvD..70j3529F}. Of course, since
+the energy spectrum contributes a multiplicative factor to the \DM\
+annihilation flux, the effect of different annihilation spectra can be
+incorporated by simply multiplying our results with the appropriate
+factor.
 
@@
 We can now write
-\begin{equation}
+\begin{equation}\label{eq:totboost}
 \begin{split}
@@
 
-\begin{figure*}[t]
-\includegraphics[clip=]{2dboost8_65.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{2dboost7_65.eps}\\%%BoundingBox: 190 291 430 474 
-\includegraphics[clip=]{2dredboost8_65.eps}%%BoundingBox: 170 277 398 487 
-\includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 192 277 432 487
-\caption{The total (defined in Eq.~[\ref{eq:totalboost}]) and reduced
-boost (defined in Eq.~[\ref{eq:redboost}]) for subhalos of two
-different mass scales. The dark matter profile of all the subhalos is
-assumed to be well-described by an \NFW\ profile.}%
-\label{fig:subboost}%
-\end{figure*}
-
 In the previous section we calculated $\lumsmoothc$ for the different
@@
 
-On the scale of clusters there is a well-defined relationship between
-the mass of a \DM\ halo and its one-dimensional velocity dispersion
-which one can derive from adiabatic scaling arguments
-\cite{1998ApJ...495...80B} given by
-\begin{equation}\label{eq:vdispscaling}
-\sigma \propto m^{1/3}\, .
-\end{equation}
-This scaling relation can be easily derived in the case of an
-isothermal distribution function \cite{2008gady.book.....B}, since
-then
-\begin{equation}
-\rho(r) = \frac{\sigma^2}{2\pi G r^2}\, ,
-\end{equation}
-from which \eqnname\ (\ref{eq:vdispscaling}) immediately follows. This
-scaling relation has been found to describe numerical simulations of
-cluster-sized halos extremely well
-\cite{1991ApJ...383...95E,1995MNRAS.275..720N,1995AJ....110...21C,1996ApJ...469..494E,1996MNRAS.281..716C,1998ApJ...495...80B}. From
-recent numerical simulations it seems that we can use this relation
-even for subhalos of subhalos in a Galaxy sized dark matter halo. In
-Ref.~\cite{2008MNRAS.391.1685S} it is shown that the relation between
-the maximum circular velocity of a subhalo and its mass is well fit by
-a power-law of the type
-\begin{equation}
-m \propto \vmax^{3.5}\, .
-\end{equation}
-By a simple estimate of the velocity dispersion using the virial
-theorem, we can use the circular velocity at \rminustwo\ as a measure
-of the halos velocity dispersion. This would give $\sigv \propto
-m^{1/3.5}$. Alternatively, we could use the circular velocity at the
-virial radius as a measure of the velocity dispersion. In the case of
-an \NFW\ profile we have that
-\begin{equation}
-\frac{\vmax^2}{\vcirc(\Rvir)^2} \propto \frac{c}{f(c)}\, ,
-\end{equation}
-in which $f(x) = \ln(1+c) - c/(1+c) \approx 2.6 (c/33)^{0.4}$ in the
-relevant concentration range. Using \eqnname\ (\ref{eq:concscaling}),
-this gives
-\begin{equation}
-\sigv \propto M^{0.32}\, .
-\end{equation}
-Therefore, we will use (as a rough approximation) the scaling relation
-given in \eqnname\ (\ref{eq:vdispscaling}). We calibrate this
-relationship at the mass scale of the dwarf spheroidals by using the
-velocity and mass of Canes Venatici I from Ref.~\cite{2007ApJ...670..313S}.
-
-\subsection{The total and the ``reduced'' boost}
+\begin{figure*}[t]
+\includegraphics[clip=]{2dboost8_65.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost7_65.eps}\\%%BoundingBox: 190 291 430 474 
+\includegraphics[clip=]{2dredboost8_65.eps}%%BoundingBox: 170 277 398 487 
+\includegraphics[clip=]{2dredboost7_65.eps}%%BoundingBox: 192 277 432 487
+\caption{The total (defined in Eq.~[\ref{eq:totalboost}]) and reduced
+boost (defined in Eq.~[\ref{eq:redboost}]) for subhalos of two
+different mass scales. The dark matter profile of all the subhalos is
+assumed to be well-described by an \NFW\ profile.}%
+\label{fig:subboost}%
+\end{figure*}
 
@@
 
-We define the total boost for a given subhalo mass as
+\begin{figure*}[t]
+\includegraphics[clip=]{2dboost8_65_-0.7.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost8_65_-1.0.eps}\\%%BoundingBox: 190 291 430 474
+\includegraphics[clip=]{2dboost8_65_-3.eps}%%BoundingBox: 178 291 400 474
+\includegraphics[clip=]{2dboost8_65_-12.eps}%%BoundingBox: 190 291 430 474
+\caption{The total boost for a $M=10^8$ \Msun\ for different slopes
+$n$ for the subhalo mass function (\emph{top panels}) and for
+different values of the dark matter thermal free streaming limit
+$m_0$ (\emph{bottom panels}).}%
+\label{fig:subboostslope}%
+\end{figure*}
+
+On the scale of clusters there is a well-defined relationship between
+the mass of a \DM\ halo and its one-dimensional velocity dispersion
+which one can derive from adiabatic scaling arguments
+\cite{1998ApJ...495...80B} given by
+\begin{equation}\label{eq:vdispscaling}
+\sigma \propto m^{1/3}\, .
+\end{equation}
+This scaling relation can be easily derived in the case of an
+isothermal distribution function \cite{2008gady.book.....B}, since
+then
+\begin{equation}
+\rho(r) = \frac{\sigma^2}{2\pi G r^2}\, ,
+\end{equation}
+from which \eqnname\ (\ref{eq:vdispscaling}) immediately follows. This
+scaling relation has been found to describe numerical simulations of
+cluster-sized halos extremely well
+\cite{1991ApJ...383...95E,1995MNRAS.275..720N,1995AJ....110...21C,1996ApJ...469..494E,1996MNRAS.281..716C,1998ApJ...495...80B}. From
+recent numerical simulations it seems that we can use this relation
+even for subhalos of subhalos in a Galaxy sized dark matter halo. In
+Ref.~\cite{2008MNRAS.391.1685S} it is shown that the relation between
+the maximum circular velocity of a subhalo and its mass is well fit by
+a power-law of the type
+\begin{equation}
+m \propto \vmax^{3.5}\, .
+\end{equation}
+By a simple estimate of the velocity dispersion using the virial
+theorem, we can use the circular velocity at \rminustwo\ as a measure
+of the halos velocity dispersion. This would give $\sigv \propto
+m^{1/3.5}$. Alternatively, we could use the circular velocity at the
+virial radius as a measure of the velocity dispersion. In the case of
+an \NFW\ profile we have that
+\begin{equation}
+\frac{\vmax^2}{\vcirc(\Rvir)^2} \propto \frac{c}{f(c)}\, ,
+\end{equation}
+in which $f(x) = \ln(1+c) - c/(1+c) \approx 2.6 (c/33)^{0.4}$ in the
+relevant concentration range. Using \eqnname\ (\ref{eq:concscaling}),
+this gives
+\begin{equation}
+\sigv \propto M^{0.32}\, .
+\end{equation}
+Therefore, we will use (as a rough approximation) the scaling relation
+given in \eqnname\ (\ref{eq:vdispscaling}). We calibrate this
+relationship at the mass scale of the dwarf spheroidals by using the
+velocity and mass of Canes Venatici I from Ref.~\cite{2007ApJ...670..313S}.
+
+\subsection{The total and the ``reduced'' boost}
+
+We define the total boost for a given subhalo mass as (see \eqnname\
+[\ref{eq:totboost}])
 \begin{equation}\label{eq:totalboost}
@@
 
-\begin{figure*}[t]
-\includegraphics[clip=]{2dboost8_65_-0.7.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{2dboost8_65_-1.0.eps}\\%%BoundingBox: 190 291 430 474
-\includegraphics[clip=]{2dboost8_65_-3.eps}%%BoundingBox: 178 291 400 474
-\includegraphics[clip=]{2dboost8_65_-12.eps}%%BoundingBox: 190 291 430 474
-\caption{The total boost for a $M=10^8$ \Msun\ for different slopes
-$n$ for the subhalo mass function (\emph{top panels}) and for
-different values of the dark matter thermal free streaming limit
-$m_0$ (\emph{bottom panels}).}%
-\label{fig:subboostslope}%
-\end{figure*}
-
 However, the Sommerfeld enhancement has not saturated in all regions
@@
 the higher saturation level for smaller $\mv/\mdm$ values.
+
+The total boost factors obtained by assuming an \NFW\ dark matter
+density profile do not depend much on the inner slope of the density
+profile. This is mainly because the calculation of the boost only
+considers \emph{relative} luminosities, for which the difference in
+normalization of the $\lum \propto \rhos^2 \rs^3/\dist^2$ relation for
+different inner slopes is irrelevant, and because the mass of the
+subhalo is only weakly affected by the value of the inner slope ---the
+function $f(c)$ is only a weak function of the inner slope. This is
+confirmed by calculating the total boost for a $M = 10^8$ \Msun\
+subhalo for a subhalo with a very shallow inner cusp as well as for a
+subhalo with a very steep inner slope. These are shown in Figure
+\ref{fig:subboostprofiles} (\emph{top panels}), and, from comparing
+this with the total boost shown in Figure \ref{fig:subboost}, it is
+clear that the inner slope does not affect the total boost due to
+substructure, neither the overall magnitude of the effect nor the
+structure in the (\mv/\mdm,$\alpha$) plane. The main difference in
+\DM\ annihilation flux between the different dark matter density
+profiles is therefore given by the normalization of the $\lum \propto
+\rhos^2 \rs^3/\dist^2$ relation, shown in Figure
+\ref{fig:normgammaNFW}.
+
+If the \DM\ density profiles of halos is better described by an
+Einasto profile then the substructure boost is affected more strongly
+by the shape parameter \alphaEinasto\, which basically describes how
+cored the \DM\ density profile is. Changing \alphaEinasto\ has a
+greater impact on changing the overall \DM\ density profile and
+therefore also plays a larger role in the relation between the
+characteristic density, the concentration, and the mass of the
+subhalo. The total boosts for a $M$ = $10^8$ \Msun\ subhalo is shown
+in Figure \ref{fig:subboostprofiles} (\emph{bottom panels}) for two
+different values of the shape parameter \alphaEinasto\, which span the
+range found in numerical simulations (for values of \alphaEinasto\
+greater than 0.2 the behavior basically saturates at that of
+\alphaEinasto\ = 0.2). For \alphaEinasto\ = 0.1 the total boost is
+about the same as for the \GNFW\ profile shape, which is
+understandable given that the Einasto profile is more cuspy for
+smaller values of \alphaEinasto. The same dependence on the parameter
+\mv/\mdm\ as is seen for the \GNFW\ profile is displayed for the
+Einasto profile. In the case that \alphaEinasto\ is larger the
+substructure boost reaches slightly larger values than for the \GNFW\
+density profile. Thus, as was the case for the \GNFW\ profile, the
+impact of the different normalization of the
+$\rhominustwo^2\rminustwo^3/\dist^2$ for different values of
+\alphaEinasto\ will play a leading role in the difference between the
+different profile shapes.
 
@@
 
-The total boost factors obtained by assuming an \NFW\ dark matter
-density profile do not depend much on the inner slope of the density
-profile. This is mainly because the calculation of the boost only
-considers \emph{relative} luminosities, for which the difference in
-normalization of the $\lum \propto \rhos^2 \rs^3/\dist^2$ relation for
-different inner slopes is irrelevant, and because the mass of the
-subhalo is only weakly affected by the value of the inner slope ---the
-function $f(c)$ is only a weak function of the inner slope. This is
-confirmed by calculating the total boost for a $M = 10^8$ \Msun\
-subhalo for a subhalo with a very shallow inner cusp as well as for a
-subhalo with a very steep inner slope. These are shown in Figure
-\ref{fig:subboostprofiles} (\emph{top panels}), and, from comparing
-this with the total boost shown in Figure \ref{fig:subboost}, it is
-clear that the inner slope does not affect the total boost due to
-substructure, neither the overall magnitude of the effect nor the
-structure in the (\mv/\mdm,$\alpha$) plane. The main difference in
-\DM\ annihilation flux between the different dark matter density
-profiles is therefore given by the normalization of the $\lum \propto
-\rhos^2 \rs^3/\dist^2$ relation, shown in Figure
-\ref{fig:normgammaNFW}.
-
-If the \DM\ density profiles of halos is better described by an
-Einasto profile then the substructure boost is affected more strongly
-by the shape parameter \alphaEinasto\, which basically describes how
-cored the \DM\ density profile is. Changing \alphaEinasto\ has a
-greater impact on changing the overall \DM\ density profile and
-therefore also plays a larger role in the relation between the
-characteristic density, the concentration, and the mass of the
-subhalo. The total boosts for a $M$ = $10^8$ \Msun\ subhalo is shown
-in Figure \ref{fig:subboostprofiles} (\emph{bottom panels}) for two
-different values of the shape parameter \alphaEinasto\, which span the
-range found in numerical simulations (for values of \alphaEinasto\
-greater than 0.2 the behavior basically saturates at that of
-\alphaEinasto\ = 0.2). For \alphaEinasto\ = 0.1 the total boost is
-about the same as for the \GNFW\ profile shape, which is
-understandable given that the Einasto profile is more cuspy for
-smaller values of \alphaEinasto. The same dependence on the parameter
-\mv/\mdm\ as is seen for the \GNFW\ profile is displayed for the
-Einasto profile. In the case that \alphaEinasto\ is larger the
-substructure boost reaches slightly larger values than for the \GNFW\
-density profile. Thus, as was the case for the \GNFW\ profile, the
-impact of the different normalization of the
-$\rhominustwo^2\rminustwo^3/\dist^2$ for different values of
-\alphaEinasto\ will play a leading role in the difference between the
-different profile shapes.
-
 \begin{figure*}[t]
@@
 \caption{Predicted \DM\ annihilation fluxes for various dwarf
-Spheroidal satellite galaxies. The properties of the \dSphs\ used are
-given in Table \ref{tab:dsphs}.}%
+Spheroidal satellite galaxies. The background flux of extragalactic
+photons in the relevant energy range is given under the name of the
+dwarf Spheroidal. The properties of the \dSphs\ used are given in
+Table \ref{tab:dsphs}.}%
 \label{fig:dsph}%
@@
 been found to be rather poor, with detection with \Fermi\ being
-unlikely \cite{2008arXiv0812.1494P}. Some of the \dSphs\ have already
-been the subject of observational searches for \DM\ annihilation and
-some limits on the \DM\ annihilation signal have already been set
+unlikely \cite{2008arXiv0812.1494P,2009arXiv0902.4330P}. Some of the
+\dSphs\ have already been the subject of observational searches for
+\DM\ annihilation and some limits on the \DM\ annihilation signal have
+already been set
 \cite{2008ApJ...678..594W,2008ApJ...679..428A,2008APh....29...55A,2008PhRvD..78h7101D,2009ApJ...691..175A,2008arXiv0810.3561M}.
@@
 predicted to be coming from the \dSphs\ is very large in a large class
-of models. For example, the flux predicted for Draco exceeds the upper
-limit set on the \DM\ annihilation flux set by
-Ref.~\cite{2008ApJ...679..428A} for all models parametrized by
-\mv/\mdm\ and $\alpha$ by many orders of magnitude. We have to stress
-here that our predicted fluxes are all for the particle physics
-specific parameters given in Section \ref{sec:partphys}, \ie,
-\sigmaannv\ = $3 \times 10^{-26}$ cm$^3$ s$^{-1}$, \mdm\ = 700 GeV and
-a particular form for the \DM\ annihilation spectrum. To predict the
-gamma-ray flux for a particular model with different values of these
-particle physics parameters one is required to calculate the particle
-physics parameter specific factor to the \DM\ annihilation flux and
-appropriately rescale the values given here. Nevertheless, any model
-that employs the Sommerfeld enhancement mechanism to predict the very
-large annihilation cross sections necessary to account for the
-PAMELA/ATIC excesses, and we are thinking here specifically about
-models that try to fit the PAMELA data with very large boost factors
-with respect to the fiducial annihilation cross section, give rise to
-these very large predicted gamma-ray fluxes coming from the
-\dSphs. Therefore, unless the boosts from substructure are somehow
-suppressed or the assumptions about the subhalo mass function,
-concentration-mass relation for subhalos, and/or any of the other
-asumptions about subhalos beyond the resolution limits of the
-state-of-the-art numerical simulations are violated for physical
-subhalos, these models are in tension with the experimental limits on
-the gamma-ray flux from \dSphs.
-
-%dSphs good candidates for \DM\ annihilation:
-%2000PhRvD..61b3514B,%Baltz/SUSY 2002PhRvD..66b3509T,%Tyler,Draco
-%2004PhRvD..70b3512B,%Baltz,general 
-%2004PhRvD..69l3501E,%Evans,general
-%2004PhRvD..70d3503P,%Peirani, Draco, Sagit, M31, M87.
-%2004PhRvL..93p1302H,%Hooper: Sagit: 511 keV
-%2006PhRvD..73f3510B,%Bergstrom,Draco 
-%2006JCAP...03..003P,%Profumo,Draco, SUSY 
-%2007PhRvD..75b3513C,%Colafrancesco,Draco
-%2007PhRvD..76l3509S,%Sanchez-Conde,Draco
-%2007PhRvD..75h3526S,%StrigariI 
-%2008ApJ...678..614S,%StrigariII
-%2008arXiv0809.2781G,%Geha 
-%2008arXiv0812.1494P,%Pieri, no Draco, Umi,Carina or Sextans with Fermi
-
-%Limits:
-%2008ApJ...678..594W,%Wood: Whipple: Draco and Umi
-%2008ApJ...679..428A,%Albert et al, MAGIC: <10^-11
-%2008APh....29...55A,%HESS: Sagg: 
-%2008PhRvD..78h7101D,%Driscoll: Draco with STACEE: limit on the spectrum
-%2009ApJ...691..175A,%Hess: Canis Major
-%2008arXiv0810.3561M,%MAGIC: Willman 1 < 10^-12
-
-%dSph modeling:
-%2007PhRvD..75h3526S,%StrigariI
-%2007ApJ...667L..53W,%Walker: vdisp profiles
-%2008ApJ...672..904P,%Penarubia+CDM
-%2009AJ....137.3109W,%Walker, EM contamination
+of models. We have to stress here that our predicted fluxes are all
+for the particle physics specific parameters given in Section
+\ref{sec:partphys}, \ie, \sigmaannv\ = $3 \times 10^{-26}$ cm$^3$
+s$^{-1}$, \mdm\ = 700 GeV and a particular form for the \DM\
+annihilation spectrum. To predict the gamma-ray flux for a particular
+model with different values of these particle physics parameters one
+is required to calculate the particle physics parameter specific
+factor to the \DM\ annihilation flux and appropriately rescale the
+values given here. In order to compare these signals to the expected
+background flux of extragalactic gamma rays (which is the only
+relevant background since these \dSphs\ are at high Galactic
+lattitude), we have integrated the extragalactic gamma ray flux as
+measured by EGRET over the energy range between 4 GeV -- 250 GeV and
+multiplied it with the solid angle that the \dSphs\ span on the
+sky. The extragalactic gamma ray flux is given by a power law
+\cite{1998ApJ...494..523S}
+\begin{equation}
+\frac{\dd \Ngamma}{\dd E \,\dd A \, \dd t\,\dd \Omega} = k\,\left(\frac{E}{E_0}\right)^{-\alpha}\, ,
+\end{equation}
+where $k = 7.3 \times 10^{-9}$, $\alpha$ = 2.10, and $E_0 = 0.45$
+GeV. We have included the background flux of extragalactic photons in
+Figure \ref{fig:dsph}. This background flux is in all cases of the
+order of 10$^{-13}$ photons cm$^{-2}$ s$^{-1}$, such that the \DM\
+annihilation boosted by substructure and Sommerfeld enhancement
+outshine this background in all of the considered models, and by many
+orders of magnitude in most models. Given that the flux above 100 MeV
+over a 55 days time-span needed to separate a point source of \DM\
+annihilation from the background by \Fermi\ to give a 5$\sigma$
+detection is about 10$^{-8}$ photons cm$^{-1}$ s$^{-1}$ and that the
+necessary fluxes scale roughly as $t_{\mbox{{\footnotesize
+exp}}}^{-1/2}$ \cite{2008JCAP...07..013B}, the \DM\ annihilation of
+these \dSphs\ should be detectable by \Fermi\ in many of the models
+parameterized by ($\mv/\mdm,\alpha$). We conclude that the prospects
+for detecting \DM\ annihilation from the \dSphs\ with \Fermi\ are very
+good in models that include a significant Sommerfeld enhancement.
 
@@
 \includegraphics{hist_-2_-3_10_3.eps}
-\caption{Total gamma-ray flux from subhalos in \VL\ II for 1,000 random
-observers at 8.5 kpc from the center of the main \VL\ halo for
-\mv/\mdm\ = 10$^{-2}$ and $\alpha = 10^{-3}$.}%\
+\caption{Total gamma-ray flux from subhalos in \VL\ II for 1,000
+random observers at 8.5 kpc from the center of the main \VL\ halo for
+\mv/\mdm\ = 10$^{-2}$ and $\alpha = 10^{-3}$. For comparison we show
+the total gamma-ray flux from subhalos in the absence of Sommerfeld
+enhancement as the dashed histogram and the top horizontal axis.}%\
 \label{fig:VL23_3}%
@@
 small in this case, as is shown in Figure \ref{fig:VL23_3}, in which
-the total gamma-ray flux is shown for 1,000 random observer at a
+the total gamma-ray flux is shown for 1,000 random observers at a
 distance of 8.5 kpc from the galactic center. At most observation
@@
 high-end tail not extending further than about half an order of
-magnitude from this. This is still a gamma-ray flux which is quite
-large, showing that the cumulative signal gives rise to the largest
-gamma-ray fluxes.
+magnitude from this. For comparison the dashed histogram in Figure
+\ref{fig:VL23_3} shows the same distribution of total gamma-ray flux
+for different random observers in the absence of any Sommerfeld
+enhancement. This confirms that the total gamma-ray flux with
+Sommerfeld enhancement is about an order of magnitude larger than
+without Sommerfeld enhancement.
 
@@
 predicted flux in this case is many orders of magnitude larger than
-the total flux in the other extremum of the Sommerfeld
-enhancement. Thus, the exact value of the parameters \mv/\mdm\ and
-$\alpha$, upon which the Sommerfeld enhancement depends, matters a
-great deal for the predicted gamma-ray flux.
+the total flux in the other extremum of the Sommerfeld enhancement and
+about seven orders of magnitude larger than the total gamma-ray flux
+in the absence of any Sommerfeld enhancement. Thus, the exact value of
+the parameters \mv/\mdm\ and $\alpha$, upon which the Sommerfeld
+enhancement depends, matters a great deal for the predicted gamma-ray
+flux.
 
@@
 it opens up new possibilities for detecting \DM\ annihilation and
-learning about the properties of the dark sector. Given a detailed
-knowledge of the density structure and velocity dispersion profile of
-the dwarf Spheroidal galaxies, much can be learned if we are able to
-detect the \DM\ annihilation signal coming from these \dSphs, by
-considering the relative signal and using the results of
-Sec.~\ref{sec:dsphs}, or by using the detailed velocity dispersion
-profile of the \dSphs\ \cite{2009arXiv0902.0362R}.
+learning about the properties of the dark sector. As we showed above,
+the combination of the Sommerfeld enhancement and the substructure
+that is believed to exists down to small scales leads to very large
+boost factors because of the combined effect of the higher densities
+of small subhalos as well as the fact that they are kinematically
+colder than the larger halos. This leads to predictions for the \DM\
+annihilation gamma-ray flux that are much larger than was previously
+believed. As we showed in Sec.~\ref{sec:dsphs}, in models in which
+there is a significant Sommerfeld enhancement at the level of the
+Galactic halo and its subhalos, the \DM\ annihilation signal from the
+dwarf Spheroidals outshines the extragalactic background flux and the
+predicted gama-ray fluxes are so large that they should be detectable
+by \Fermi. Given a detailed knowledge of the density structure and
+velocity dispersion profile of the dwarf Spheroidal galaxies, much can
+then be learned if we are able to detect the \DM\ annihilation signal
+coming from these \dSphs, by considering the relative signal and using
+the results of Sec.~\ref{sec:dsphs}, or by using the detailed velocity
+dispersion profile of the \dSphs\ \cite{2009arXiv0902.0362R}.
 

### 2009-02-27 tex/somm.tex
Updated figures + text

@@
 \newcommand{\degree}{^\circ}
+\newcommand{\Jtilde}{\ensuremath{\tilde{J}}}
 
@@
 the total flux of gamma-rays from \DM\ annihilation that would be
-detectable by Fermi and find that in most of the models that posit the
-Sommerfeld enhancement the \DM\ annihilation signal from the dwarf
-Spheroidals shines above the diffuse extragalactic gamma-ray
-background, which is the only background signal at the high Galactic
-lattitudes of these satellite galaxies. Finally, we also consider the
-prospects for detecting smaller subhalos of the Galaxy by computing
-annihilation fluxes of subhalos in the \VL\-2 simulation. We will see
-that whether these subhalos will be detectable in the near future is
-highly dependent on the exact details of the Sommerfeld enhancement.
+detectable by the Fermi Gamma-ray Space Telescope (\Fermi) and find
+that in most of the models that posit the Sommerfeld enhancement the
+\DM\ annihilation signal from the dwarf Spheroidals shines above the
+diffuse extragalactic gamma-ray background, which is the only
+background signal at the high Galactic lattitudes of these satellite
+galaxies. Finally, we also consider the prospects for detecting
+smaller subhalos of the Galaxy by computing annihilation fluxes of
+subhalos in the \VL\ II simulation. We will see that whether these
+subhalos will be detectable in the near future is highly dependent on
+the exact details of the Sommerfeld enhancement.
 
@@
 \begin{tabular}{lrr@{.}lr@{}lr@{}lr@{.}lr@{.}lr@{.}l@{ $\times$ }lc}
-dSph & \dist\ & \multicolumn{2}{c}{\sigv}\ & \multicolumn{2}{c}{\rmax} & \multicolumn{2}{c}{\vmax} & \multicolumn{2}{c}{\rhos} & \multicolumn{2}{c}{\rs} & \multicolumn{3}{c}{$M(<\rs)$} & References\\
+dSph & \dist\ & \multicolumn{2}{c}{$\sigma_0$}\ & \multicolumn{2}{c}{\rmax} & \multicolumn{2}{c}{\vmax} & \multicolumn{2}{c}{\rhos} & \multicolumn{2}{c}{\rs} & \multicolumn{3}{c}{$M(<\rs)$} & References\\
 & [kpc] & \multicolumn{2}{c}{[km s$^{-1}$]} & \multicolumn{2}{c}{[kpc]} & \multicolumn{2}{c}{[km s$^{-1}$]} & \multicolumn{2}{c}{[$\Msun$ kpc$^{-3}$]} & \multicolumn{2}{c}{[kpc]} & \multicolumn{3}{c}{[$\Msun$]}&\\[4pt]
@@
 
+At this point we can also make a connection to another quantity that
+is often used to describe the astrophysical contribution to the \DM\
+annihilation signal, $J$. This quantity is generally defined as
+\begin{equation}
+J \equiv \frac{1}{\Delta \Omega} \int_{\Delta \Omega} \dd \Omega\, \int_{\mbox{{\footnotesize los}}} \dd l\, \rho^2\, ,
+\end{equation}
+and is therefore equal to the line-of-sight integral over the density
+squared averaged over solid angle. We can define a slight variant of
+this which takes into account the Sommerfeld enhancement
+\begin{equation}
+\Jtilde \equiv \frac{1}{\Delta \Omega} \int_{\Delta \Omega} \dd \Omega\, \int_{\mbox{{\footnotesize los}}} \dd l\, \rho^2 \somm\, ,
+\end{equation}
+in which \somm\ is the Sommerfeld enhancement averaged over the
+distribution of relative velocities. If we work this out for the
+dwarf Spheroidals, using the expression from \eqnname\
+(\ref{eq:lumrhosrsNFW}) for the line-of-sight integral averaged over
+the solid angle spanned by the scale radius \rs\ we find that this
+\Jtilde\ is approximately equal to the total boost \totalboost, with a
+pre-factor \order(1). Therefore in the case of the \dSphs\ studied
+here the \Jtilde\ can be approximately read off from the top panels of
+Figure \ref{fig:subboost}.
+
 It is clear from Figure \ref{fig:dsph} that the \DM\ annihilation flux
@@
 values given here. In order to compare these signals to the expected
-background flux of extragalactic gamma rays (which is the only
-relevant background since these \dSphs\ are at high Galactic
-lattitude), we have integrated the extragalactic gamma ray flux as
-measured by EGRET over the energy range between 4 GeV -- 250 GeV and
-multiplied it with the solid angle that the \dSphs\ span on the
-sky. The extragalactic gamma ray flux is given by a power law
+background flux we will only consider contributions to the background
+flux of extragalactic gamma rays. Although there are also
+contributions to the background flux from pion-decay, inverse Compton
+scattering, and Bremsstrahlung, at the high Galactic lattitudes at
+which the dwarf Spheroidals are found these are all of the same order
+or smaller than the extragalactic background
+\cite{2004ApJ...613..962S}. We have integrated the extragalactic gamma
+ray flux as measured by EGRET over the energy range between 4 GeV --
+250 GeV and multiplied it with the solid angle that the \dSphs\ span
+on the sky. The extragalactic gamma ray flux is given by a power law
 \cite{1998ApJ...494..523S}
@@
 \end{equation}
-where $k = 7.3 \times 10^{-9}$, $\alpha$ = 2.10, and $E_0 = 0.45$
-GeV. We have included the background flux of extragalactic photons in
-Figure \ref{fig:dsph}. This background flux is in all cases of the
-order of 10$^{-13}$ photons cm$^{-2}$ s$^{-1}$, such that the \DM\
+where $k = 7.3 \times 10^{-6}$ cm$^{-2}$ s$^{-1}$ sr$^{-1}$
+GeV$^{-1}$, $\alpha$ = 2.10, and $E_0 = 0.45$ GeV. We have included
+the background flux of extragalactic photons in Figure
+\ref{fig:dsph}. This background flux is in all cases of the order of
+10$^{-9}$ -- 10$^{-10}$ photons cm$^{-2}$ s$^{-1}$, such that the \DM\
 annihilation boosted by substructure and Sommerfeld enhancement
-outshine this background in all of the considered models, and by many
+outshine this background in most of the parameter space, and by a few
 orders of magnitude in most models. Given that the flux above 100 MeV
@@
 these \dSphs\ should be detectable by \Fermi\ in many of the models
-parameterized by ($\mv/\mdm,\alpha$). We conclude that the prospects
-for detecting \DM\ annihilation from the \dSphs\ with \Fermi\ are very
-good in models that include a significant Sommerfeld enhancement.
+parameterized by ($\mv/\mdm,\alpha$). The Ursa Major II \dSph\ is the
+only \dSph\ for which the \DM\ annihilation dominates the background
+in all the regions of parameter space. This shows that the newly
+discovered, most-dark-matter-dominated satellite galaxies are possibly
+the best targets for the indirect detection of \DM. We conclude that
+the prospects for detecting \DM\ annihilation from the \dSphs\ with
+\Fermi\ are very good in models that include a significant Sommerfeld
+enhancement.
 
@@
 
-{\bf Acknowledgements:} It is a pleasure to thank Mulin Ding, David
-Hogg, and Neal Weiner for comments and assistance. This research was partially
-supported by NASA (grant NNX08AJ48G).
+{\bf Acknowledgements:} It is a pleasure to thank Ilias Cholis, Mulin
+Ding, David Hogg, and Neal Weiner for comments and assistance. This
+research was partially supported by NASA (grant NNX08AJ48G).
 

### 2009-03-02 tex/somm.tex
As submitted to PRD

@@
 Speroidal galaxies in the Galactic halo are such that they outshine
-the extragalactic background gamma-ray flux and should be detectable
+the background gamma-ray flux and should be detectable
 by \Fermi.
@@
 \end{equation}
-By a simple estimate of the velocity dispersion using the virial
-theorem, we can use the circular velocity at \rminustwo\ as a measure
-of the halos velocity dispersion. This would give $\sigv \propto
-m^{1/3.5}$. Alternatively, we could use the circular velocity at the
-virial radius as a measure of the velocity dispersion. In the case of
-an \NFW\ profile we have that
+Using the circular velocity at \rminustwo\ as a measure of the halo's
+velocity dispersion gives $\sigv \propto m^{1/3.5}$. Alternatively, we
+could use the circular velocity at the virial radius as a measure of
+the velocity dispersion. In the case of an \NFW\ profile we have that
 \begin{equation}
@@
 \begin{equation}
-\sigv \propto M^{0.32}\, .
+\sigv \propto m^{0.32}\, .
 \end{equation}

### 2009-03-03 tex/somm.tex
Removed old figures; fixed small integration limit error
Fixed some spelling mistakes, made figures nicer

@@
 factors for the detection of dark matter annihilation from dwarf
-Speroidal galaxies in the Galactic halo are such that they outshine
+Spheroidal galaxies in the Galactic halo are such that they outshine
 the background gamma-ray flux and should be detectable
@@
 typically of the order 1 -- 10. Therefore, a different mechanism must
-be reponsible for the large boosts necessary. One such mechanism that
+be responsible for the large boosts necessary. One such mechanism that
 has received much attention lately is the Sommerfeld enhancement
@@
 diffuse extragalactic gamma-ray background, which is the only
-background signal at the high Galactic lattitudes of these satellite
+background signal at the high Galactic latitudes of these satellite
 galaxies. Finally, we also consider the prospects for detecting
@@
 \caption{Sommerfeld enhancement factor as a function of the particle
-physics parameters assuming a one-particle Maxwell-Boltzman velocity
+physics parameters assuming a one-particle Maxwell-Boltzmann velocity
 distribution with velocity dispersion \sigv\ for four different orders
@@
 \end{equation}
-In this \Eth\ is the treshold energy of the detector, $i$ denotes
+In this \Eth\ is the threshold energy of the detector, $i$ denotes
 (possibly) different final states of the \DM\ annihilation, $\frac{\dd
@@
 \begin{equation} 
-\int_{4 \mbox{{\footnotesize GeV}}}^{250 \mbox{{\footnotesize GeV}}}\dd E\, \frac{\dd
+\int_{4 \mbox{ {\footnotesize GeV}}}^{250 \mbox{ {\footnotesize GeV}}}\dd E\, \frac{\dd
 \Ngamma}{\dd E} \approx 0.5\, .
@@
 the results we obtain below. This concentration-halo mass
-concentration has not just been establised in numerical simulations:
+concentration has not just been established in numerical simulations:
 statistical weak lensing analysis of luminous red galaxies and
@@
 \mv/\mdm and especially in resonance regions, the reduced boost can
-reach high values, signalling a large extra boost due to the
+reach high values, signaling a large extra boost due to the
 kinematically coldness of the substructure.  For small values of
@@
 contributions to the background flux from pion-decay, inverse Compton
-scattering, and Bremsstrahlung, at the high Galactic lattitudes at
+scattering, and Bremsstrahlung, at the high Galactic latitudes at
 which the dwarf Spheroidals are found these are all of the same order
@@
 these \dSphs\ should be detectable by \Fermi\ in many of the models
-parameterized by ($\mv/\mdm,\alpha$). The Ursa Major II \dSph\ is the
+parametrized by ($\mv/\mdm,\alpha$). The Ursa Major II \dSph\ is the
 only \dSph\ for which the \DM\ annihilation dominates the background
@@
 \caption{Left: Histogram of the predicted gamma-ray flux coming from
-indivual subhalos contained within the tidal radius of the main halo
+individual subhalos contained within the tidal radius of the main halo
 in \VL\ II for three random observer placed at a distance of 8.5 kpc
@@
 physics parameters \mv/\mdm\ and $\alpha$ which give rise to the
-smallest Sommerfeld enahcements in Figures \ref{fig:VL23_2} and
+smallest Sommerfeld enhancements in Figures \ref{fig:VL23_2} and
 \ref{fig:VL23_3}, \ie, \mv/\mdm\ = 10$^{-2}$ and $\alpha$ = 0.001. The
@@
 cross section forces us to reconsider statements made in the past
-about the expected signal of \DM\ anihilation from cold substructures
+about the expected signal of \DM\ annihilation from cold substructures
 in the halo of the Galaxy. Since the Sommerfeld enhancement depends
-criticically on the velocity dispersion of the substructure in
+critically on the velocity dispersion of the substructure in
 question, as well as on the details of the particle physics model used
@@
 dwarf Spheroidals outshines the extragalactic background flux and the
-predicted gama-ray fluxes are so large that they should be detectable
+predicted gamma-ray fluxes are so large that they should be detectable
 by \Fermi. Given a detailed knowledge of the density structure and
@@
 
-{\bf Acknowledgements:} It is a pleasure to thank Ilias Cholis, Mulin
+{\bf Acknowledgments:} It is a pleasure to thank Ilias Cholis, Mulin
 Ding, David Hogg, and Neal Weiner for comments and assistance. This

### 2009-04-16 tex/somm.tex
sommboost: updates to the text after the referee report

@@
 radius. Another set of parameters is given by a concentration
-parameter and a mass. The concentration for both the \GNFW\ profile as
-for the Einasto profile is given by
+parameter and a mass. The virial radius \Rvir\ of a halo is defined to
+be the radius within which the mean density of the halo is $\Delta$
+times the critical density $\rhoc$, where $\Delta = 178\,
+\omegam^{0.45} \approx 100$
+\cite{1996MNRAS.282..263E,1998ApJ...495...80B}. The virial mass is
+simply the mass contained within a sphere of radius \Rvir
+\begin{equation}\label{eq:virialmass}
+\Mvir \equiv \frac{4\pi}{3} \Delta \rhoc \Rvir^3\, .
+\end{equation}
+The concentration for both the \GNFW\ profile as for the Einasto
+profile is then defined by
 \begin{equation}\label{eq:concdef}
@@
 incomplete gamma function. The density can then be expressed in terms
-of the concentration by using the definition of the virial mass:
-$\Mvir \equiv 4\pi/3 \Delta \rhoc \Rvir^3$, where $\Delta = 178\,
-\omegam^{0.45} \approx 100$
-\cite{1996MNRAS.282..263E,1998ApJ...495...80B}, and \rhoc\ is the
-critical density.
+of the concentration by using the definition of the virial mass given
+in \eqnname~(\ref{eq:virialmass}).
 
@@
 $\alpha$ is small. Indeed, in the 1/$v$ regime the Sommerfeld
-enhancement is $\sim \alpha/\beta$ $\approx$ $10^{-2}$ for $\alpha =
+enhancement is $\sim \alpha/\beta$ $\approx$ $10^{2}$ for $\alpha =
 10^{-3}$, such that the Sommerfeld enhancement has fully saturated

### 2009-04-16 tex/somm.tex
sommboost: some more small updates

@@
 \end{equation}
-A characteristic mass is given by the virial mass \Mvir. The virial
-mass of the \GNFW\ profile is
+A characteristic mass of a halo is given by the virial mass defined
+above. The virial mass of the \GNFW\ profile in terms of the
+characteristic density \rhos\ and the characteristic radius \rs\ is
 \begin{equation}\label{eq:Mvirdef}
@@
 form $f(c) = \ln(1+c)-c/(1+c)$. The virial mass for the Einasto
-profile is
+profile in terms of the characteristic density \rhominustwo\ and the
+characteristic radius \rminustwo\ is
 \begin{equation}

### 2009-04-16 tex/somm.tex
sommboost: updated figures that were too small everywhere

@@
 \includegraphics{hist_-2_-3_10_2.eps}
-\caption{Left: Histogram of the predicted gamma-ray flux coming from
+\caption{Top: Histogram of the predicted gamma-ray flux coming from
 individual subhalos contained within the tidal radius of the main halo
@@
 from the center of the main \VL\ halo for \mv/\mdm\ = 10$^{-2}$ and
-$\alpha = 10^{-3}$. Right: Histogram of the total boosts of the
+$\alpha = 10^{-3}$. Bottom: Histogram of the total boosts of the
 subhalos for the same values of \mv/\mdm\ and $\alpha$.}%\

### 2009-04-16 py/2denhance.py

@@ -75,12 +75,12 @@ params = {'backend': 'ps',
 rcParams.update(params)
 #rc('text',usetex=True)
 fig= figure()
-im= imshow(data['S'],origin='lower',#cmap=cm.gray,
+im= imshow(data['S'],origin='lower',cmap=cm.gist_yarg,
        extent=[-3,-1,-5,-2],aspect=2./3.)
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)
 xlabel(r'$\log_{10}\alpha$')#,fontsize=16)
-text(-1.75,-4.75,r'$\beta = 10^{-4}$')#,fontsize=16)
+text(-1.75,-4.75,r'$\beta = 10^{'+sys.argv[1]+'}$',color='w')#,fontsize=16)
 CB1= colorbar(im,shrink=0.87)
 levels=linspace(0,5,11)
 #cont= contour(data['S'],levels,origin='lower',linewidths=1,colors='k',

### 2009-04-16 py/avgenhance.py

@@ -92,12 +92,12 @@ params = {'backend': 'ps',
 rcParams.update(params)
 #rc('text',usetex=True)
 fig= figure()
-im= imshow(S,origin='lower',#cmap=cm.gray,
+im= imshow(S,origin='lower',cmap=cm.gist_yarg,
        extent=[-3,-1,-5,-2],aspect=2./3.)
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)
 xlabel(r'$\log_{10}\alpha$')#,fontsize=16)
-text(-1.75,-4.75,r'$\sigma_v = 10^{'+str(logb)+r'}$')#,fontsize=16)
+text(-1.75,-4.75,r'$\sigma_v = 10^{'+str(logb)+r'}$',color='w')#,fontsize=16)
 CB1= colorbar(im,shrink=0.87)
 levels=linspace(0,5,11)
 #cont= contour(data['S'],levels,origin='lower',linewidths=1,colors='k',

### 2009-04-16 py/calc_reducedboost.py

@@ -142,12 +142,12 @@ params = {'backend': 'ps',
 rcParams.update(params)
 #rc('text',usetex=True)
 fig= figure()
-im= imshow(redboostS,origin='lower',#cmap=cm.gray,
+im= imshow(redboostS,origin='lower',cmap=cm.gist_yarg,
        extent=[-3,-1,-5,-2],aspect=2./3.)
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)
 xlabel(r'$\log_{10}\alpha$')#,fontsize=16)
-text(-2,-4.75,r'$M = 10^{'+str(logMh)+r'} M_{\odot}$',color='w')#,fontsize=16)
+text(-2,-4.75,r'$M = 10^{'+str(logMh)+r'} M_{\odot}$')#,color='w')#,fontsize=16)
 CB1= colorbar(im,shrink=0.87)
 levels=linspace(0,5,11)
 #cont= contour(data['S'],levels,origin='lower',linewidths=1,colors='k',

### 2009-04-16 py/hist_VL.py

@@ -193,15 +193,15 @@ plotfilename2= 'hist_'+sys.argv[1]+'_'+sys.argv[2]+'_'+sys.argv[3]+'_2.eps'
 plotfilename3= 'hist_'+sys.argv[1]+'_'+sys.argv[2]+'_'+sys.argv[3]+'_3.eps'
 
 #Plotting parameters
-fig_width = 3.25  # width in inches
-fig_height = 1.6      # height in inches
+fig_width = 3#3.25  # width in inches
+fig_height =4.5      # height in inches
 fig_size =  [fig_width,fig_height]
 params = {'backend': 'ps',
           'axes.labelsize': 8,
           'text.fontsize': 8,
           'legend.fontsize': 8,
-          'xtick.labelsize':8,
-          'ytick.labelsize':8,
+          'xtick.labelsize':10,
+          'ytick.labelsize':10,
           'text.usetex': True,
           'figure.figsize': fig_size}
 rcParams.update(params)
@@ -213,25 +213,40 @@ fig= figure(1)
 nbins= 20
 
 subplots_adjust(left=None, bottom=None, right=None, top=None,wspace=None, hspace=0.4)
-subplot(1,2,1)
+subplot(2,1,1)
 hist(sublumobs[0,:,0],bins=nbins,histtype='step',ec='black')
 hist(sublumobs[1,:,0],bins=nbins,histtype='step',ec='black',linestyle='dashed')
 hist(sublumobs[2,:,0],bins=nbins,histtype='step',ec='black',linestyle='dashdot')
 hist(sublumobs[3,:,0],bins=nbins,histtype='step',ec='black',linestyle='dotted')
 axis([0,3,0,2000])
 ylabel(r'$N_{\mbox{{\footnotesize sub}}}$')#,fontsize=16)
-xlabel(r'$\log_{10}D$ [kpc]')#,fontsize=16)
+xlabel(r'$\log_{10}D$ $[\mathrm{kpc}]$')#,fontsize=16)
 
-ax=subplot(1,2,2)
+ax=subplot(2,1,2)
+fig.subplots_adjust(hspace=.3)
 hist(sublum[:,9],bins=nbins,histtype='step',ec='black')
 axis([3,8,0,2000])
[…]

### 2009-04-16 py/plot_2dboost.py

@@ -93,12 +93,12 @@ params = {'backend': 'ps',
 rcParams.update(params)
 #rc('text',usetex=True)
 fig= figure()
-im= imshow(boostS,origin='lower',#cmap=cm.gray,
+im= imshow(boostS,origin='lower',cmap=cm.gist_yarg,
        extent=[-3,-1,-5,-2],aspect=2./3.)
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)
 xlabel(r'$\log_{10}\alpha$')#,fontsize=16)
-text(-2,-4.75,r'$M = 10^{'+str(logMh)+r'} M_{\odot}$')#,fontsize=16)
+text(-2,-4.75,r'$M = 10^{'+str(logMh)+r'} M_{\odot}$',color='w')#,fontsize=16)
 CB1= colorbar(im,shrink=0.87)
 levels=linspace(0,5,11)
 #cont= contour(data['S'],levels,origin='lower',linewidths=1,colors='k',

### 2009-04-16 py/plot_2dboostabundance.py

@@ -121,12 +121,12 @@ params = {'backend': 'ps',
 rcParams.update(params)
 #rc('text',usetex=True)
 fig= figure()
-im= imshow(boostS,origin='lower',#cmap=cm.gray,
+im= imshow(boostS,origin='lower',cmap=cm.gist_yarg,
        extent=[-3,-1,-5,-2],aspect=2./3.)
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)
 xlabel(r'$\log_{10}\alpha$')#,fontsize=16)
-text(-1.75,-4.75,r'$n$ = '+str(n))#,fontsize=16)
+text(-1.75,-4.75,r'$n$ = '+str(n),color='w')#,fontsize=16)
 CB1= colorbar(im,shrink=0.87)
 levels=linspace(0,5,11)
 #cont= contour(data['S'],levels,origin='lower',linewidths=1,colors='k',

### 2009-04-16 py/plot_2dboostfree.py

@@ -130,12 +130,12 @@ params = {'backend': 'ps',
 rcParams.update(params)
 #rc('text',usetex=True)
 fig= figure()
-im= imshow(boostS,origin='lower',#cmap=cm.gray,
+im= imshow(boostS,origin='lower',cmap=cm.gist_yarg,
        extent=[-3,-1,-5,-2],aspect=2./3.)
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)
 xlabel(r'$\log_{10}\alpha$')#,fontsize=16)
-text(-2.12,-4.75,r'$m_0 = 10^{'+sys.argv[5]+r'} M_{\odot}$')#,fontsize=16)
+text(-2.12,-4.75,r'$m_0 = 10^{'+sys.argv[5]+r'} M_{\odot}$',color='w')#,fontsize=16)
 CB1= colorbar(im,shrink=0.87)
 levels=linspace(0,5,11)
 #cont= contour(data['S'],levels,origin='lower',linewidths=1,colors='k',

### 2009-04-16 py/plot_2dboostprofiles.py

@@ -112,15 +112,15 @@ params = {'backend': 'ps',
 rcParams.update(params)
 #rc('text',usetex=True)
 fig= figure()
-im= imshow(boostS,origin='lower',#cmap=cm.gray,
+im= imshow(boostS,origin='lower',cmap=cm.gist_yarg,
        extent=[-3,-1,-5,-2],aspect=2./3.)
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)
 xlabel(r'$\log_{10}\alpha$')#,fontsize=16)
 if einasto==1:
-    text(-1.75,-4.75,r'$\alpha$ = '+str(gamma))#,fontsize=16)
+    text(-1.75,-4.75,r'$\alpha$ = '+str(gamma),color='w')#,fontsize=16)
 else:
-    text(-1.75,-4.75,r'$\gamma$ = '+str(gamma))#,fontsize=16)
+    text(-1.75,-4.75,r'$\gamma$ = '+str(gamma),color='w')#,fontsize=16)
 CB1= colorbar(im,shrink=0.87)
 levels=linspace(0,5,11)
 #cont= contour(data['S'],levels,origin='lower',linewidths=1,colors='k',

### 2009-04-16 py/plot_dsph.py

@@ -164,7 +164,7 @@ restorefile.close()
 
 
 #also calculate the background in this range
-back= log10(2.69156*10**-10*2*pi*(1.-cos(rs/D)))
+back= log10(2.69156*10**-7*2*pi*(1.-cos(rs/D)))
 
 #Then plot
 plotfilename= 'flux_'+sys.argv[1]+'_'+str(nas)
@@ -187,7 +187,7 @@ params = {'backend': 'ps',
 rcParams.update(params)
 #rc('text',usetex=True)
 fig= figure()
-im= imshow(flux,origin='lower',#cmap=cm.gray,
+im= imshow(flux,origin='lower',cmap=cm.gist_yarg,
        extent=[-3,-1,-5,-2],aspect=2./3.)
 axis([-3,-1,-5,-2])
 ylabel(r'$\log_{10}(m_{\phi}/m_\chi)$')#,fontsize=16)

### 2009-04-16 tex/response.txt
sommboost: added response to the referee

Dear Editor,

Please find below the response to the referees' comments for the
submission "Substructure boosts to dark matter annihilation from
Sommerfeld enhancement" by myself (submission # DC10513).

I have made changes following the referee's suggestions and have also
update figures 1, 2, 5, 6, and 7 to make the label in these plots more
clear to the reader.

A detailed list of the changes is:

- I slightly rewrote the paragraph containing Eq. 20 to include the
definition of the virial radius;

- I fixed a type which the referee found in the paragraph below Eq. 39;

- I changed the color of the label in figures 1, 2, 5, 6, and 7 from
  black to white;

- I enlarged figures 9, 10, 12, 14, arranging the two panels now below
  each other. I updated the caption of figure 10 to reflect this
  change.

Please contact me by the information given below with any questions
regarding this re-submission.

With kind regards,

Jo Bovy

-
Jo Bovy
Ctr. for Cosmology & Particle Physics, 
Dept. of Physics
New York University
4 Washington Place
New York, NY 10003
jb2777@nyu.edu

In what follows the referee's comments are indented.

==================================================================

	Dear Dr. Bovy:

	The manuscript ``Substructure boosts to dark matter
	annihilation from Sommerfeld enhancement'' (dc10513) by Bovy,J
	has been reviewed by one of our referees.  Comments from the
	report are enclosed for your consideration.

	When you resubmit your manuscript, please include a summary of
	the changes made, and a brief response to all recommendations
	and criticisms.

	Sincerely,

	Rashmi Ray
        Senior Assistant Editor
        Physical Review D

	----------------------------------------------------------------
	Report:
	----------------------------------------------------------------
	This is a very good paper. It makes a careful, intelligent
	contribution to the hot topic of boost factors in Galactic
	dark matter annihilation.  I am also impressed with the
	clarity of exposition.

	I have two small suggestions and a typo to convey to the
	author, for his/her optional consideration.

	1. Typo: toward the end of the paragraph following Eq 39, an
           enhancement of 10^-2 is reported; I believe the author
           means 10^+2.

That is correct, I did mean 10^+2. This has been corrected in the text.

	2. In Eq 20, the virial radius r_vir is used; I don't
	   believe it is defined in the text.

I have added the definition of the virial radius in the paragraph
containing Eq.20 before the definition of the concentration. I have
also moved up the definition of the virial mass from where it was
before (it used to be at the end of that same paragraph).

	3. Figs 9, 10, 12, 14 are too small in the copy I received
	   from APS.  Perhaps this is not the author's issue.

This was my fault. I have replaced these figures with larger versions.
-------------------------------------------------------------------------

### 2009-04-17 tex/somm.tex
sommboost: forgot a ,

@@
 while the velocity dispersion of the Galactic halo might be of the
-order of 100 km s$^{-1}$ the actual velocity dispersion relevant to
+order of 100 km s$^{-1}$, the actual velocity dispersion relevant to
 the Sommerfeld enhancement might be much smaller for halo \DM\

### 2009-04-18 tex/response.txt
sommboost: new version

@@
 
-A detailed list of the changes is:
+A detailed list of the main changes is:
 
@@
 - I enlarged figures 9, 10, 12, 14, arranging the two panels now below
-  each other. I updated the caption of figure 10 to reflect this
-  change.
+  each other. I updated the text and the caption of figure 10 to
+  reflect this change.
+
+- In the conclusion I elaborated a fit further on the implications of
+  the possibility that the Galactic halo consists for a large part of
+  cold dark matter streams, adding a reference to a paper by Afshordi,
+  Mohayaee, and Bertschinger which discusses boosts due to such tidal
+  debris.
+
+- I also updated all of the references to papers which have gotton
+  published since I first submitted this paper.
+
+In addition to these I made a number of cosmetic changes and slight
+rewordings to the text to improve the clarity of the exposition, the
+most important of which are:
+
+- Some rewordings in the abstract (including spelling out the proper
+  name of the Fermi telescope, and emphasizing more clearly what is
+  new in the present paper);
+
+- I renamed Sec. III A "Some details of the particle physics model"
+  instead of "Some details of the particle physics"
+
+- I moved up the sentence in the paragraph above Eq. 13 which states
+  that the DM annihilation flux depends on the density squared to the
+  beginning of that section, such that the point of discussing the
+  dark matter profile at length is clear at the start of the section
+  on the shape of the profile.
 

### 2009-04-18 tex/somm.tex
sommboost: new version

@@
 \begin{abstract}
-I discuss the implications of the recently introduced Sommerfeld
-enhancement of the dark matter annihilation cross section for the
-detection of subhalos in the Galactic halo.  In addition to the boost
-to the dark matter annihilation cross section from the high densities
-of these subhalos with respect to the main halo, an additional boost
-due to Sommerfeld enhancement results from the fact that they are
-kinematically colder than the Galactic halo. If we further believe the
-generic prediction of \CDM\ that in each subhalo there is an abundance
-of substructure which is approximately self-similar to that of the
-Galactic halo, then additional boosts coming from the density
-enhancements of these small substructures and their small velocity
-dispersions enhance the dark matter annihilation cross section even
-further. I show that very large boost factors ($10^5$ -- $10^9$) are
-obtained in a large class of models. The implications of these boost
-factors for the detection of dark matter annihilation from dwarf
-Spheroidal galaxies in the Galactic halo are such that they outshine
-the background gamma-ray flux and should be detectable
-by \Fermi.
+The recently introduced Sommerfeld enhancement of the dark matter
+annihilation cross section has important implications for the
+detection of dark matter annihilation in subhalos in the Galactic
+halo. In addition to the boost to the dark matter annihilation cross
+section from the high densities of these subhalos with respect to the
+main halo, an additional boost caused by the Sommerfeld enhancement
+results from the fact that they are kinematically colder than the
+Galactic halo. If we further believe the generic prediction of \CDM\
+that in each subhalo there is an abundance of substructure which is
+approximately self-similar to that of the Galactic halo, then I show
+that additional boosts coming from the density enhancements of these
+small substructures and their small velocity dispersions enhance the
+dark matter annihilation cross section even further. I find that very
+large boost factors ($10^5$ to $10^9$) are obtained in a large class
+of models. The implications of these boost factors for the detection
+of dark matter annihilation from dwarf Spheroidal galaxies in the
+Galactic halo are such that, generically, they outshine the background
+gamma-ray flux and are detectable by the Fermi Gamma-ray Space
+Telescope.
 \end{abstract}
@@
 self-annihilation seems unlikely considering the standard
-self-annihilation cross section --- $\sigmaannv_{\mathrm{ann}}
-\lesssim 3 \times 10^{-26} $ cm$^3$ s$^{-1}$ for consistency with the
-observed relic density. However, there are various mechanisms to boost
-the annihilation signal to a value larger by a few orders of
-magnitude, e.g., clumpiness of the dark halo
+self-annihilation cross section; $\sigmaannv_{\mathrm{ann}} \lesssim 3
+\times 10^{-26} $ cm$^3$ s$^{-1}$ for consistency with the observed
+relic density. However, there are various mechanisms to boost the
+annihilation signal to a value larger by a few orders of magnitude,
+e.g., clumpiness of the dark halo
 \cite{1993ApJ...411..439S,1999PhRvD..59d3506B,2008A&A...479..427L,2008ApJ...686..262K}
@@
 (PAMELA) satellite, which saw a sharp upturn in the positron fraction
-$e^+/(e^++e^-)$ from 10 -- 100 GeV \cite{Adriani:2008zr}, and the
+$e^+/(e^++e^-)$ from 10 to 100 GeV \cite{Adriani:2008zr}, and the
 Advanced Thin Ionization Calorimeter (ATIC) group, which reported an
-excess in the combined number of electrons and positrons at energies
-500 -- 800 GeV \cite{2008Natur.456..362C}. The PAMELA result on the
-positron fraction confirms earlier results
+excess over the expected background in the combined number of
+electrons and positrons at energies 500 to 800 GeV
+\cite{2008Natur.456..362C}. The PAMELA result on the positron fraction
+confirms earlier results
 \cite{1969ApJ...158..771F,1975ApJ...199..669B,1987ApJ...312..183M,1994ApJ...436..769G,Barwick:1997ig,Beatty:2004cy,Aguilar:2007yf}
@@
 \cite{2008arXiv0810.5557H,ArkaniHamed:2008qn,2008arXiv0811.0399F,2008arXiv0811.1555I,2008arXiv0811.3357C,2008arXiv0811.3641C,2008arXiv0812.2196A}),
-a description which also naturally accounts for other experimental
+a description that also naturally accounts for other experimental
 anomalies, such as the ``WMAP haze''
@@
 
-Of course, the largest obstacle currently to modeling the PAMELA/ATIC
-signal with \DM\ annihilation is the very large boost factors ($10^4$
--- $10^5$) with respect to the fiducial $\langle \sigma v
-\rangle_{\mathrm{ann}} \approx 3 \times 10^{-26} $ cm$^3$ s$^{-1}$
-needed to lead to an observable annihilation flux. The boost factor
-from the clumpiness of the dark matter halo is expected to be small
+The largest obstacle to modeling the PAMELA/ATIC signal with \DM\
+annihilation is the very large boost factor ($10^4$ to $10^5$) with
+respect to the fiducial $\langle \sigma v \rangle_{\mathrm{ann}}
+\approx 3 \times 10^{-26} $ cm$^3$ s$^{-1}$ required to produce an
+observable annihilation flux. The boost factor from the clumpiness of
+the dark matter halo is expected to be small
 \cite{2008A&A...479..427L,2008Natur.454..735D,2008ApJ...686..262K,2008Natur.456...73S},
-typically of the order 1 -- 10. Therefore, a different mechanism must
+typically of the order 1 to 10. Therefore, a different mechanism must
 be responsible for the large boosts necessary. One such mechanism that
@@
 Sommerfeld enhancement is a generic effect present whenever there is
-an attractive force acting between the dark matter particles, \eg, due
-to a Yukawa interaction or a gauge interaction through vector bosons
+an attractive force acting between the dark matter particles, \eg, a
+Yukawa interaction or a gauge interaction through vector bosons
 \cite{ArkaniHamed:2008qn}. The main effect of this attractive force is
 to enhance the annihilation cross section $\sigma v$ with a factor
-proportional to $\beta^{-1}$ ($\beta \equiv v/c$), this is known as
+proportional to $\beta^{-1}$ ($\beta \equiv v/c$); this is known as
 ``$1/v$'' enhancement. There are regions in the parameter space made
@@
 in which the enhancement is near a resonance, where it approximately
-behaves as $\beta^{-2}$. However, the enhancement levels off at a
-certain $\beta$ due to the finite range of the attractive force ---
-large ($\alpha \sim\!10^{-1}$ -- $10^{-3}$) attractive forces with
-infinite range lead to a burst of \DM\ annihilation in the first \DM\
-halos formed at $z$ $\sim$ 100 -- 200 which would lead to observable
-effects incompatible with measurements of the diffuse extragalactic
-gamma-ray background today and the cosmic microwave background
-\cite{2008arXiv0810.3233K}. Since the typical velocities of dark
-matter particles in the Galactic halo today are of the order of the
-velocity dispersion of the halo, which is non-relativistic ($\beta\sim
-10^{-3}$), the cross section of dark matter particles annihilating in
-the halo could be significantly enhanced by this Sommerfeld
-enhancement.
+behaves as $\beta^{-2}$. However, the enhancement saturates at a
+certain $\beta$ because of the finite range of the attractive
+force---large ($\alpha \sim\!10^{-1}$ to $10^{-3}$) attractive forces
+with infinite range lead to a burst of \DM\ annihilation in the first
+\DM\ halos formed at $z$ $\sim$ 100 -- 200 which would lead to
+observable effects incompatible with measurements of the diffuse
+extragalactic gamma-ray background today and the cosmic microwave
+background \cite{2008arXiv0810.3233K}. Since the typical velocities of
+dark matter particles in the Galactic halo today are of the order of
+the velocity dispersion of the halo, which is non-relativistic
+($\beta\sim 10^{-3}$), the cross section of dark matter particles
+annihilating in the halo could be significantly enhanced by this
+Sommerfeld effect.
 
@@
 substructure. These substructures are both in the form of subhalos,
-self-bound objects orbiting in the Galactic halo, as well as in
+self-bound objects orbiting in the Galactic halo, and in the form of
 streams, relics of tidally disrupted subhalos. Since these
 substructures are overdense regions in the halo, the dark matter
-annihilation cross section will be correspondingly higher, since
-annihilation cross sections scale as the density squared. The effects
+annihilation cross section will be correspondingly higher:
+Annihilation cross sections scale as the density squared. The effects
 of these kinds of substructure boosts have been studied extensively
@@
 of magnitude larger than the Sommerfeld enhancement acting in the
-Galactic halo.
-
-In addition to this larger boost due to the small
-velocity dispersion in the subhalos, these subhalos themselves are
-filled with an abundance of substructure which is approximately
-self-similar to that of the main halo. Each of these even smaller
-substructure has a larger density than the subhalo they are orbiting
-in (quantified by the well established concentration-mass
-relationships for \CDM\ halos
-\cite{2001MNRAS.321..559B,2001ApJ...554..114E}) as well as a smaller
-velocity dispersion. This opens up the possibility that the dark
-matter annihilation signal in these already cold subhalos can be
+Galactic halo taken as a whole.
+
+In addition to this larger boost caused by the small velocity
+dispersion in the subhalos, these subhalos themselves are filled with
+an abundance of substructure which is approximately self-similar to
+that of the main halo. Each of these even smaller substructure has a
+larger density than the subhalo in which they are orbiting (quantified
+by the well established concentration-mass relationships for \CDM\
+halos \cite{2001MNRAS.321..559B,2001ApJ...554..114E}) as well as a
+smaller velocity dispersion. This opens up the possibility that the
+dark matter annihilation signal in these already cold subhalos can be
 boosted even further by contribution of their substructure.
@@
 total combined boost from Sommerfeld enhancement and substructure can
-reach very large values (10$^5$ -- 10$^9$), and these boosts are only
+reach very large values (10$^5$ to 10$^9$), and these boosts are only
 weakly dependent on the assumptions that we have to make about the
@@
 smaller subhalos of the Galaxy by computing annihilation fluxes of
-subhalos in the \VL\ II simulation. We will see that whether these
-subhalos will be detectable in the near future is highly dependent on
-the exact details of the Sommerfeld enhancement.
+subhalos in the \VL\ II simulation. Whether these subhalos will be
+detectable in the near future depends on the exact details of the
+Sommerfeld enhancement.
 
@@
 The simplest example of the Sommerfeld effect that contains all of the
-generic features described in the introduction is that of the
-Sommerfeld enhancement due to an attractive Yukawa type force mediated
-by a spin-0 boson $\phi$ between the \DM\ particles, and we adopt that
-description here. The Sommerfeld effect is a non-perturbative effect
-in the sense that in the language of quantum field theory it comes
-about from Feynman diagrams in which the force carrier is exchanged
-many times before the annihilation actually happens, giving rise to so
-called ``ladder diagrams''. However, we can study the Sommerfeld
-effect in simple, non-relativistic quantum mechanics (see the appendix
-in Ref.~\cite{ArkaniHamed:2008qn}). We are instructed to solve the
-radial Schr{\"o}dinger equation (in natural units)
+generic features described in the introduction is that from an
+attractive Yukawa type force mediated by a spin-0 boson $\phi$ between
+the \DM\ particles, and we adopt that description here. The Sommerfeld
+effect is a non-perturbative effect in the sense that in the language
+of quantum field theory it comes about from Feynman diagrams in which
+the force carrier is exchanged many times before the annihilation
+actually happens, giving rise to so called ``ladder
+diagrams''. However, we can study the Sommerfeld effect in simple,
+non-relativistic quantum mechanics (see the appendix in
+Ref.~\cite{ArkaniHamed:2008qn}). We are instructed to solve the radial
+Schr{\"o}dinger equation (in natural units)
 \begin{equation}\label{eq:sommdiff1}
@@
 \end{equation}
-From \eqnname\ (\ref{eq:sommdiff2}) it can be seen that the Sommerfeld
-enhancement is only a function of two variables, which can be chosen
-as $\alpha/\beta$ and $\alpha \mdm/\mv$. The qualitative features of
-\eqnname\ (\ref{eq:sommdiff2}) have been widely discussed before and
-are easily derived, and, therefore, we will merely mention them here
-without proof (see, \eg,
+\Eqnname~(\ref{eq:sommdiff2}) shows that the Sommerfeld enhancement is
+only a function of two variables, which can be chosen as
+$\alpha/\beta$ and $\alpha \mdm/\mv$. The qualitative features of
+\eqnname\ (\ref{eq:sommdiff2}) have been widely discussed and,
+therefore, we will merely mention them here without proof (see, \eg,
 Refs.~\cite{ArkaniHamed:2008qn,2008arXiv0812.0360L}). When the kinetic
@@
 high-energy photons in the early Universe \cite{2008arXiv0810.3233K},
-levels off due to the finite range of the Yukawa interaction at a
-velocity $\beta \approx \mv/\mdm$ and settles on a value $\somm \approx
-\alpha \mdm/\mv$. However, at some parameter combinations the Yukawa
-potential has bound states, at which there are large boosts to the
-annihilation cross section which go approximately as $\somm \propto
-\beta^{-2}$. These resonance regions, however, are cut-off at small
-$\beta$ due to the finite lifetime of the bound state. We expect these
-three different behaviors ($1/v$, saturation at low $v$, and resonance
-effects due to the existence of bound states) to be generic among
-different models of the Sommerfeld enhancement (see also the
-discussion in the appendix of Ref.~\cite{ArkaniHamed:2008qn}).
+saturates as a consequence of the finite range of the Yukawa
+interaction at a velocity $\beta \approx \mv/\mdm$ and settles on a
+value $\somm \approx \alpha \mdm/\mv$. However, at some parameter
+combinations the Yukawa potential has bound states, at which there are
+large boosts to the annihilation cross section which go approximately
+as $\somm \propto \beta^{-2}$. These resonance regions, however, are
+cut-off at small $\beta$ by the finite lifetime of the bound state. We
+expect these three different behaviors ($1/v$, saturation at low $v$,
+and resonance effects from the existence of bound states) to be
+generic among different models of the Sommerfeld enhancement (see also
+the discussion in the appendix of Ref.~\cite{ArkaniHamed:2008qn}).
 
@@
 In the Coulomb regime ($\mdm \beta^2 \gg \alpha \mv$) this means that
-the starting value should be such that $r \gg \alpha^2/\beta^2$;
-otherwise we should start at $r \gg \alpha \mdm/\mv$. Figure
-\ref{fig:sommb34} shows the Sommerfeld enhancement in the parameter plane
-made up off $\alpha$ and $\mv/\mdm$. All of the qualitative features
-we discussed above are present: In the Coulomb regime (lower part of
-the two panels) we see that the enhancement scales as $\alpha$,
-leveling off at $\somm \approx 1$ when $\alpha$ becomes small; The
-fact that the upper left corner of both panels is approximately the
-same shows that the Sommerfeld enhancement saturates at $\beta \approx
-\mv/\mdm$; The sharp diagonal lines in the upper right of both panels
-is the resonance region, in which the enhancement is sharply
-peaked. Because of the scaling properties of \eqnname\
-(\ref{eq:sommdiff2}) shifting $\log\beta$ is equivalent to shifting
-both $\log\alpha$ and $\log \mv/\mdm$ by the same amount. Therefore,
-the $\beta$ behavior of the enhancement can also be read off from
-these figures along diagonals. Inspection of the resonance diagonals
-in the second panel of Figure \ref{fig:sommb34} shows that the enhancement
-grows as $\beta^{-2}$ in these regions.
+the initial value for the integration should be such that $r \gg
+\alpha^2/\beta^2$; otherwise we should start at $r \gg \alpha
+\mdm/\mv$. Figure \ref{fig:sommb34} shows the Sommerfeld enhancement
+in the parameter plane made of $\alpha$ and $\mv/\mdm$. All of the
+qualitative features we discussed above are present: In the Coulomb
+regime (lower part of the two panels) we see that the enhancement
+scales as $\alpha$, saturating at $\somm \approx 1$ when $\alpha$
+becomes small; The fact that the upper left corner of both panels is
+approximately the same shows that the Sommerfeld enhancement saturates
+at $\beta \approx \mv/\mdm$; The sharp diagonal lines in the upper
+right of both panels show the resonance region, in which the
+enhancement is sharply peaked. Because of the scaling properties of
+\eqnname\ (\ref{eq:sommdiff2}), shifting $\log\beta$ is equivalent to
+shifting both $\log\alpha$ and $\log \mv/\mdm$ by the same
+amount. Therefore, the $\beta$ behavior of the enhancement can also be
+read off from these figures along diagonals. Inspection of the
+resonance diagonals in the second panel of Figure \ref{fig:sommb34}
+shows that the enhancement grows as $\beta^{-2}$ in these regions.
 
@@
 the halo has a constant velocity dispersion over the halo, \ie, that
-the halo is isothermal. This is not necessarily the case ---indeed, it
+the halo is isothermal. This is not necessarily the case---indeed, it
 is not if we believe that the density profiles of \DM\ halos are well
-fit by Einasto or \NFW\ profiles--- but it is not a bad assumption for
+fit by Einasto or \NFW\ profiles---but it is not a bad assumption for
 intermediate distances from the halo's center, \ie, between the
@@
 \end{equation}
-since the truncation does not significantly influence the
-result. Similarly, in the resonance region the $1/v^2$ behavior is
-conserved by the averaging. When the enhancement has leveled off the
-averaging has no effect.
+since the truncation does not significantly influence the result of
+these integrals. Similarly, in the resonance region the $1/v^2$
+behavior is conserved by the averaging. When the enhancement has
+leveled off the averaging has no effect.
 
@@
 \ref{fig:sommb34} is indeed conserved well by the averaging: this has
-to do with the fact that the most probable relative velocity is given
-by the velocity dispersion for a Maxwell-Boltzmann relative velocity
-distribution. The Sommerfeld enhancement attains very large values for
+to do with the fact that the most probable relative velocity in a
+Maxwell-Boltzmann relative velocity distribution is the velocity
+dispersion. The Sommerfeld enhancement attains very large values for
 subhalos with very small velocity dispersions, mostly, but not
 exclusively, in the resonance regions. This means that the smallest
-subhalos that exists in \CDM\ will get large boosts to their \DM\
+subhalos that exist in \CDM\ will get large boosts to their \DM\
 annihilation from Sommerfeld enhancement.
@@
 & \times \int_{\Delta\Omega}\,\frac{\dd\Omega}{4\pi}\,\int_{\mathrm{\los}} \dd l\,
-\rho^2(l)\, .
+\rho^2(l)\, ,
 \end{split}
 \end{equation}
-In this \Eth\ is the threshold energy of the detector, $i$ denotes
+where \Eth\ is the threshold energy of the detector, $i$ denotes
 (possibly) different final states of the \DM\ annihilation, $\frac{\dd
@@
 final annihilation products, such as the electrons and positrons
-supposedly seen by PAMELA/ATIC.
+observed by PAMELA/ATIC.
 
@@
 This structure quantity does not only depend on the mass and internal
-properties of the \DM\ halos in which we are interest, it depends on
+properties of the \DM\ halos in which we are interested, it depends on
 the particle physics parameters as well. We will keep all of the other
@@
 briefly describe the other quantities appearing in \eqnname\
-(\ref{eq:flux}) here.
-
-\subsection{Some details of the particle physics}\label{sec:partphys}
-
-For the annihilation cross section without taking account of any
-Sommerfeld enhancement we will adopt the fiducial value $\sigmaannv =
-3 \times 10^{-26} $ cm$^3$ s$^{-1}$, and we choose a fiducial mass for
-the \DM\ particle of \mdm\ = 700 GeV. The energy spectrum of the \DM\
-annihilation depends on the model used to describe the \DM\ and its
-interactions and it is outside of the scope of this paper to provide a
-detailed treatment of this for different \DM\ models. Therefore, we
-will adopt an annihilation spectrum for \DM\ annihilation into photons
+(\ref{eq:flux}).
+
+\subsection{Some details of the particle physics model}\label{sec:partphys}
+
+For the annihilation cross section prior to any Sommerfeld enhancement
+we will adopt the fiducial value $\sigmaannv = 3 \times 10^{-26} $
+cm$^3$ s$^{-1}$, and we choose a fiducial mass for the \DM\ particle
+of \mdm\ = 700 GeV. The energy spectrum of the \DM\ annihilation
+depends on the model used to describe the \DM\ and its interactions;
+it is outside of the scope of this paper to provide a detailed
+treatment of this for different \DM\ models. Therefore, we will adopt
+an annihilation spectrum for \DM\ annihilation into photons
 appropriate for the newer models that were proposed to explain the
@@
 \end{equation}
-We find a similar number if we integrate the spectrum due to one
+We find a similar number if we integrate the spectrum from one
 cascade step. It is interesting to note that these photon spectra have
 fewer photons in them, about an order of magnitude less, than the
-annihilation spectra due to neutralino annihilation which are often
+annihilation spectra from neutralino annihilation which are often
 used \cite{1998APh.....9..137B,2004PhRvD..70j3529F}. Of course, since
@@
 
-The density profile of dark matter halos has been discussed
-extensively in the last few years. While the shape of the outer
-density profile seems to have converged in the various numerical
+Since the photon flux from \DM\ annihilations depends on the density
+squared of the \DM\, the exact form of the inner density profile has
+important ramifications for the detectability of \DM\
+annihilation. The density profile of dark matter halos has been
+discussed extensively in the last few years. While the shape of the
+outer density profile seems to have converged in the various numerical
 simulations of \DM\ halos and a consensus has been reached, the
-detailed shape of the inner density profile is still a hot topic of
+detailed shape of the inner density profile is still a matter of
 debate. The main point of contention is whether the density profile is
 cusped near the center of the \DM\ halo, and if so, how steep this
-cusp is, or whether the density stays finite near the center. Since
-this discussion has a non-negligible impact on the prospects for \DM\
-annihilation detection, we will briefly discuss the various plausible
-density profiles.
+cusp is, or whether the density asymptotes to a constant value at the
+center. Since this discussion has a non-negligible impact on the
+prospects for \DM\ annihilation detection, we will briefly discuss the
+various plausible density profiles.
 
@@
 \cite{2003MNRAS.338...14P}. Even the universality of the density
-profile was brought into question
+profile has been brought into question
 \cite{2000ApJ...529L..69J,2000ApJ...535...30J}.
@@
 
-Since the photon flux from \DM\ annihilations depends on the density
-squared of the \DM\, the exact form of the inner density profile has
-important ramifications for the detectability of \DM\
-annihilation. This has been widely discussed before in the context of
-various \DM\ annihilation models (\eg,
+The implications of the exact form of the inner density profile have
+been widely discussed before in the context of various \DM\
+annihilation models (\eg,
 \cite{1998APh.....9..137B,1999PhRvL..83.1719G,2000PhRvD..62l3005C,2000PhLB..494..181G,2002PhRvD..66b3509T,2002PhRvD..66l3502U,2003MNRAS.339..505T,2004ApJ...601...47A,2008JCAP...07..013B,2008arXiv0811.3744B,2008arXiv0812.3895B}).
@@
 form does \emph{not} include the exact form of Moore's profile; this
-need not concern us since inner cusps as steep as Moore's profile have
-been as good as excluded by the latest numerical simulations.
+need not concern us since inner cusps as steep as Moore's profile are
+excluded by the latest numerical simulations.
 
@@
 halo. For reasonable values of the inner slope this fraction is in the
-range 80 -- 95\%.
+range 80 to 95\%.
 
@@
 free-streaming scale, a \DM\ model-dependent quantity, which in many
-\DM\ models lies in the range 10$^{-6}$ -- 10$^{-12}$ \Msol\
+\DM\ models lies in the range 10$^{-6}$ to 10$^{-12}$ \Msol\
 \cite{1999PhRvD..59d3517S,2001PhRvD..64h3507H,2004MNRAS.353L..23G,2005PhRvD..71j3520L,2006PhRvL..97c1301P}. The
@@
 the Yukawa coupling parameter space this reduced boost is equal to one
-such that there is no extra boosts due to the lower velocity
+such that there is no extra boosts caused by the lower velocity
 dispersion in the substructures. This is mainly the case for large
@@
 \mv/\mdm and especially in resonance regions, the reduced boost can
-reach high values, signaling a large extra boost due to the
+reach high values, signaling a large extra boost caused by the
 kinematically coldness of the substructure.  For small values of
@@
 to two orders of magnitude larger than what we would expect from
-naively combining the Sommerfeld enhancement due to the velocity
+naively combining the Sommerfeld enhancement from the velocity
 dispersion of the subhalo with the boost from density enhancements
-alone. This can be easily understood as being due to the fact that the
-Sommerfeld enhancement has not saturated yet at the mass scale of the
-top-level subhalo, such that it can be boosted up to it saturation
-point by the contribution from the kinematically colder substructure.
+alone. This can be easily understood as being a consequence of the
+fact that the Sommerfeld enhancement has not saturated yet at the mass
+scale of the top-level subhalo, such that it can be boosted up to it
+saturation point by the contribution from the kinematically colder
+substructure.
 
@@
 boosts in different resonances seems to vary greatly, but this is
-mostly due to the finite resolution of this figure. Extra boosts up to
-$\sim 10^5$ are obtained in these resonances, and these extra boosts
-increase along the resonances when decreasing $\alpha$, again due to
-the higher saturation level for smaller $\mv/\mdm$ values.
+mostly caused by the finite resolution of this figure. Extra boosts up
+to $\sim 10^5$ are obtained in these resonances, and these extra
+boosts increase along the resonances when decreasing $\alpha$, again
+caused by the higher saturation level for smaller $\mv/\mdm$ values.
 
@@
 different inner slopes is irrelevant, and because the mass of the
-subhalo is only weakly affected by the value of the inner slope ---the
+subhalo is only weakly affected by the value of the inner slope---the
 function $f(c)$ is only a weak function of the inner slope. This is
@@
 this with the total boost shown in Figure \ref{fig:subboost}, it is
-clear that the inner slope does not affect the total boost due to
+clear that the inner slope does not affect the total boost caused by
 substructure, neither the overall magnitude of the effect nor the
@@
 such a shallow slope is to seriously reduce the substructure boost,
-which is simply due to the fact that for a shallower slope less
-substructure is found in smaller and more concentrated clumps. A
+which is simply a consequence of the fact that for a shallower slope
+less substructure is found in smaller and more concentrated clumps. A
 steeper slope such as $n = -1$, as was found in the recent \VL\
@@
 the \DM\ density profiles of the \dSphs\ were not completely
-determined by the available data ---a degeneracy between the
+determined by the available data---a degeneracy between the
 parameters of the \NFW\ profile used to describe the \DM\ density and
 the King profile used to fit the luminous component of the galaxies
-was found--- and the concentration-mass relation from \CDM\
+was found---and the concentration-mass relation from \CDM\
 \cite{2001MNRAS.321..559B,2001ApJ...554..114E} was used to break the
@@
 \cite{2004ApJ...613..962S}. We have integrated the extragalactic gamma
-ray flux as measured by EGRET over the energy range between 4 GeV --
+ray flux as measured by EGRET over the energy range between 4 GeV to
 250 GeV and multiplied it with the solid angle that the \dSphs\ span
@@
 \ref{fig:dsph}. This background flux is in all cases of the order of
-10$^{-9}$ -- 10$^{-10}$ photons cm$^{-2}$ s$^{-1}$, such that the \DM\
+10$^{-9}$ to 10$^{-10}$ photons cm$^{-2}$ s$^{-1}$, such that the \DM\
 annihilation boosted by substructure and Sommerfeld enhancement
@@
 halo \CDM\ predicts that the halo is filled with hundreds of
-individual subhalos ---a fact which gives rise to the so-called
+individual subhalos---a fact which gives rise to the so-called
 ``missing satellites problem''
@@
 for all three observers. We see that most of the subhalos have masses
-in the range 10$^4$ -- 10$^6$ \Msun, with subhalos with masses of the
+in the range 10$^4$ to 10$^6$ \Msun, with subhalos with masses of the
 order of the \dSphs\ being relatively rare. The lower limit is of
 course set by the finite numerical resolution of the simulation and
-doesn't correspond to a physical cut-off ---indeed, the physical
+doesn't correspond to a physical cut-off---indeed, the physical
 cut-off lies at the thermal free-streaming scale. The distribution of
 the distance to the subhalos shows that the distribution of the
-subhalos is strongly radially anti-biased, with a sharp cut-off due to
-our cut at the tidal radius of the main halo. This radial anti-bias is
-due to the fact that there is much more space far away from the center
-of the galaxy, and because of the strong tidal forces near the center
-of the galaxy, which strongly affects halos which pass near the center
-of the galaxy.
+subhalos is strongly radially anti-biased, with a sharp cut-off caused
+by our cut at the tidal radius of the main halo. This radial anti-bias
+is a consequence of the fact that there is much more space far away
+from the center of the galaxy, and because of the strong tidal forces
+near the center of the galaxy, which strongly affects halos which pass
+near the center of the galaxy.
 
@@
 \ref{fig:VL23_3}, \ie, \mv/\mdm\ = 10$^{-2}$ and $\alpha$ = 0.001. The
-left panel of Figure \ref{fig:VL23_2} shows the distribution of
+top panel of Figure \ref{fig:VL23_2} shows the distribution of
 gamma-ray fluxes from \DM\ annihilation for the subhalos, while the
-right panel shows the distribution of the total boost factors. As can
+bottom panel shows the distribution of the total boost factors. As can
 be seen, the total boost factors are quite small: the typical boost
@@
 the combination of the Sommerfeld enhancement and the substructure
-that is believed to exists down to small scales leads to very large
+that is believed to exist down to small scales leads to very large
 boost factors because of the combined effect of the higher densities
@@
 space are in no way unrealistic. Indeed, in one of the most
-straightforward interpretations of the ATIC/PAMELA results the sharp
+straightforward interpretations of the ATIC/PAMELA results, the sharp
 cut-off around 1 TeV seen in the ATIC spectrum is interpreted as being
-due to the fact that the dark matter mass is about 700 GeV. The
-absence of an excess in anti-protons similar to the excess in
+a consequence of the fact that the dark matter mass is about 700
+GeV. The absence of an excess in anti-protons similar to the excess in
 positrons seen by PAMELA can be explained by positing that the dark
@@
 We have not discussed the substructure boost to the \DM\ annihilation
-signal from unresolved substructure in the Galactic halo. This is,
-however, interesting in the light of the Sommerfeld enhancement, since
-it might lead to a different prediction about the relative magnitudes
-of the annihilation signal coming from the resolved Galactic halo and
-unresolved halo, as well from the resolved substructures in the
-Galactic halos. It is already clear that in the generic 1/$v$
-Sommerfeld enhancement the main halo, with its typical velocity
-dispersion of $\sim\!150$ km s$^{-1}$, will receive a much smaller
-enhancement than the \dSphs, with their typical velocity dispersions
-of 10 km s$^{-1}$. This, however, ignores the effects of saturation
-and resonances, which might be important in the specific particle
-physics model adopted for the Sommerfeld enhancement. Additionally,
-while the velocity dispersion of the Galactic halo might be of the
-order of 100 km s$^{-1}$, the actual velocity dispersion relevant to
-the Sommerfeld enhancement might be much smaller for halo \DM\
-particles. If a large fraction of the \DM\ in the Galactic halo is
-part of a cold stream ---and various numerical simulations predict
-that large parts of the halo, if not all of the halo, is made up of
-such cold streams with typical velocity dispersions of the order of 1
-km s$^{-1}$ \cite{1999MNRAS.307..495H,2003MNRAS.339..834H}--- then the
-relevant relative velocity of \DM\ particles is set by the velocity
-dispersion of particles in the stream, and not by the velocity
-dispersion of the streams in the halo, which is what the 100 km
-s$^{-1}$ Galactic halo velocity dispersion corresponds to in these
-scenarios. Thus, the expected \DM\ annihilation signal from the main
-halo could be much larger than that naively expected for a halo with a
-velocity dispersion of 100 km s$^{-1}$.
+signal from unresolved substructure, such as very small subhalos and
+\DM\ tidal debris from disrupted subhalos, in the Galactic halo. This
+is, however, interesting in the light of the Sommerfeld enhancement,
+since it might lead to a different prediction about the relative
+magnitudes of the annihilation signal coming from the resolved
+Galactic halo and unresolved halo, as well from the resolved
+substructures in the Galactic halos. It is already clear that in the
+generic 1/$v$ Sommerfeld enhancement the main halo, with its typical
+velocity dispersion of $\sim\!150$ km s$^{-1}$, will receive a much
+smaller enhancement than the \dSphs, with their typical velocity
+dispersions of 10 km s$^{-1}$. This, however, ignores the effects of
+saturation and resonances, which might be important in the specific
+particle physics model adopted for the Sommerfeld
+enhancement. Additionally, while the velocity dispersion of the
+Galactic halo might be of the order of 100 km s$^{-1}$, the actual
+velocity dispersion relevant to the Sommerfeld enhancement might be
+much smaller for halo \DM\ particles. If a large fraction of the \DM\
+in the Galactic halo is part of a cold stream---and various numerical
+simulations predict that large parts of the halo, if not all of the
+halo, are made up of such cold streams with typical velocity
+dispersions of the order of 1 km s$^{-1}$
+\cite{1999MNRAS.307..495H,2003MNRAS.339..834H}---then, in addition to
+any boost caused by density enhancements in the stream
+\cite{2008arXiv0811.1582A}, the boost from the Sommerfeld enhancement
+could also play a large role, as the relevant relative velocity of
+\DM\ particles is set by the velocity dispersion of particles in the
+stream, and not by the velocity dispersion of the streams in the halo,
+which is what the 100 km s$^{-1}$ Galactic halo velocity dispersion
+corresponds to in these scenarios. Thus, the expected \DM\
+annihilation signal from the main halo could be much larger than that
+naively expected for a halo with a velocity dispersion of 100 km
+s$^{-1}$.
 
